### 2014-10-20 SI.tex
First commit

%\begin{description}

%\vspace{1ex} \includegraphics[width=0.7\textwidth]{Figs/e_fig7s-crop.pdf}\\
%{\bf Figure \SFpef. Performance of shallow and deep networks.}
%
%\vspace{2ex} \includegraphics[width=0.7\textwidth]{Figs/e_fig4b.pdf}\\
%{\bf Figure \SFexp. Explanation power vs.~performance rank.} $P < 0.001$. %Enter optional descriptive information here.
%
%\vspace{2ex} \includegraphics[width=0.7\textwidth]{Figs/e_fig5d.pdf}\\
%{\bf Figure \SFaln. Invariance and selectivity subspace alignment vs.~performance rank.} Both $P < 0.001$.
%
%\vspace{2ex} \includegraphics[width=0.7\textwidth]{Figs/e_fig5b.pdf}\\
%{\bf Figure \SFinc. Invariance subspace capacity vs.~performance rank.} $P < 0.001$.
%
%\vspace{2ex} \includegraphics[width=0.7\textwidth]{Figs/e_fig5a.pdf}\\
%{\bf Figure \SFfda. Fitness-distance diagram.} $P$ value? %Corr of Random Walk, Multi Corr after adding that? why r is low? kernel selection.
%
%\vspace{2ex} \includegraphics[width=0.7\textwidth]{Figs/e_fig6b.pdf}\\
%{\bf Figure \SFenc. Encoding specificity vs.~performance rank.} $P < 0.001$.

%\item{\bf Figure S0. All OS.}
%\item\bf Figure S6. All corr plot.} PARTIAL correlation!?
%\item{\bf Text S1. LFW EXP DETAILS (FT DIM, SVM SETTING)} retinotopic. %latex textblock?
%\item{\bf Text S2. Details of numerical optimzation (invariance subspace; encoding specificity).} %enc spc: regularized and more structural, especially important for highly noise tolerant networks; multi-resolution
% $\lambda\left\|\nabla^{2}x\right\|_{F}$

%\item{\bf Text S1. Details of numerical optimization (invariance subspace; encoding specificity).}
%\item enc spc: regularized and more structural, especially important for highly noise tolerant networks; multi-resolution

%\end{description}

\newcommand{\beginsupplement}{\setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
\beginsupplement

\begin{figure}[H]
\centering \includegraphics[width=0.6\textwidth]{Figs/e_fig7s-crop.pdf} 
\caption{ 
{\bf Distributions of shallow and deep network performances.} Performances of the randomly generated 100 shallow and 100 deep networks are obtained via 10-fold cross validations of 6,000 face image pairs. Feature vectors used in the linear SVM classifier are generated via sliding the receptive fileds throughout entire images and concatenating the population representations (as defined in convolutional neural networks). Maximum and mean of performances of the deep networks are both significantly higher than those of the shallow networks (with permutation tests, $p<0.001$ and $p=0.008$ respectively). See \cite{cox2011beyond} for more details.}
\label{fig:SFpef}
\end{figure}

\begin{figure}[H]
\centering \includegraphics[width=0.6\textwidth]{Figs/e_fig4b.pdf} 
\caption{ 
{\bf Explanation power vs.~performances of deep networks.} Each distribution from left to right shows the top 1\% explainabilities of the task-related stimuli within each of the 32 top-layer unit representations of a deep network. Performance ranks, instead of performance values, are used for visualization purposes. Means of distributions are plotted as gray circles and the linear regression as dashed black line. Significance of slope using permutation test has $p < 0.001$.}
\label{fig:SFexp}
\end{figure}

\begin{figure}[H]
\centering \includegraphics[width=0.6\textwidth]{Figs/e_fig5d.pdf}
\caption{ 
{\bf Invariance and selectivity subspace alignments vs.~performances of deep networks.} Each red distribution from left to right shows the invariance subspace alignments against the task-related stimuli of population representations of a deep network given the 16 reference stimuli. Alignment scores are normalized by subtracting the corresponding reference stimuli's own alignment scores. Performance ranks, instead of performance values, are used for visualization purposes. Means of distributions are plotted as red circles and the linear regression as dashed red line. Blue distributions for selectivity subspace alignments follow the same definitions. Significances of slopes using permutation tests both have $p < 0.001$.}
\label{fig:SFaln}
\end{figure}

\begin{figure}[H]
\centering \includegraphics[width=0.6\textwidth]{Figs/e_fig5b.pdf}
\caption{ 
{\bf Invariance subspace capacities vs.~performances of deep networks.} Each distribution from left to right shows the invariance subspace capacities of population representations of a deep network given the 16 reference stimuli. Performance ranks, instead of performance values, are used for visualization purposes. Means of distributions are plotted as gray circles and the linear regression as dashed black line. Significance of slope using permutation test has $p < 0.001$.}
\label{fig:SFinc}
\end{figure}

\begin{figure}[H]
\centering \includegraphics[width=0.6\textwidth]{Figs/e_fig5a.pdf}
\caption{ 
{\bf Fitness-distance diagram of population representations within deep networks.} Red, blue, and green dots indicate results of invariance and selectivity path searches, and random walks, where brighter shades of a color are of better performing networks, and 
darker shades of poor performing networks. Means of results are plotted as solid lines in corresponding colors. As visualized, there is no apparent strong correlations between invariance and selectivity path potentials and performance. Random walk results vs.~performance, on the other hand, has $R^2 = 0.18$ correlation which may arise from the differences in sensitivities to noises, though incorporating it into representation measures does not improve the final multiple correlation.}
\label{fig:SFfda}
\end{figure}

\begin{figure}[H]
\centering \includegraphics[width=0.6\textwidth]{Figs/e_fig6b.pdf}
\caption{ 
{\bf Encoding specificities vs.~performances of deep networks.} Each distribution from left to right shows the SSIM scores of the reconstructed stimuli of population representations of a deep network against the 16 reference stimuli. Performance ranks, instead of performance values, are used for visualization purposes. Means of distributions are plotted as gray circles and the linear regression as dashed black line. Significance of slope using permutation test has $p < 0.001$.}
\label{fig:SFenc}
\end{figure}

### 2014-10-20 discussion.tex
First commit

%why depth? bad about universal approx theorem? 
%L2 distance vs. cosine distance? 
%v1, poor selectivity, large size?

\subsection*{Why Do We Need a New Framework?} % NOVELTY OF METHODS

As aforementioned, the study of sensory representations of biological neural networks has been an extremely important research topic for decades and numerous methods have been proposed; nevertheless, most methods can be viewed as implementation instances of functional frameworks of certain generalized concepts \cite{wu2006complete, dimattina2013adaptive}. We argue than existing frameworks being actively used are still mostly highly parametric, in terms of both the model assumption and stimulus generation. As moving away from peripheral or primary sensory circuitry and going deeper into higher-level sensory cortices, the modeling power of parametric models becomes inevitably insufficient, and even though parametric stimulus may facilitate the speed of representation characterization, its inherently assumptive nature can strongly bias the results. Although there also exist approaches which utilize the nonparametric artificial neural networks as models of sensory circuitry, only usage of shallow networks has been reported \cite{lau2002computational, prenger2004nonlinear}, and interpreting the resultant model remains difficult and largely approximate \cite{wu2006complete}. In fact, this also directly links to the fact that attempts of characterizing/interpreting artificial neural networks themselves are also relatively difficult and scarce, and sometimes algorithms using handcrafted representations (e.g.~HoG and SIFT \cite{szeliski2010computer}) are still preferred for their interpretability. Such problems only get worse when deeper networks are being adopted (either as system identification models of biological circuitry, or as algorithms directly applied in tasks), since understanding them also becomes even harder. Another important issue is the lack of principled and efficient approach for characterizing neuronal populations, due to the fact that, e.g., there is no straightforward generalization of the definition of optimal stimulus to neural populations, and existing methods can suffer from the exponential growth of required measurements, with respect to the population size \cite{dimattina2013adaptive}.

Major benefits of the proposed framework can be categorized as follows. (1) Generality: as aforementioned, the proposed framework tackles the challenges in studying sensory representation in a highly generalizable and unbiased way via supporting both nonparametric models and nonparametric stimuli. Although not explicitly tested in this work, such setting in practice can be easily reconfigured and extended, if desired, to support both parametric models, by numerically analyzing them as nonparametric models (especially when analytical solutions are not easily derivable), and parametric stimuli, by simply reframing the optimization in the parameter space $p \in \mathbb{R}^P$ and evaluating the fitness through the generative function $f\left(x\left(p\right)\right)$ (especially when certain stimulus type or deformation is confirmed to be highly relevant). Also, the proposed framework supports characterizing both unit and population representations, as described in \nameref{sec:methods}; the fitness-distance diagram can be viewed as a ``generalized tuning curve'' characterization of the landscape, which is also dimensionality- and population size-insensitive. (2) Efficiency: the proposed framework is designed to be efficient using low-rank and stochastic approximations of the complex high-dimensional tuning landscape and its properties. As reported in multiple existing works \cite{touryan2002isolation, rust2004spike} which utilized second-order model for system identification and representation characterization, often only eigenvectors corresponding to few of the most positive and negative eigenvalues have significant and interpretable structures. Directly searching for invariance and selectivity paths eases this problem and saves potentially costly measurement requirement, which is on the order of $\mathcal{O}\left(N^2\right)$, to $\mathcal{O}\left(N\right)$, in addition to the fact that non-local characterization can be obtained in this work as well. Stochastic sampling of the solution space of the invariance subspace capacity provides an efficient approximation of the intrinsic dimension estimation approach \cite{kegl2002intrinsic, levina2004maximum}, and, of the encoding specificity, an efficient alternative of the maximally informative ensemble estimation approach \cite{machens2002adaptive}, both of which can require substantially more measurements in order to yield robust and meaningful results. As aforementioned, the back-end solver with relatively simple constraints also positively contributes to the speed of this framework. %sharpee2004analyzing

\subsection*{How Close Are We to Fully Understanding Deep Networks?} % CONTRIBUTION/MEANING OF FINDINGS, REPRESENTATION-WISE

Via the proposed numerical optimization framework, the following important aspects of sensory representation research are enabled or facilitated: (1) Decoding unit or population representation (as optimal or reconstructed stimulus). (2) Discovering what changes in stimulus the representation is invariant or selective to. (3) Explaining how deep network's representation may affect its task performances. Main findings of this work can be summarized as:
\begin{itemize} % Main Findings
\item Complexity of representation increases along network's depth.
\item Unlike deep representation which is both invariant and selective, shallow representation is only invariant, not selective, and its capacity of invariance is significantly lower than deep representation's.
\item How well the optimal stimulus can explain task-related stimuli, and how specific the representation encoding of task-related stimuli is, both decently explain network's task performance.
\item How well the invariance and selectivity of representation align with the actual ``distribution'' of task-related stimuli also partially explains network's task performance.
\end{itemize}
We argue the importance of these findings as follows. First, the experiments are conducted in large scale and all results are statistically significant. Second, the results are obtained with relatively unbiased methods (i.e.~nonparametric model plus nonparametric stimulus). Third, the representation measures match what are known to be crucial properties of performing accurate visual recognition, and explain network's performance decently. Finally, the results also point out the practicability of the relatively less explored paradigm, where not only population representation can be examined, but extrinsic properties (i.e.~measures based on task-related stimuli) also can be integrated to more effectively characterize various aspects of the deep network, in addition to only using intrinsic properties (i.e.~measures based on optimal stimuli). In fact, if just using the four intrinsic properties (i.e. OSSC, INPP, SLPP and INSC as shown in Fig.~\ref{fig:ind_mea}; not included in Fig.~\ref{fig:all_R^2}) of the unit representations of the deep networks (with measures averaged from all 32 top-layer neurons within a network), we can only explain the network's performance with $R^2 = 0.34$, which is significantly lower than the correlation based on extrinsic properties (i.e.~$R^2 = 0.71$; combining both properties only marginally increase $R^2$ to 0.73). This also corresponds to our observation that the ranking of deep networks may differ from task to task (e.g.~face pair matching in this work vs.~object recognition in \cite{pinto2009high}), and the fact that neural circuitry can (and maybe constantly) dynamically shape its tuning (i.e.~representation) according to the tasks being performed \cite{gilbert2007brain} such that ``intrinsic'' properties may not mean much. Also, combining the results from machine learning and computer vision studies about deep networks' advanced theoretical efficiency \cite{delalleau2011shallow, montufar2014number} and practical performances over shallow networks', we argue that network depth is possibly the most plausible way of implementing efficient sensory processing.

% finding not just by unit or intrinsic. [intriguing,malik]

However, there are still undeniably plenty of unsolved and even unaddressed mysteries about deep networks. Here we list the ones we think to be of the most importance. First, even though we have successfully visualized the representations of certain deep neurons, as what is argued in \nameref{sec:results} about the optimal construction of tuning landscapes, how these representations explicitly facilitate generalization (i.e.~invariance) across features that are distinct but of the same class, and discrimination (i.e.~selectivity) between features that are similar but of different classes, remains ambiguous. Second, even though some smaller-scale studies using heuristic evaluation \cite{zeiler2014visualizing} or theoretical characterization \cite{saxe2013exact} of representations' evolution have been conducted, how the training algorithms (e.g.~backpropagation) explicitly (and orientedly) shape the representations, and how different training settings alter the resultant representations, remain largely unclear. Third, the ability of ``bottom-up attention'' seems to be prominently functioning in these feedforward deep networks \cite{zeiler2014visualizing, simonyan2013deep} for unaligned and mildly cluttered visual object recognition; even though pooling operations are generally considered capable of providing such function, to what degree it (or other operations) contributes to this, how it explicitly (or implicitly) resolves, e.g., objects' labels from cluttered scenes, and how much variation can be tolerated, remain unclear. Finally, some ``higher-level vision'' tasks also have been reported working decently using deep networks (e.g.~image style recognition \cite{karayev2013recognizing} and scene understanding \cite{CVPR14_Khosla}), in which what might the representations being used are can be even more ambiguous and undefined, and more ``holistic'', instead of part-based, representations, are likely to dominate. As argued in \cite{cox2014we}, object recognition alone is a limited operational definition of high-level vision, and may not even be the right framing problem for studying high-level vision and visual representations any more. Like \cite{olshausen2005close}, we start this subsection by asking how close we are to solving this puzzle, yet with all the remaining inexplicable properties of deep networks, maybe asking how far we are from solving this, and how to precisely address these mysteries, would be more important and fruitful questions to be asked.

\subsection*{Challenges and Possible Directions}

In addition to the inexplicable properties of deep networks listed above, which still lack clearer definitions and directions, here in the following, with no specific order, we also give a list of more defined challenges that we will need to solve, and possible directions to specifically tackle the challenges. Arguably, sensory representations of the most interest are at neither the bottom layer nor the top layer, but intermediate layers, since the former is relatively well understood and usually consist of Gabor (or Gabor-like) filters, and the later is highly class selective, as directly suggested by the ``behavioral'' (i.e.~classification) results and the visualizations shown in \cite{donahue2014decaf, azizpour2014generic}. However, for most advanced deep networks \cite{szegedy2014going, simonyan2014very} which can have around 20 layers of operations, even intermediate layers can be very computationally expensive to measure and may have way more complex tuning landscapes (compared to the two-level deep networks experimented in this work, which consist of up to 8 layers of operations). Instead of CPU-based deep network simulation adopted in this work, GPU- or FPGA-/ASIC-based simulation may be required to tackle the potential speed problem. The back-end numerical solver used in this work may not suffice the needs for even deeper networks either; nevertheless, the proposed framework itself is neutral to the selection of solvers, thus more advanced numerical optimization algorithms can always be adopted to improve the quality of numerical solutions. Similar challenges may be encountered when moving from purely feedforward architectures into more biologically plausible ones---networks with feedback pathways, or recurrent networks---for their widely known nature of complex dynamics. Though in this work, around 70\% of the variance of performance differences on face pair matching can be explained with the proposed representation measures, yet we are not clear how high this number can be driven, and if these measures can generalize well enough onto even deeper representations or radically different types of tasks (e.g.~phoneme recognition). Therefore, the validation/search of existing/new representation measures shall still be an urgent direction for future research.

Another difficult yet potentially fruitful future challenge is to apply this framework to real biological deep networks. While results are exclusively shown for artificial networks, the methods presented in this work by nature require a small budget of neuronal measurements that they could potentially be applied to real neurons in real experiments, even when spiking variability is taken into account, given that at least two effective techniques---stochastic averaging and measurement reweighting---are built in the core of CMA-ES, the back-end solver, for robustness against measurement noises. In fact, we argue that the mathematical formulation of CMA-ES can be viewed as an extension/generalization of the spike-trigger covariance method, which is known to be working effectively for biological neurons \cite{rust2004spike}. Potential inefficiency of measurement arising from neural adaptation \cite{kohn2007visual} can also benefit from, e.g., Particle Swarm CMA-ES \cite{muller2009particle} or other multipopulation/multimodal optimization methods, where evaluation of stimuli from different local searches can be interwoven to minimize neural adaption and maintain search speed. Overall, as aforementioned, since the proposed framework supports efficient nonparametric search and analysis, a biological neural network can be characterized \emph{as is} to reduce the chances of obtaining premature or biased numerical results. On the other hand, based upon recently reported results of deep networks being able to produce representations similar to mammalian visual systems' \cite{yamins2014performance, cadieu2014deep}, which though may still seem incomplete, we argue that the possibility of deep networks being a good surrogate model for understanding biologically plausible ways of representation encoding, and even guiding the research direction of ventral visual pathway studies, should not be neglected. In addition to the tools proposed in this work, neurophysiological methods like patch clamping, lesioning, ``gene'' (i.e.~hyperparameters) knockout/knockdown, etc.~may be as well used to inspire the development of new tools for characterizing artificial deep networks. For instance, the effect of surround suppression is also observed in experimented artificial neurons in this work as visualized in Fig.~\ref{fig:ind_res} (i.e.~flat gray peripheries in the optimal stimulus patterns; responses drop when manually place, e.g., dots in peripheries) and, in our preliminary tests, can be easily shutdown via ``lesioning'' the prenormalization operations (i.e.~early-stage lateral inhibition) of the network, which is consistent with neurophysiological findings \cite{cavanaugh2002selectivity}.

Finally, though via the proposed numerical framework, we have demonstrated how previously unknown or unconfirmed properties can be positively identified with high statistical significance, we know this does not entirely cover what will be needed to ``fully'' explain observed properties and ultimately everything about sensory representations within deep networks---a theoretical framework yet to be constructed/completed. Notable recent works on bottom-up construction of theoretical frameworks for feedforward convolutional deep networks include magic theory \cite{anselmi2013unsupervised}, scattering transform \cite{mallat2012group}, convolutional kernel network \cite{MairalKHS14}, etc. Saxe et al.~\cite{saxe2011random} and Szegedy et al.~\cite{szegedy2013intriguing} also addressed particular phenomena observed in deep networks using top-down theoretical approaches, furthermore to the theoretical studies \cite{delalleau2011shallow, montufar2014number} on deep network's properties mentioned above. While undoubtedly all different theoretical approaches should keep being actively researched and interactively inspired by each other, we argue that the proposed numerical framework can nicely function as an independent \emph{diagnostic tool} for examining, verifying, and facilitating the construction of theories.

### 2014-10-20 introduction.tex
First commit

The mammalian ventral visual pathway consists of a hierarchical cascade of visual areas that progressively reformat visual information from pixel-like retinotopic representations into formats suitable for high-level tasks \cite{dicarlo2007untangling, dicarlo2012does}, such as recognizing visual objects. Recently, computer vision algorithms inspired by the hierarchical structure of the visual system---convolutional neural networks \cite{fukushima1980neocognitron, lecun1998gradient, riesenhuber1999hierarchical}, or so-called ``deep learning'' approaches \cite{krizhevsky2012imagenet}---have become the focus of tremendous attention in the computer vision and machine learning communities, due to their success in a variety of practical domains. However, we are still far from achieving human-level capabilities for unconstrained visual tasks \cite{ghodrati2014feedforward}, and even for the most successful current deep learning systems which, on controlled visual tasks, perform relatively close to human subjects \cite{serre2007feedforward, cirecsan2012multi, russakovsky2014imagenet, taigman2013deepface, sun2014deep, viglarge} and produce representations similar to mammalian visual systems' \cite{yamins2014performance, cadieu2014deep}, inexplicable failures still occur \cite{szegedy2013intriguing}. Undeniably, we lack a comprehensive functional understanding of how these system work at the theoretical level. Here, we take the success of artificial ``deep learning'' architectures as an opportunity to develop tools for studying and gaining insight about the function of deep networks of nonlinear units---tools that can in principle be used to study both artificial and real neuronal systems. 
%MORE REF FOR DNN'S PROBLEM??
% WHAT'S DEEP NET, DEEP NEURON? WHY IMPORTANT

%\suppressfloats
%\begin{table}
%\caption{\bf{Definition of mathematical terms.}} %Glossary List of recurring
%%\small
%\begin{tabular}{ll}
%$N$ & Receptive field size, which is organized as, e.g., $\sqrt{N}\times\sqrt{N}$ for static visual stimulus. \\
%$x$ & Stimulus, \\
%$\arg\max$ & optimizing a target fitness function \\
%$x^+_{\delta} x^-_{\delta}$ & \\
%$\| \cdot \|$ & \\
%$\| \cdot \|_{F}$ & \\
%$\| \cdot \|_{*}$ & \\
%$\| \cdot \|$ & \\
%$\| \cdot \|_{F}$ & \\
%$\| \cdot \|_{*}$ & \\
%$\| \cdot \|$ & \\
%$\| \cdot \|_{F}$ & \\
%$\| \cdot \|_{*}$ & \\
%\end{tabular}
%%\begin{flushleft} Table caption \end{flushleft}
%\label{tab:label}
%\end{table}

Explaining the sensory representations of neurons, one of the most important aspects of studying the underpinnings of sensory processing circuitry, has been a widely researched topic in both theoretical and experimental neuroscience, where artificial and biological neural networks are mainly used and often cross-studied for developing better explanations of sensory processing mechanisms. Brief reviews of methods used in both fields are as follows. In theoretical neuroscience studies, methods can be categorized by the artificial models being used \cite{wu2006complete}: parametric or nonparametric (which are usually analytically or only numerically analyzable). Berkes et al.~\cite{berkes2006analysis} studied quadratic networks, $f\left(x\right) = \frac{1}{2}x^{T}Qx+L^{T}x+c$, in which the optimal stimulus uniquely exists and can be efficiently computed, and local invariance and selectivity directions can be analytically derived through eigendecomposition of the symmetric quadratic term $Q$ (i.e. the Hessian), which are eigenvectors of the least and most negative eigenvalues. Saxe et al.~\cite{saxe2011random} studied single-level convolutional networks $f(x) = \left\| K \otimes x \right\|_{F}$ which can be viewed as the building block of deep learning networks, and showed the optimal stimulus can be analytically approximated using a Gabor-like filter with its frequency component collocated with the peak of the Fourier spectrum of convolution kernel $K$, whether $K$ itself is structural or random. Zeiler et al.~\cite{zeiler2014visualizing} and Simonyan et al.~\cite{simonyan2013deep} studied multi-layer convolutional networks (i.e.~deep learning networks), numerically derived and approximated the optimal stimuli of neurons of various depths, and qualitatively showed in deeper layers, neurons are tuned to gradually more complex visual patterns. Zeiler et al.~\cite{zeiler2014visualizing} also used parametric deformations (translating, rotating, and scaling) to test the invariance of neurons as in \cite{goodfellow2009measuring}. Although methods in \cite{zeiler2014visualizing, simonyan2013deep} didn't explicitly require the network being analytical, the capability of the network being able to perform backpropagation is however needed. Le et al.~\cite{ngiam2010tiled} extended the method in \cite{berkes2006analysis} onto multi-layer convolutional networks through numerically estimating the optimal stimulus and the Hessian. Le et al.~\cite{le2012building} also visualized the optimal stimulus of multi-layer convolutional networks, but unlike \cite{zeiler2014visualizing, simonyan2013deep}, treated the network as non-analytical black box. Erhan et al.~\cite{erhan2010understanding} studied multi-layer networks and proposed to characterize invariance in a larger extent, through numerically searching for non-local (i.e.~farther to the optimal stimulus) solutions, instead of only locally approximating it through Hessian decomposition. Szegedy et al.~\cite{szegedy2013intriguing} also used similar approach to study the selectivity of top-layer neurons inside multi-layer convolutional networks. In experimental neuroscience studies, methods can be categorized by the stimulus used to characterize the biological---naturally nonparametric---models \cite{wu2006complete}: parametric or nonparametric. An $N$ dimensional parametric stimulus $x$ can be either functionally generated through a $P$ dimensional controlling parameter $p$, i.e.~$x\left(p\right) \in \mathbb{R}^N$ and $p \in \mathbb{R}^P$ where $P < N$, or sampled from a stimulus dictionary of limited size that mostly only spans part of the $N$ dimensional space \cite{field1987relations}, and e.g., moving bars, sine gratings, natural images, etc.~fall into this category. It is an efficient and commonly adopted stimulus for studying the sensory representation of neurons in various stages along the sensory pathways, and can be coupled with characterization procedures ranging from closed-loop methods like genetic algorithm \cite{bleeck2003using, yamane2008neural}, to open-loop methods like spike-triggered average (or reverse correlation) \cite{ringach2004reverse, hansen2004parametric, dotsch2012reverse} and Bayesian methods \cite{naselaris2009bayesian, nishimoto2011reconstructing}. Contrarily, an $N$ dimensional nonparametric stimulus $x \in \mathbb{R}^N$ can have all its variables valued independently and thus spans the entire $N$ dimensional space. White noise is the most commonly used modality, which can be coupled with characterization procedures ranging from closed-loop methods like hill climbing \cite{harth1974alopex}, to open-loop methods like spike-triggered covariance \cite{touryan2002isolation, rust2004spike} and clearly spike-triggered average as well \cite{ringach2004reverse}. Methods focusing on closed-loop characterization are also reviewed and summarized by DiMattina et al.~in \cite{dimattina2013adaptive}. %ISO RESPONSE?

%brenner2000adaptive
%or sampled from a stimulus dictionary $D$ of limited size $S$ that mostly only spans part of the $N$ dimensional space \cite{field1987relations}, [i.e.~$x \sim \left\lbrace D_i \right\rbrace$ where $D \in \mathbb{R}^{N \times S}$ and $\mathrm{rank}\left(D\right) < N$]

In this work, we propose a \emph{unified framework} based on modern numerical optimization techniques to help us uncover the sensory representation encoded by neurons deep inside neural networks, where the most flexible and generalizable settings---nonparametric network model with nonparametric stimulus---are supported to maximize the applicability of this framework, with the inherent difficulties and inefficiency resolved by carefully constructing and constraining the numerical framework. Characterization of the ``tuning landscape'' of neurons to the high-dimensional nonparametric stimulus includes both first- and quasi-second-order ``landscape features'', i.e.~optimal stimulus and its invariance and selectivity directions---the most significant structural features of the surrounding landscape---where the inefficiency of modeling the entire Hessian can be avoided. Through incorporating multiple randomized searches, the invariance and selectivity subspaces can be efficiently estimated as well. Representation measures for analyzing of numerical searching results are also designed to perform dimensionality-insensitive characterization of deep networks. Using the proposed framework for sensory representation characterization, we directly address two important questions: (1) Why are deep networks better than shallow networks? (2) Among experimented networks of the same depth, why are certain networks better than others? %%Interesting findings include, ...

### 2014-10-20 methods.tex
First commit

\begin{figure}
\centering \includegraphics[width=0.5\textwidth]{Figs/concept.pdf} 
\caption{ 
{\bf Schematic of numerical optimization framework.} Deep network $f$: hierarchical cascade of layers (shaded tall rectangles) of artificial neurons (circles) transforming the stimulus $x$ (bottom layer, leftmost) into high-level representations $r$ (top layer, rightmost) suitable for linear classification. Optimization framework: closed-loop algorithm that repeatedly measures the response(s) of one or multiple neuron(s), calculate the fitness, and adaptively generate stimuli to characterize the target neuron(s) via optimizing the fitness functions defined for different purposes (see Eq.~(\ref{eq:O1}--\ref{eq:S2})); when characterizing multiple neurons for population representation of given optimal or reference stimulus, response of an imaginary neuron (dashed circle) tuned to the given stimulus is optimized (see Eq.~(\ref{eq:O2}, \ref{eq:I2}, \ref{eq:S2})) equivalently.}
\label{fig:concept}
\end{figure}

Figure \ref{fig:concept} illustrates and defines the setup of the closed-loop numerical optimization framework for studying sensory representations encoded within the deep network of interest. The proposed framework consists of two main methods: (1) optimal stimulus search and (2) invariance/selectivity path search, or the first-order and quasi-second-order characterization methods, which equivalently correspond to the linear analysis and low-rank quadratic analysis (i.e.~only eigenvectors of the largest and smallest eigenvalues considered) respectively. Representation $r$, in this work, is equivalent to the response(s) of artificial neuron(s), and is denoted as unit representation (or scalar representation, i.e.~$r \in \mathbb{R}$) when referring to single neurons and population representation (or vector representation, i.e.~$r \in \mathbb{R}^R$) when referring to groups of neurons (with group size $R$). Arguably, for studying both artificial and biological neural networks, especially when evaluating task-related performances, the ability of characterizing population representations is usually more important than of characterizing unit representations, considering the differences in representational powers. This framework supports characterizing both unit (using Eq.~(\ref{eq:O1}, \ref{eq:I1}, \ref{eq:S1}), though Eq.(\ref{eq:O2}, \ref{eq:I2}, \ref{eq:S2}) can be used as well) and population (using Eq.(\ref{eq:O2}, \ref{eq:I2}, \ref{eq:S2})) representations in a unified way.

As aforementioned, this framework also treats network as non-analytical black boxes and poses no assumption about networks' properties. Nevertheless, all stimuli used in this work follow the constant energy constraint $\left\| x \right\| = E$, as suggested by previous studies where simply enhancing the contrast of a fixed stimulus can as well cause increased firing rates \cite{albrecht1982striate, cheng1994comparison}, which in most cases leads to meaningless outcomes since the spatial ``patterns'' of stimulus are not changed. For simplicity of mathematical formulation, the setting of $E=1$ is used for the rest of the paper, while in experiments $E$ is set to the average of task-related stimulus energy before normalization. In both methods, numerical optimization are executed multiple times from randomized initial settings, for two main reasons: (1) increasing robustness of the numerical solutions of, e.g., optimal stimulus and invariance/selectivity path searches (2 runs executed), and (2) providing statistical samplings of the solution spaces of, e.g., encoding specificity and invariance/selectivity subspace searches (10 and 20 runs executed). The phase of optimal stimulus search may be omitted when studying the invariance, selectivity and encoding specificities of representations of given reference stimuli.

%(STATE LOCAL VS GLOBAL OPTIMAL, WHY NAMED FIRST, Q-SECOND, WHY IN/SL ARE IMPORTANT)

\subsection*{First-Order Characterization: Optimal Stimulus Search and Analysis}

\begin{figure}
\centering \includegraphics[width=0.5\textwidth]{Figs/methods.pdf} 
\caption{
{\bf Visualization of numerical optimization procedures.} Spherical constraint: solution space of the stimulus with energy constraint $\left\| x \right\| = 1$, an $N$ dimensional sphere centered on the origin. Optimal stimulus search trajectory: best solutions (patterns with black frame boarders) from sequential search iterations with adaptive landscape modeling (gray eclipses with varying diameters); visualization following \cite{muller2010black}. Distance constraints (gray circles): solution spaces of stimuli with $0.1\pi$ to $0.5\pi$ distances from the optimal stimulus. Invariance path (dashed red curve): solutions of the invariance path searches. Selectivity path (dashed blue curve): solutions of the selectivity path searches. Target neuron in this visualization is optimally tuned to a $45^{\circ}$ Gabor pattern, invariant to phase changes, and selective to orientation changes.}
\label{fig:methods}
\end{figure}

In this work, optimal stimulus is numerically derived through the iterative optimization as \begin{equation} \label{eq:O1} \hat{x} = \underset{x} {\arg\max} f\left(x\right) = \underset{x_{g} \in \Omega \left( x_{g-1} , m_{g-1} \right)}{\arg\max} f\left(x_{g}\right),\end{equation} where $m_{g}=U\left(\left\lbrace f\left(x_{g}\right) \right\rbrace , m_{g-1}\right)$ and $1 \le g \le G$, subject to $\left\| x \right\| = 1$. The algorithm starts from random initial point $x_{0}$ on an $N$ dimensional sphere, as the noisy pattern shown in the bottom of Fig.~\ref{fig:methods}, and samples a set of $\lambda$ neighboring points (through function $\Omega$) as candidates to be evaluated, from an initial model $m_{0}$ with null distribution (in this work, multivariate Gaussian distribution starting with covariance matrix $\Sigma = I$). Through measuring the fitnesses of sampled points $\left\lbrace f\left(x_{g}\right) \right\rbrace$, the algorithm updates its model $m_{g}$ (i.e.~mean and covariance of the distribution, through function $U$) for generating samples in the following iterations, where the same operations repeat. Usually the maximally allowed iteration number $G$ is set to prevent the algorithm from running unreasonably long; nevertheless, early termination is also practically possible when optimality is reached and the neighborhood size shrinks below threshold, as shown in the center of Fig.~\ref{fig:methods}. Though the fitness landscape can be highly nonlinear, such sequential optimization procedure is usually capable of gradually accumulating knowledge of the landscape and adjusting its search directions to climb onto higher fitness areas, and the smoother the landscape is, the faster the algorithm converges. As exemplified in Fig.~\ref{fig:methods}, the ``most salient pattern'', or direction leading to high fitness areas usually can be rapidly extracted (see Fig.~\ref{fig:ind_res} for more examples). In most scenarios, fixing the maximally allowed number of fitness evaluations $\lambda G$ to, e.g., $100N$ (as adopted for the optimal stimulus search in this work) can lead to reasonably good convergence speed vs.~quality trade-off. The aforementioned sequential optimization concept is in fact commonly adopted in modern stochastic optimization \cite{spall2005introduction}, and the CMA-ES (Covariance Matrix Adaptation Evolution Strategy) algorithm \cite{hansen2001completely} is chosen as the back-end solver of this numerical framework, for its decent convergence speed and capability of handling rugged landscapes. Readers can refer to \cite{hansen2001completely} for detailed definitions of model $m$ and functions $\Omega$ and $U$. The energy constraint of stimulus is simply handled through the spherical projection before evaluation, i.e.~$f\left(p_s\left(x\right)\right)$ where $p_{s}\left(x\right) = {x} / {\left\|x\right\|}$, while the solver works unconstrainedly.

For analyzing the results of optimal stimulus searches, the following measures are adopted: (1) \emph{Spectral complexity}, which is estimated through the $L^{1}$ norm of the (2 dimensional) Fourier power spectrum of optimal stimulus, i.e.~$\left\| \mathcal{F}(\hat{x}) \right\|_{1}$, where higher value suggests higher non-sparsity (i.e.~more spectral components required to represent the signal). (2) \emph{Explanation power}, which is estimated as the mean of the linear explainability of task-related stimuli---rectified inner-product distances between a set of $n$ task-related stimuli $\left\lbrace x^{t} \right\rbrace$ and the optimal stimulus itself, i.e.~$\frac{1}{n}\sum_{i=1}^{n}\max\left(\langle x^{t}_{i} , \hat{x} \rangle , 0\right)$---and measures how well the optimal stimulus can linearly ``approximate'' task-related stimuli (or intuitively how much neuronal response might be elicited). (3) \emph{Encoding specificity}, which utilizes the optimal stimulus search as an inverse function $f^{-1}$ of a representation $r$ (i.e.~searching for the optimal stimulus of an imaginary neuron, as shown in Fig.~\ref{fig:concept}, tuned to the representation of a reference stimulus) and measures the average of structural similarities (SSIM) \cite{wang2004image} between task-related reference stimulus $x^{*}$ and $n$ reconstructed stimuli (with randomized initializations), i.e.~$\frac{1}{n}\sum_{i=1}^{n} \mathrm{SSIM}\left(x^{*} , \hat{x}^{*}_{i} \right)$ where \begin{equation} \label{eq:O2} \hat{x}^{*} = \underset{x}{\arg\max} \left( e^{-\left\|f\left(x\right)-r\right\|} \right) \end{equation} and $r = f\left(x^{*}\right)$, indicating how specific (or non-confounding) a representation $r$ is encoded.

%\underset{x}{\arg\max} f^*\left(x\right)
% ssim human vision similarity

\subsection*{Quasi-Second-Order Characterization: Invariance and Selectivity Path Search and Analysis}

\begin{figure}
\centering \includegraphics[width=0.5\textwidth]{Figs/fda.pdf} 
\caption{
{\bf Fitness-distance diagram.} Invariance (red) and selectivity (blue) curves: fitnesses of invariance and selectivity path search results plotted against distances from the optimal stimulus. Baseline (dashed gray) curve: graph of cosine function; invariance and selectivity curves of single inner-product neuron. Subsequent analyses, including path potential, subspace capacity and subspace alignment, can be performed. Invariance and selectivity path potentials can be defined on either the arccosine normalized diagram (shaded red and blue areas), or on the original diagram (line integrals).}
\label{fig:fd_diag}
\end{figure}

%\underset{x_{\delta}}{\arg\max} \hat{f}\left(x_{\delta}\right)
With respect to the optimal stimulus $\hat{x}$, the searches of invariance and selectivity paths, which consist of sets of invariant stimuli ($\left\lbrace x^{+}_{\delta} \right\rbrace$, i.e.~optimally excitatory) and selective stimuli ($\left\lbrace x^{-}_{\delta} \right\rbrace$, i.e.~optimally inhibitory) respectively, are formulated as
\begin{align}
x^{+}_{\delta} &= \underset{x_{\delta}}{\arg\max} f\left(x_{\delta}\right) \label{eq:I1} \\
&= \underset{x_{\delta}}{\arg\max} \left( e^{-\left\|f\left(x_{\delta}\right)-f\left(\hat{x}\right)\right\|} \right); \label{eq:I2} \\
x^{-}_{\delta} &= \underset{x_{\delta}}{\arg\min} f\left(x_{\delta}\right) \label{eq:S1} \\
&= \underset{x_{\delta}}{\arg\min} \left( e^{-\left\|f\left(x_{\delta}\right)-f\left(\hat{x}\right)\right\|} \right); \label{eq:S2}
\end{align}
where $0 < \delta \le \frac{\pi}{2}$, subject to $\left\| x_{\delta} \right\| = 1$ and $\langle x_{\delta} , \hat{x} \rangle = \cos\left(\delta\right)$. As visualized in Fig.~\ref{fig:methods}, the invariance path characterizes the $N$ dimensional ``curve'' that leads toward (or ideally maintains at) fitness as high as possible while moving away from the optimal stimulus, and the selectivity path characterizes the ``curve'' that leads toward fitness as low as possible while moving away. The search process is implemented as multiple runs of maximization/minimization on discretized $\delta \in \left\lbrace 0.1\pi, 0.2\pi, 0.3\pi, 0.4\pi, 0.5\pi\right\rbrace$, as the circular distance constraints shown in Fig.~\ref{fig:methods}, where each run is initialized with the result from previous run (and the $0.1\pi$ run directly with optimal stimulus $\hat{x}$) to increase the path continuity and searching speed. Each $\delta$ of both paths has a budget of $20N$ fitness evaluations which in this work appears to be practically sufficient. The maximization and minimization simply use the same back-end solver, and the linear constraint $\langle x_{\delta} , \hat{x} \rangle = \cos\left(\delta\right)$ can be easily handled through an extra conic projection before evaluation, i.e.~$f\left(p_c\left(p_s\left(x\right)\right)\right)$ where $p_c\left(x\right) = \cos\left(\delta\right)\hat{x} + \sin\left(\delta\right)\dot{x} \mathbin{/} \left\|\dot{x}\right\|$ and $\dot{x} = x - \langle\hat{x},x\rangle \hat{x}$. The way the simple linear constraint is constructed to enforce the exploration of a larger extent of the fitness landscape is the one of the main differences compared to \cite{erhan2010understanding}.
In this work, the distance constraint $\delta$ only goes up to $\frac{\pi}{2}$, since on the $N$ dimensional sphere, stimuli fall in $\frac{\pi}{2} < \delta \le \pi$ are simply ``negatives'' of those in $0 < \delta \le \frac{\pi}{2}$, which are of less uniqueness and interest; nevertheless, going up to the full range $0 < \delta \le \pi$ is numerically supported.

For analyzing the results of invariance/selectivity path searches, the following measures are adopted: (1) \emph{Path potential}: while the search results of invariance/selectivity paths can be visualized via the \emph{fitness-distance diagram} \cite{jones1995fitness} as shown in Fig.~\ref{fig:fd_diag} where perfect invariance and selectivity are flat lines at highest and lowest fitnesses respectively, the baselines of invariance and selectivity also can be intuitively (and analytically) defined as paths of the simplest form of neural networks---an inner-product neuron, $f(x) = w^{T}x$---which precisely overlap and follow the monotonic cosine falloff from its optimal stimulus $\hat{x}=w$, as $f(x_{\delta}) = \cos(x_{\delta})$ by definition; the invariance (and selectivity, similarly) path potential of a unit representation can thus be defined as $\int_{0}^{\frac{\pi}{2}}{\left| \cos^{-1}\left(f\left( x^{+}_{\delta} \right)\right) - \delta \right|}\mathrm{d}\delta \mathbin{/} {\frac{\pi}{2}}$, the area sandwiched between the invariance (and selectivity) and baseline curves (in $\cos^{-1}$ domain such that both potential values fall in the range $\left[0,1\right]$), measuring how invariant (and selective) a target neuron is compared to the baseline (i.e.~zero invariance/selectivity potential); for population representation where no baseline can be easily identified, the invariance (and selectivity) path potential is alternatively defined as $\int_{0}^{\frac{\pi}{2}}{\exp\left(-\left\|f\left(x^{+}_{\delta}\right)-f\left(\hat{x}\right)\right\|\right)}\mathrm{d}\delta \mathbin{/} {\frac{\pi}{2}}$. (2) \emph{Subspace capacity}: compared to path potential, which is designed to characterize the best invariance/selectivity path even when only one of such exists, subspace capacity estimates the ``dimensionality'' (i.e.~how diverse different paths can be) of the linear subspace formed by multiple path search results via the nuclear norm of concatenation of $n$ results $\left\|\left[x_{\delta,1},\dots,x_{\delta,n}\right]\right\|_{*}$ (in this work $n=20$ runs at $\delta=0.1\pi$). (3) \emph{Subspace alignment}, which measures the alignment between subspaces formed by task-related stimuli and invariance/selectivity paths (i.e.~how likely the invariance/selectivity can benefit, e.g., stimulus recognition) via estimating the sparsity of $n$ path search results projected onto the principal component vectors $V$ of task-related stimuli $\left\lbrace x^{t} \right\rbrace$, i.e.~$\frac{1}{n}\sum_{i=1}^{n} \left\| Vx_{\delta,i} \right\|_{1}$ ($n=20$ runs at $\delta=0.1\pi$ as well). %SUBSPACE OVERLAP?

### 2014-10-20 results.tex
First commit

%SI??? few line intro to HT algorithm [REFS]; methods not limited to HT only

The methods and measures proposed in this work are tested on the Spatio-Temporal Hierarchical Object Representation (STHOR) network \cite{pinto2009high, sthor}, an instance of various convolutional neural network (i.e.~deep learning algorithm) implementations \cite{fukushima1980neocognitron, lecun1998gradient, riesenhuber1999hierarchical, krizhevsky2012imagenet}, which also consists of the standard cascade of levels of ``convolution, nonlinear activation, pooling, and normalization'' layers (which together define a single ``level'' in this work), for its highly parameterizable design (i.e.~wide diversity of network structures can be tested) and decent performance on various visual recognition tasks \cite{pinto2009high, cox2011beyond, viglarge} even without any convolution kernel training (i.e.~fast experiments on large number of networks can be achieved, using kernels with random values). In this work, 100 shallow networks (i.e.~one-level, or L1, models) and 100 deep networks (i.e.~two-level, or L2, models) are randomly generated (all with 32 top-layer neurons) and tested to see how their representations vary with networks' depths and affect networks' performances (as shown in Fig.~\SFpef{}) on face pair matching tasks---accuracy of identifying pairs of different pictures from the same person and rejecting those from different persons---against the Labeled Faced in the Wild (LFW/LFW-a) dataset \cite{LFWTech, wolf2011effective}. Stimulus dimensionalities of shallow and deep neurons are $N=121$ and $N=441$ (i.e.~$11\times11$ and $21\times21$ receptive field sizes) respectively.

\subsection*{Shallow vs.~Deep Representations}

\begin{FPfigure} %\begin{figure}
\begin{minipage}{\textwidth} \centering \includegraphics[width=0.9\textwidth]{Figs/pic1.pdf} \end{minipage}
\caption{
{\bf Visualization of shallow and deep representations.} (A) Optimal stimulus search trajectories of shallow neurons. Each column of (A) demonstrates the optimal stimulus search trajectory of one neuron, by showing the initial ${1}/{f}$ random stimulus in (A.1), resultant optimal stimulus in (A.5), and 3 intermediate stimuli corresponding to 3 largest second derivatives (i.e.~curvatures) of the fitness history in (A.2--4). (B) Invariance paths of shallow neurons. Each column of (B) demonstrates the invariance path search results of one neuron, starting from corresponding optimal stimulus as shown in (A.5) and moving away with distance constraints from $\delta = 0.1\pi$ to $0.5\pi$ as shown in (B.1--5) accordingly. (C) Selectivity paths of shallow neurons. Definitions follow (B). (D) Invariance subspaces of shallow neurons. Each column of (D) randomly shows 5 results out of 20 runs of invariance path searches at $\delta = 0.1\pi$. (E--H) Optimal stimulus search trajectories, invariance paths, selectivity paths, and invariance subspaces, of deep neurons, respectively. Definitions follow (A--D). Color indicates fitness (i.e.~response) of a neuron (definition of color map follows Fig.~\ref{fig:methods}).}
\label{fig:ind_res}
\end{FPfigure} %\end{figure}

Figure \ref{fig:ind_res}A and \ref{fig:ind_res}E exemplify the trajectories of optimal stimulus searches of 32 top-layer neurons from the best performing shallow network and 32 from the best performing deep network respectively. Worth to note, to further increase the searching speed, the initial stimulus is selected from 1000 ${1}/{f^{\alpha}}$ random stimuli with the best fitness, where $\alpha \in \left\lbrace -4,-3,-2,-1,0 \right\rbrace$, without sacrificing the nature of random initialization. From the results, two major differences can be observed. First, the fitness landscapes of deep neurons are significantly more complex than those of shallow neurons, as a large portion of searching trajectories for deep neurons actually go down then up before reaching the optima, which also suggests this numerical framework can handle non-convexity reasonably well. Second, the optimal stimuli of deep neurons are also relatively more complex (i.e.~consisting of more textural components) than those of shallow neurons (which are more uni-textured or Gabor-like).

%is also directly confirmed in this work. %to our knowledge, for the first time.

\begin{figure}
\centering \includegraphics[width=0.5\textwidth]{Figs/e_fig3a_2-nup-crop.pdf}
\caption{
{\bf Comparison of shallow and deep fitness landscapes.} (A) Fitness-distance diagram of shallow neurons. Red, blue and green dots indicate results of invariance and selectivity path searches, and random walks. Means of results are plotted as solid lines in corresponding colors, and the cosine baseline curve as dashed line in gray. (B) Fitness-distance diagram of deep neurons. Definitions follow (A).}
\label{fig:ind_fdd} % NUMERICAL PRECISION!! WHY NO SLSC?
\end{figure}

Figure \ref{fig:ind_res}B/\ref{fig:ind_res}C and \ref{fig:ind_res}F/\ref{fig:ind_res}G exemplify the invariance/selectivity paths on the same sets of shallow and deep neurons as in Fig.~\ref{fig:ind_res}A and \ref{fig:ind_res}D. It can be observed that, while invariance paths are mostly phase changes and selectivity paths are leading toward meaningless noises (all at the same falloff rate) for shallow neurons, both types of paths consist of sophisticated shape deformations for deep neurons. Also, although most shallow neurons are selective to manually rotated optimal stimuli (i.e.~Gabor filters of different orientations) as well, their fitnesses still do not drop faster then the nonparametric numerical solutions as presented in Fig.~\ref{fig:ind_res}C. In fact, this intriguing difference generalizes across all shallow vs.~deep neurons tested in this work. As the fitness-distance diagrams shown in Fig.~\ref{fig:ind_fdd}, on average, though shallow neurons (Fig.~\ref{fig:ind_fdd}A) have good invariance (i.e.~red curve stays high), they do not have any selectivity (i.e.~blue curve drops slow and completely overlaps with baseline), while deep neurons (Fig.~\ref{fig:ind_fdd}B) show both good invariance and selectivity (i.e.~red curve stays high; blue curve drops fast). Random walks on the fitness landscape \cite{jones1995fitness}, shown as the green curve in Fig.~\ref{fig:ind_fdd}, can be defined as $f\left(\cos\left(\delta\right)\hat{x} + \sin\left(\delta\right)\tilde{x}\right)$ where $\tilde{x}$ is any random stimulus such that $\left\| \tilde{x} \right\| = 1$ and $\left\langle \hat{x},\tilde{x} \right\rangle = 0$, which can be obtained through random projection of $\mathrm{Null}\left(\hat{x}\right)$. Comparing the gaps between selectivity and random-walk curves in Fig.~\ref{fig:ind_fdd}A and \ref{fig:ind_fdd}B again shows the selectivity of deep neurons are in fact way more ``selective'' than of shallow neurons. To further characterize the invariance properties, multiple runs of invariance path searches are executed and results are visualized in Fig.~\ref{fig:ind_res}D and \ref{fig:ind_res}H. The diversity/dimensionality (i.e.~subspace capacity) of the ``invariance plateau'' (e.g.~the central high fitness region in Fig.~\ref{fig:methods}) of deep neurons is also more apparent than that of shallow neurons. Worth to note, it can be seen in Fig.~\ref{fig:ind_fdd} that small fractions of invariance path search results in fact have higher fitness than the optimal stimuli due to nature of non-convex problems, which however does not cause noticeable differences on resultant statistics.

\begin{figure}
\centering \includegraphics[width=0.5\textwidth]{Figs/e_fig23b-crop.pdf}
\caption{
{\bf Comparisons of shallow and deep representation measures.} Representation measures from left to right are, optimal stimulus spectral complexity (OSSC), invariance path potential (INPP), selectivity path potential (SLPP), and invariance subspace capacity (INSC), where data points from shallow networks (left distributions) and deep networks (right distributions) are shown side by side. Significances of differences of means are obtained through permutation tests, and $p < 0.001$ in all four measures. Sensitivity indexes ($d'$) between distributions of shallow and deep representation measures from left to right are 1.63, 0.42, 4.73, and 3.67.} %Means and standard deviations from left to right are, $0.53\pm0.12$, $0.74\pm0.13$, $0.67\pm0.22$, $0.76\pm0.17$, $0.00\pm0.00$, $0.47\pm0.14$, $0.30\pm0.03$, and $0.42\pm0.04$.
\label{fig:ind_mea}
\end{figure}

Comparisons of spectral complexities, invariance/selectivity path potentials, and invariance subspace capacities of shallow and deep neurons (i.e.~unit representations) are summarized in Fig.~\ref{fig:ind_mea}, where all measures are normalized to have analytical upper and lower bounds being 1 and 0. These measures break down the differences between shallow and deep representations, in addition to the straightforward representation separability/classifiability as used in previous studies \cite{donahue2014decaf, zeiler2014visualizing}. For instance, subtle differences of visual features have better chances being distinguished when the ``gap'' between invariance and selectivity curves (i.e.~the dynamic range of neuronal responses, or amplification ratio of differences) is large. The increasing of complexity in deeper neurons' preferred/optimal stimuli also agrees with neurobiological findings in higher visual cortical areas (e.g.~V2 and V4 \cite{hegde2000selectivity, pasupathy2001shape}). In addition, the highly structural representations of neurons from deep networks with purely random convolution kernels found in the experiments explain why such type of random deep networks can still have good performances, on top of the theoretical analysis of shallow networks \cite{saxe2011random}.

%, which is also observed in this work (as shown in Fig.~\SFpef{})
%and futher point out the architecture is special/good
%[lecun's random]
%resize of OS

\subsection*{Good vs.~Bad Representations} %\label{sec:good_bad}

\begin{figure}
\centering \includegraphics[width=0.9\textwidth]{Figs/pic2.pdf}
\caption{
{\bf Explanation power of optimal stimulus.} (A) Deep network with the best average explanation power of optimal stimulus. Each column of row (A.1) shows the optimal stimulus of the corresponding neuron. Each column of row (A.3--5) shows the task-related stimuli with the first, second, and third highest explainabilities (i.e.~inner-product distances to the optimal stimulus) respectively. Each column of row (A.2) shows the average of task-related stimuli with top 1\% explainabilities. (B--D) Deep networks with the second best, third best, and worst average explanation power of optimal stimulus, respectively. Definitions of rows follow (A). Color indicates explainability, and higher means better (definition of color map follows Fig.~\ref{fig:methods}).}
\label{fig:ind_exp}
\end{figure}

% how're stimuli sampled (face regions) from LFW-a
% DIFF BETWEEN FORMULA AND HERE??
%(calculated using only the top 1\% nearest task-related stimuli for visualization reasons; full set of task-related stimuli yields similar $R^2$ as well)
%Point out interesting cells. %reciprocal
% only deep nets, what runs are combined (by paragraph)
%; using all stimuli gives similar result of measure vs.~performance correlation.
%(though not improving $R^2$ when added into multiple correlation test ??)

To test how well optimal stimulus may linearly explain task-related stimuli and how the explanation power correlates with network's performance, 10,000 image patches are randomly sampled out of the predefined facial regions of pictures from the LFW-a dataset \cite{wolf2011effective} and compared against the optimal stimuli of all 32 top-layer neurons from all 100 deep networks. Figure \ref{fig:ind_exp} shows the results of best 3 (in \ref{fig:ind_exp}A--C) and worst 1 (in \ref{fig:ind_exp}D) networks in terms of explanation power. For simplicity and clarity of presentation, the explanation power measure is calculated using the top 1\% (i.e.~100) of the task-related stimuli with highest explainabilities, without biasing the results compared to using all task-related stimuli. It can be observed that, though none of the task-related stimuli (or arguably, of all natural stimuli) is highly similar to any optimal stimulus, there is an $R^2=0.36$ correlation (with details shown in Fig.~\SFexp{}) between network's explanation power and performance, and when alternatively comparing the average of the top 1\% of task-related stimuli and the optimal stimuli, the similarities become apparent. %representativeness

\begin{figure}
\centering \includegraphics[width=0.5\textwidth]{Figs/pic3.pdf}
\caption{
{\bf Subspace alignment between invariance and selectivity paths and task-related stimuli.} (A) Reference stimuli used as starting points of invariance and selectivity path searches. (B) Best invariance subspace alignments. Each column of row (B.1) shows an invariance path search result with the best alignments against the principle component vector space of task-related stimuli (i.e.~the eigenface vector space). Each column of row (B.2) shows the difference between result in (B.1) and corresponding reference stimulus in (A) for better visual comparison. All columns do not necessary come from the same deep network. (C--E) Best selectivity subspace alignments, worst invariance subspace alignments, and worst selectivity subspace alignments, respectively. Definitions of rows follow (B). Color indicates alignment measure, and lower means better (definition of color map follows Fig.~\ref{fig:methods}).} % Show PCs or not? ref stimulus is not perfect either?
\label{fig:pop_aln}
\end{figure}

It has been shown that invariance and selectivity are extremely important for recognition in various modalities and species \cite{desimone1991face, ito1995size, quiroga2005invariant}; however, to how these properties directly affect the recognition performances remains unclear. To address this puzzle, starting from 16 randomly sampled reference stimuli (as shown in Fig.~\ref{fig:pop_aln}A), invariance and selectivity path searches are performed at $\delta = 0.1\pi$ using Eq.~(\ref{eq:I2}, \ref{eq:S2}) for population representations. Figure \ref{fig:pop_aln}B--E demonstrates the results of invariance and selectivity subspace alignment analyses. It can be observed that, good alignments (i.e.~sparser representations in principle component vector spaces) correspond to highly structural deformations or lighting changes, which are known to be important factors in visual recognition, while bad alignments correspond to mostly meaningless noises. The correlation between network's invariance/selectivity subspace alignment measure and performance is $R^2=0.19/0.31$ (with details shown in Fig.~\SFaln{}). Network's invariance subspace capacity (in terms of population representation), which though has an $R^2=0.15$ correlation as well, is in fact negatively correlated to network's performance (with details shown in Fig.~\SFinc{}), mainly due to the fact that poor performing network's invariance path search results are usually noisier (as depicted in Fig.~\ref{fig:pop_aln}D), and thus should be interpreted differently from unit representation cases. Invariance/selectivity path potentials for population representations, on the other hand, are weaker in explaining networks' performances (with $R^2=0.10/0.06$), due to the fact that all deep networks have high but less ordered invariance and selectivity (as the fitness-distance diagram shown in Fig.~\SFfda{}).

\begin{figure}
\centering \includegraphics[width=0.5\textwidth]{Figs/pic4.pdf}
\caption{
{\bf Encoding specificity of task-related stimulus.} (A) Reference stimuli used as sources of representations. (B) Best encoding specificities. Each column of row (B.1--3) shows reconstructed stimuli of corresponding reference stimulus in (A) with top 3 SSIM measures (out of 10 reconstructions), from deep network of the best average encoding specificity. All columns do not necessary come from the same deep network. (C--E) Second best, third best, and worst encoding specificities, respectively. Definitions of rows follow (B). Color indicates SSIM measure, and higher means better (definition of color map follows Fig.~\ref{fig:methods}).}
\label{fig:pop_ens}
\end{figure}

%(see Text S\ref{?})
To further understand the meaning of population representation $r$ formed by a deep network $f$, the inverse function $f^{-1}\left( r \right)$ is approximated through the numerical optimization framework to reveal what kind of stimulus can drive the deep network to give output $r$. Instead of inverting randomly generated $r$, which may not have feasible or interpretable solutions, we use $r$ from known reference stimulus $x^{*}$, and study the encoding specificity and its relationship to performance. Figure \ref{fig:pop_ens} shows the reference stimuli (in \ref{fig:pop_ens}A), top 3 (in \ref{fig:pop_ens}B--D) and worst 1 (in \ref{fig:pop_ens}A) examples of encoding specificity, and the overall correlation between encoding specificity and performance is $R^2=0.41$ (with details shown in Fig.~\SFenc{}; SSIM measured using default settings of the standard implementation \cite{wang2004image} and different local window sizes yield similar correlations; other similarity measures like SNR, PSNR, and inner-product distance yield similar correlations as well). It can be observed that, though the reconstructed stimuli are overall of lower spatial frequencies, SSIM is still able to capture the relative similarity (e.g.~reconstructions of task-related stimulus 1 and 5).

\begin{figure}
\centering \includegraphics[width=0.5\textwidth]{Figs/e_fig7-crop.pdf}
\caption{
{\bf Spearman's correlations between deep network's representation measures and performance.} Representation measures from left to right are, optimal stimulus explanation power (OSEP), invariance path potential (INPP), selectivity path potential (SLPP), invariance subspace capacity (INSC), invariance vs.~task-related stimuli subspace alignment (ITSA), selectivity vs.~task-related stimuli subspace alignment (STSA), task-related stimulus encoding specificity (TSES), and all measures together (ALL) using linear multiple correlation analysis. Pearson's correlations following the same order are 0.37, 0.02, 0.02, 0.19, 0.17, 0.23, 0.39, and 0.69.}
\label{fig:all_R^2}
\end{figure}

%single best measure is...
%second best is ... and link to ...

Spearman's correlations between all representation measures and deep networks' performances are summarized in Fig.~\ref{fig:all_R^2}, and 71\% of the variance of deep networks' performances can be explained by the proposed measures altogether. These measures further break down the differences between deep representations and suggest features like network's explanation power, invariance and selectivity subspace properties, and encoding specificity against the task-related stimuli can be extremely important for the effectiveness of representations. Encoding specificity, as a single representation measure, can best explain the deep network's performance. Though bearing strong similarity to the invariance path search in terms of mathematical formulation, encoding specificity is in fact a radically different and complementary measure, since its numerical optimization is not constrained through distances to the reference stimulus, while distance constraints can induce \emph{a priori} similarities; it can better function as an unconstrained global characterization of the ``encoding landscape'' (i.e.~multivariate tuning landscape of a population) and estimate if the (reconstructed) stimuli that a deep network is invariant to are ``selective'' (i.e.~visually similar to original stimuli) at the same time. Explanation power of optimal stimulus is the second best single representation measure. Following the efficient coding theory \cite{barlow1961possible, laughlin1981simple}, one may exclude the possibility of an optimal stimulus looking highly similar to the task-related stimuli. Nevertheless, in a high-dimensional stimulus space, how to optimally construct the tuning landscape (and place the optimal stimuli) remains unclear, and whether the observed phenomenon (as shown in Fig.~\ref{fig:ind_exp}), in which optimal stimuli of well performing networks in fact are highly similar to the averages of neighboring task-related stimuli, is essential to the optimal tuning landscape still needs to be further studied. Worth to note, similar phenomenon was also reported in previous studies on well performing networks \cite{le2012building, simonyan2013deep}. Such ``linearities'' around the task-related stimuli out of highly nonlinear networks may also explain why natural stimuli can be more efficient in estimating the tuning of neurons \cite{talebi2012natural}.

%1st-order predictive power vs 2nd-order (0.56, 0.58)
%For second-order representation measures provide relative lower

### 2014-12-17 abstract.tex
First edits

% Please keep the abstract between 250 and 300 words

A central challenge in sensory neuroscience is to describe how the activity in populations of neurons represents properties of the external environment in a format suitable for guiding behavior.
However, while neurophysiologists have long been able to record the responses of neurons in awake, behaving animals, it is another matter entirely to say what a given neuron ``does.''
A key problem is that in many sensory systems, the space of all possible stimuli that one might encounter is effectively infinite.  For instance, in vision, scenes are combinatorially complex---an individual human will encounter only a small fraction of the possible scenes in the world; indeed only a small fraction of the possible images have \emph{ever} been experienced in the collective history of the human race.
Thus, investigations of neuronal function---real and simulated---are almost always critically limited by the number of stimuli that can be considered. 
In some early stages of sensory processing, a combination of intuition and systems identification theory has led to reasonable models of early neuron response functions (e.g. Gabor wavelet models in cortical area V1), but cascading nonlinearities have tended to foil our best efforts in later stages of sensory processing.

Here, we propose a closed-loop experimental framework for characterizing the response properties of sensory neurons, building on past efforts in closed-loop experimental methods, and leveraging recent advances in artificial neural network models to serve as as a proving ground for our techniques.
To the extent that modern deep learning models increasingly serve as \emph{de facto} working hypotheses for biological vision \ref{jim, others}, we argue that a unified approach of developing methods to interrogate

### 2014-12-17 introduction.tex
First edits

@@
+A central drive in the study of sensory systems is the production of models that
+
 The mammalian ventral visual pathway consists of a hierarchical cascade of visual areas that progressively reformat visual information from pixel-like retinotopic representations into formats suitable for high-level tasks \cite{dicarlo2007untangling, dicarlo2012does}, such as recognizing visual objects.
@@
 Szegedy et al.~\cite{szegedy2013intriguing} also used similar approach to study the selectivity of top-layer neurons inside multi-layer convolutional networks.
+
 In experimental neuroscience studies, methods can be categorized by the stimulus used to characterize the biological---naturally nonparametric---models \cite{wu2006complete}: parametric or nonparametric.
@@
 
-In this work, we propose a \emph{unified framework} based on modern numerical optimization techniques to help us uncover the sensory representation encoded by neurons deep inside neural networks, where the most flexible and generalizable settings---nonparametric network model with nonparametric stimulus---are supported to maximize the applicability of this framework, with the inherent difficulties and inefficiency resolved by carefully constructing and constraining the numerical framework.
+In this work, we propose a unified framework based on modern numerical optimization techniques to help us uncover the sensory representation encoded by neurons deep inside neural networks, where the most flexible and generalizable settings---nonparametric network model with nonparametric stimulus---are supported to maximize the applicability of this framework, with the inherent difficulties and inefficiency resolved by carefully constructing and constraining the numerical framework.
 Characterization of the ``tuning landscape'' of neurons to the high-dimensional nonparametric stimulus includes both first- and quasi-second-order ``landscape features'', i.e.~optimal stimulus and its invariance and selectivity directions---the most significant structural features of the surrounding landscape---where the inefficiency of modeling the entire Hessian can be avoided.

### 2014-12-17 methods.tex
First edits

@@
 \caption{ 
-{\bf Schematic of numerical optimization framework.} Deep network $f$: hierarchical cascade of layers (shaded tall rectangles) of artificial neurons (circles) transforming the stimulus $x$ (bottom layer, leftmost) into high-level representations $r$ (top layer, rightmost) suitable for linear classification.
-Optimization framework: closed-loop algorithm that repeatedly measures the response(s) of one or multiple neuron(s), calculate the fitness, and adaptively generate stimuli to characterize the target neuron(s) via optimizing the fitness functions defined for different purposes (see Eq.~(\ref{eq:O1}--\ref{eq:S2})); when characterizing multiple neurons for population representation of given optimal or reference stimulus, response of an imaginary neuron (dashed circle) tuned to the given stimulus is optimized (see Eq.~(\ref{eq:O2}, \ref{eq:I2}, \ref{eq:S2})) equivalently.}
+{\bf Schematic of numerical optimization framework.} (A) Here we consider a deep network, which consists of a hierarchical cascade of layers (shaded tall rectangles) of artificial neurons (circles) transform a pixel stimulus $x$ (bottom layer, leftmost) into a high-level feature representation $r$ (top layer, rightmost) suitable for classification.
+To study the properties of units within these networks, we employ an optimization framework shown in (B).
+The framework consists of a closed-loop algorithm that iteratively measures the response(s) of one or multiple neuron(s), computes the value of a predefined ``fitness'' function, and adaptively generates new stimuli to characterize the target neuron(s).
+Different fitness functions allow us to probe different properties of the units (see Eq.~(\ref{eq:O1}--\ref{eq:S2})).
+%optimizing the fitness functions defined for different purposes (see Eq.~(\ref{eq:O1}--\ref{eq:S2})); when characterizing multiple neurons for population representation of given optimal or reference stimulus, response of an imaginary neuron (dashed circle) tuned to the given stimulus is optimized (see Eq.~(\ref{eq:O2}, \ref{eq:I2}, \ref{eq:S2})) equivalently.}
 \label{fig:concept}
@@
 Representation $r$, in this work, is equivalent to the response(s) of artificial neuron(s), and is denoted as unit representation (or scalar representation, i.e.~$r \in \mathbb{R}$) when referring to single neurons and population representation (or vector representation, i.e.~$r \in \mathbb{R}^R$) when referring to groups of neurons (with group size $R$).
-Arguably, for studying both artificial and biological neural networks, especially when evaluating task-related performances, the ability of characterizing population representations is usually more important than of characterizing unit representations, considering the differences in representational powers.
+Arguably, for studying both artificial and biological neural networks, the ability to characterize population representations is arguably at leaast as importnat, if not more important than of characterizing unit representations \cite{churchland}.
 This framework supports characterizing both unit (using Eq.~(\ref{eq:O1}, \ref{eq:I1}, \ref{eq:S1}), though Eq.(\ref{eq:O2}, \ref{eq:I2}, \ref{eq:S2}) can be used as well) and population (using Eq.(\ref{eq:O2}, \ref{eq:I2}, \ref{eq:S2})) representations in a unified way.
 
-As aforementioned, this framework also treats network as non-analytical black boxes and poses no assumption about networks' properties.
-Nevertheless, all stimuli used in this work follow the constant energy constraint $\left\| x \right\| = E$, as suggested by previous studies where simply enhancing the contrast of a fixed stimulus can as well cause increased firing rates \cite{albrecht1982striate, cheng1994comparison}, which in most cases leads to meaningless outcomes since the spatial ``patterns'' of stimulus are not changed.
-For simplicity of mathematical formulation, the setting of $E=1$ is used for the rest of the paper, while in experiments $E$ is set to the average of task-related stimulus energy before normalization.
-In both methods, numerical optimization are executed multiple times from randomized initial settings, for two main reasons: (1) increasing robustness of the numerical solutions of, e.g., optimal stimulus and invariance/selectivity path searches (2 runs executed), and (2) providing statistical samplings of the solution spaces of, e.g., encoding specificity and invariance/selectivity subspace searches (10 and 20 runs executed).
+Our framework treats networks as non-analytical black boxes: it makes no assumption about network structure or properties (e.g.
+we do not assume an analytical gradient is available, even though such gradients are available for many kinds of artificial networks).
+We made this strategic decision in order to allow our methods to transfer over to experiments with real neurons, where only the output of the neuron is available.
+While no specific knowledge of the network itself was assumed, we did restrict the space of stimuli considered, based on prior knowledge of real and artificial neuronal response properties.
+In particular, all stimuli that we consider in this work follow a constant energy constraint $\left\| x \right\| = E$, following the general observation that real and artificial neurons are often modulated by stimulus contrast \cite{albrecht1982striate, cheng1994comparison}, but that this modulation is less interesting when considering the form/pattern selectivity of the unit.
+Limiting the stimulus search space in this manner vastly reduces the range of possible stimuli to consider, and it avoids degenerate solutions that simply maximize stimulus contrast.
+% DDC: what does this mean? Make clearer
+For simplicity of mathematical formulation, the setting of $E=1$ is used for the rest of the paper, while in experiments, $E$ is set to the average of task-related stimulus energy before normalization.
+
+In all experiments, numerical optimization was executed multiple times, starting from different random initial stimuli.
+Random  for two main reasons: (1) increasing robustness of the numerical solutions of, e.g., optimal stimulus and invariance/selectivity path searches (2 runs executed), and (2) providing statistical samplings of the solution spaces of, e.g., encoding specificity and invariance/selectivity subspace searches (10 and 20 runs executed).
 The phase of optimal stimulus search may be omitted when studying the invariance, selectivity and encoding specificities of representations of given reference stimuli.
@@
 
-In this work, optimal stimulus is numerically derived through the iterative optimization as \begin{equation} \label{eq:O1} \hat{x} = \underset{x} {\arg\max} f\left(x\right) = \underset{x_{g} \in \Omega \left( x_{g-1} , m_{g-1} \right)}{\arg\max} f\left(x_{g}\right),\end{equation} where $m_{g}=U\left(\left\lbrace f\left(x_{g}\right) \right\rbrace , m_{g-1}\right)$ and $1 \le g \le G$, subject to $\left\| x \right\| = 1$.
-The algorithm starts from random initial point $x_{0}$ on an $N$ dimensional sphere, as the noisy pattern shown in the bottom of Fig.~\ref{fig:methods}, and samples a set of $\lambda$ neighboring points (through function $\Omega$) as candidates to be evaluated, from an initial model $m_{0}$ with null distribution (in this work, multivariate Gaussian distribution starting with covariance matrix $\Sigma = I$).
-Through measuring the fitnesses of sampled points $\left\lbrace f\left(x_{g}\right) \right\rbrace$, the algorithm updates its model $m_{g}$ (i.e.~mean and covariance of the distribution, through function $U$) for generating samples in the following iterations, where the same operations repeat.
-Usually the maximally allowed iteration number $G$ is set to prevent the algorithm from running unreasonably long; nevertheless, early termination is also practically possible when optimality is reached and the neighborhood size shrinks below threshold, as shown in the center of Fig.~\ref{fig:methods}.
-Though the fitness landscape can be highly nonlinear, such sequential optimization procedure is usually capable of gradually accumulating knowledge of the landscape and adjusting its search directions to climb onto higher fitness areas, and the smoother the landscape is, the faster the algorithm converges.
+In this work, a unit's ``optimal stimulus'' was numerically derived through the iterative optimization as \begin{equation} \label{eq:O1} \hat{x} = \underset{x} {\arg\max} f\left(x\right) = \underset{x_{g} \in \Omega \left( x_{g-1} , m_{g-1} \right)}{\arg\max} f\left(x_{g}\right),\end{equation} where $m_{g}=U\left(\left\lbrace f\left(x_{g}\right) \right\rbrace , m_{g-1}\right)$ and $1 \le g \le G$, subject to $\left\| x \right\| = 1$.
+We note that there is no general guarantee that an optimum found by this procedure will be a global optimum; however, we argue that even exploring the response landscape around a high local maximum is useful in characterizing a unit.
+The algorithm started from random initial point $x_{0}$ on an $N$ dimensional sphere (e.g.
+the noise pattern shown in the bottom of Fig.~\ref{fig:methods}) and samples from set of $\lambda$ neighboring points (through evaluation of the function $\Omega$) were taken as candidates to be evaluated, from an initial model $m_{0}$ with null distribution (in this work, multivariate Gaussian distribution starting with covariance matrix $\Sigma = I$).
+By measuring the response magnitude (``fitness,'' in the nomenclature of derivative-free optimization) of sampled points $\left\lbrace f\left(x_{g}\right) \right\rbrace$, the algorithm updates its model $m_{g}$ (i.e.~mean and covariance of the distribution, through function $U$) for generating samples in the following iterations, where the same operations repeat.
+% why is this parameter given a name?  Is it useful to know that?
+We set a threshold on the maximum allowed number of iterations, $G$, to prevent the algorithm from running for an unreasonably long time; however, more frequently, early termination occurred when the sample neighborhood size shrank below threshold, as shown in the center of Fig.~\ref{fig:methods}.
+Though the fitness landscape can be highly nonlinear, such sequential optimization procedure is often capable of gradually accumulating knowledge of the landscape and adjusting its search directions to climb onto higher fitness areas, and the smoother the landscape is, the faster the algorithm converges.
 As exemplified in Fig.~\ref{fig:methods}, the ``most salient pattern'', or direction leading to high fitness areas usually can be rapidly extracted (see Fig.~\ref{fig:ind_res} for more examples).
 In most scenarios, fixing the maximally allowed number of fitness evaluations $\lambda G$ to, e.g., $100N$ (as adopted for the optimal stimulus search in this work) can lead to reasonably good convergence speed vs.~quality trade-off.
-The aforementioned sequential optimization concept is in fact commonly adopted in modern stochastic optimization \cite{spall2005introduction}, and the CMA-ES (Covariance Matrix Adaptation Evolution Strategy) algorithm \cite{hansen2001completely} is chosen as the back-end solver of this numerical framework, for its decent convergence speed and capability of handling rugged landscapes.
+The aforementioned sequential optimization concept is in fact commonly adopted in modern stochastic optimization \cite{spall2005introduction}, and the CMA-ES (Covariance Matrix Adaptation Evolution Strategy) algorithm \cite{hansen2001completely} is chosen as the back-end solver of this numerical framework, for its good convergence speed and capability of handling ``rugged'' response landscapes.
 Readers can refer to \cite{hansen2001completely} for detailed definitions of model $m$ and functions $\Omega$ and $U$.
-The energy constraint of stimulus is simply handled through the spherical projection before evaluation, i.e.~$f\left(p_s\left(x\right)\right)$ where $p_{s}\left(x\right) = {x} / {\left\|x\right\|}$, while the solver works unconstrainedly.
+The energy constraint on the stimulus search space was implemented by simply navigating the stimulus space in spherical coordinates, with a fixed radius.
+%through a spherical projection before evaluation, i.e.~$f\left(p_s\left(x\right)\right)$ where $p_{s}\left(x\right) = {x} / {\left\|x\right\|}$, while the solver worked in an unconstrained manner.
 
-For analyzing the results of optimal stimulus searches, the following measures are adopted: (1) \emph{Spectral complexity}, which is estimated through the $L^{1}$ norm of the (2 dimensional) Fourier power spectrum of optimal stimulus, i.e.~$\left\| \mathcal{F}(\hat{x}) \right\|_{1}$, where higher value suggests higher non-sparsity (i.e.~more spectral components required to represent the signal).
+In order to analyze the results of optimal stimulus searches, we adopted the following measures: (1) \emph{Spectral complexity}, which is estimated through the $L^{1}$ norm of the (2 dimensional) Fourier power spectrum of optimal stimulus, i.e.~$\left\| \mathcal{F}(\hat{x}) \right\|_{1}$, where higher value suggests higher non-sparsity (i.e.~more spectral components required to represent the signal).
 (2) \emph{Explanation power}, which is estimated as the mean of the linear explainability of task-related stimuli---rectified inner-product distances between a set of $n$ task-related stimuli $\left\lbrace x^{t} \right\rbrace$ and the optimal stimulus itself, i.e.~$\frac{1}{n}\sum_{i=1}^{n}\max\left(\langle x^{t}_{i} , \hat{x} \rangle , 0\right)$---and measures how well the optimal stimulus can linearly ``approximate'' task-related stimuli (or intuitively how much neuronal response might be elicited).
@@
 %\underset{x_{\delta}}{\arg\max} \hat{f}\left(x_{\delta}\right)
+
+While the optimal stimulus provides important information about the nature of units in a network, important additional information can be extracted about the response landscape around the peak response.
+The begin to quantify properties of this landscape, we define the ``selectivity'' and ``invariance'' paths.
+
 With respect to the optimal stimulus $\hat{x}$, the searches of invariance and selectivity paths, which consist of sets of invariant stimuli ($\left\lbrace x^{+}_{\delta} \right\rbrace$, i.e.~optimally excitatory) and selective stimuli ($\left\lbrace x^{-}_{\delta} \right\rbrace$, i.e.~optimally inhibitory) respectively, are formulated as
@@
 As visualized in Fig.~\ref{fig:methods}, the invariance path characterizes the $N$ dimensional ``curve'' that leads toward (or ideally maintains at) fitness as high as possible while moving away from the optimal stimulus, and the selectivity path characterizes the ``curve'' that leads toward fitness as low as possible while moving away.
-The search process is implemented as multiple runs of maximization/minimization on discretized $\delta \in \left\lbrace 0.1\pi, 0.2\pi, 0.3\pi, 0.4\pi, 0.5\pi\right\rbrace$, as the circular distance constraints shown in Fig.~\ref{fig:methods}, where each run is initialized with the result from previous run (and the $0.1\pi$ run directly with optimal stimulus $\hat{x}$) to increase the path continuity and searching speed.
+The search process was implemented as multiple runs of maximization/minimization on discretized $\delta \in \left\lbrace 0.1\pi, 0.2\pi, 0.3\pi, 0.4\pi, 0.5\pi\right\rbrace$, as the circular distance constraints shown in Fig.~\ref{fig:methods}, where each run is initialized with the result from previous run (and the $0.1\pi$ run directly with optimal stimulus $\hat{x}$) to increase the path continuity and searching speed.
 Each $\delta$ of both paths has a budget of $20N$ fitness evaluations which in this work appears to be practically sufficient.
@@
 %SUBSPACE OVERLAP?
+
+\subsection{Deep Convolutional Networks}
+
+The methods and measures proposed above were tested using random convolutional neural networks (``randnets,'' \cite{pinto2009high, sthor}), a simplified variety of a deep convolutional neural network \cite{fukushima1980neocognitron, lecun1998gradient, riesenhuber1999hierarchical, krizhevsky2012imagenet}, where convolution kernel weights are randomly assigned rather than trained.
+While these models cannot match the performance of deep networks trained on very large quantities of data (e.g.
+using backpropagation, \ref{krizhevsky2012imagenet}), they are suprisingly competitive for a variety of tasks, especially in the regime where massive quantities of training data are not available \cite{pinto2009high, cox2011beyond, viglarge}.
+Importantly, they do not require extensive training---a classifier is trained on the output of the network, but the weights of the network itself are not trained---meaning that large numbers of candidate networks can be rapidly generated with different hyperparameters (e.g.
+different kernel sizes, normalization exponents, etc.) and compared.
+Thus, random networks are a useful tool when exploring a diversity of network architectures in settings where absolute task performance is not essentail.
+We have previously shown that searching within the space of hyperparamters can yield suprisingly effective networks, even when the networks are not trained in the conventional sense \cite{james}.
+
+The random networks used here were implemented using the open source STHOR library \cite{sthor}.
+Reference code for the experiments described in this paper is available at [URL TBD].
+As with other convolutional networks, our networks consist of a standard cascade of convolution, nonlinear activation, pooling, and normalization layers (which together define a single ``level'' in this work).
+In this work, we focused 100 shallow networks (i.e.~one-level, ``L1,'' models) and 100 deep networks (i.e.~two-level, ``L2,'' models) were randomly generated (all with 32 top-layer neurons) and tested to see how their representations vary with networks' depths and affect networks' performances (as shown in Fig.~\SFpef{}) on face pair matching tasks---accuracy of identifying pairs of different pictures from the same person and rejecting those from different persons---against the Labeled Faced in the Wild (LFW/LFW-a) dataset \cite{LFWTech, wolf2011effective}.
+Stimulus dimensionalities of shallow and deep neurons are $N=121$ and $N=441$ (i.e.~$11\times11$ and $21\times21$ receptive field sizes) respectively.

### 2014-12-17 results.tex
First edits

@@
 %SI??? few line intro to HT algorithm [REFS]; methods not limited to HT only
-
-The methods and measures proposed in this work are tested on the Spatio-Temporal Hierarchical Object Representation (STHOR) network \cite{pinto2009high, sthor}, an instance of various convolutional neural network (i.e.~deep learning algorithm) implementations \cite{fukushima1980neocognitron, lecun1998gradient, riesenhuber1999hierarchical, krizhevsky2012imagenet}, which also consists of the standard cascade of levels of ``convolution, nonlinear activation, pooling, and normalization'' layers (which together define a single ``level'' in this work), for its highly parameterizable design (i.e.~wide diversity of network structures can be tested) and decent performance on various visual recognition tasks \cite{pinto2009high, cox2011beyond, viglarge} even without any convolution kernel training (i.e.~fast experiments on large number of networks can be achieved, using kernels with random values).
-In this work, 100 shallow networks (i.e.~one-level, or L1, models) and 100 deep networks (i.e.~two-level, or L2, models) are randomly generated (all with 32 top-layer neurons) and tested to see how their representations vary with networks' depths and affect networks' performances (as shown in Fig.~\SFpef{}) on face pair matching tasks---accuracy of identifying pairs of different pictures from the same person and rejecting those from different persons---against the Labeled Faced in the Wild (LFW/LFW-a) dataset \cite{LFWTech, wolf2011effective}.
-Stimulus dimensionalities of shallow and deep neurons are $N=121$ and $N=441$ (i.e.~$11\times11$ and $21\times21$ receptive field sizes) respectively.
 

### 2014-12-22 abstract.tex
More edits

@@
 
-A central challenge in sensory neuroscience is to describe how the activity in populations of neurons represents properties of the external environment in a format suitable for guiding behavior.
+A central challenge in sensory neuroscience is describing how the activity of populations of neurons can ``represent'' useful features of the external environment.
 However, while neurophysiologists have long been able to record the responses of neurons in awake, behaving animals, it is another matter entirely to say what a given neuron ``does.''
-A key problem is that in many sensory systems, the space of all possible stimuli that one might encounter is effectively infinite.
-For instance, in vision, scenes are combinatorially complex---an individual human will encounter only a small fraction of the possible scenes in the world; indeed only a small fraction of the possible images have \emph{ever} been experienced in the collective history of the human race.
-Thus, investigations of neuronal function---real and simulated---are almost always critically limited by the number of stimuli that can be considered. 
-In some early stages of sensory processing, a combination of intuition and systems identification theory has led to reasonable models of early neuron response functions (e.g.
-Gabor wavelet models in cortical area V1), but cascading nonlinearities have tended to foil our best efforts in later stages of sensory processing.
+A key problem is that in many sensory domains, the space of all possible stimuli that one might encounter is effectively infinite; in vision, for instance, natural scenes are combinatorally complex, and an organism will only encounter a tiny fraction of possible stimuli.
+As a result, even describing the response properties of sensory neurons in difficult, and investigations of neuronal function are almost always critically limited by the number of stimuli that can be considered.
 
-Here, we propose a closed-loop experimental framework for characterizing the response properties of sensory neurons, building on past efforts in closed-loop experimental methods, and leveraging recent advances in artificial neural network models to serve as as a proving ground for our techniques.
-To the extent that modern deep learning models increasingly serve as \emph{de facto} working hypotheses for biological vision \ref{jim, others}, we argue that a unified approach of developing methods to interrogate
+Here, we propose a closed-loop, optimization-based experimental framework for characterizing the response properties of sensory neurons, building on past efforts in closed-loop experimental methods, and leveraging recent advances in artificial neural network models to serve as as a proving ground for our techniques.
+Using ``deep'' convolutional neural networks as a proving ground, we asked whether modern black-box optimization techniques can be used to interrogate the response landscape of an artificial neuron in a deep, nonlinear system, without imposing significant constraints on the space of stimuli under consideration.
+We introduce a series metrics to quantify neuronal response landscapes, and show how these relate to the performance of the network in an object recognition task.
+To the extent that deep learning models increasingly serve as \emph{de facto} working hypotheses for biological vision \ref{jim, others}, we argue that developing a unified approach for studying both natural and artifical systems holds great potential to advance both fields together.

### 2014-12-22 introduction.tex
More edits

@@
-A central drive in the study of sensory systems is the production of models that
+A central drive in sensory systems neuronscience is to produce models that explain the relationship between sensory stimuli on one hand, and the firing of individual neurons and nerounal populations on the other.
+However, a major problem in many sensory domains is that the space of all possible stimuli is effectively infinite.
+For instance, in vision, natural scenes are combinatorially complex---an individual human will encounter only a small fraction of the possible scenes in the world; indeed only a small fraction of the possible images have \emph{ever} been experienced in the collective history of the human race.
+In some early stages of sensory processing, a combination of intuition and systems identification theory has led to reasonable models of early neuron response functions (e.g.
+Gabor wavelet models in cortical area V1), but cascading nonlinearities have tended to foil our best efforts in later stages of sensory processing.
 
-The mammalian ventral visual pathway consists of a hierarchical cascade of visual areas that progressively reformat visual information from pixel-like retinotopic representations into formats suitable for high-level tasks \cite{dicarlo2007untangling, dicarlo2012does}, such as recognizing visual objects.
-Recently, computer vision algorithms inspired by the hierarchical structure of the visual system---convolutional neural networks \cite{fukushima1980neocognitron, lecun1998gradient, riesenhuber1999hierarchical}, or so-called ``deep learning'' approaches \cite{krizhevsky2012imagenet}---have become the focus of tremendous attention in the computer vision and machine learning communities, due to their success in a variety of practical domains.
-However, we are still far from achieving human-level capabilities for unconstrained visual tasks \cite{ghodrati2014feedforward}, and even for the most successful current deep learning systems which, on controlled visual tasks, perform relatively close to human subjects \cite{serre2007feedforward, cirecsan2012multi, russakovsky2014imagenet, taigman2013deepface, sun2014deep, viglarge} and produce representations similar to mammalian visual systems' \cite{yamins2014performance, cadieu2014deep}, inexplicable failures still occur \cite{szegedy2013intriguing}.
-Undeniably, we lack a comprehensive functional understanding of how these system work at the theoretical level.
+% The mammalian ventral visual pathway consists of a hierarchical cascade of visual areas that progressively reformat visual information from pixel-like retinotopic representations into formats suitable for high-level tasks \cite{dicarlo2007untangling, dicarlo2012does}, such as recognizing visual objects.
+
+Recently, computer vision algorithms inspired by the hierarchical structure of the visual system---convolutional neural networks \cite{fukushima1980neocognitron, lecun1998gradient, riesenhuber1999hierarchical}, and so-called ``deep learning'' approaches \cite{krizhevsky2012imagenet}---have become the focus of tremendous attention in the computer vision and machine learning communities, due to their success in a variety of practical domains.
+In particular, these systems can perform on par with human subjects on certain controlled visual tasks \cite{serre2007feedforward, cirecsan2012multi, russakovsky2014imagenet, taigman2013deepface, sun2014deep, viglarge} and they produce internal representations that are similar to that of mammalian visual systems under certain conditions \cite{yamins2014performance, cadieu2014deep}.
+
+However, such networks are still far from achieving human-level capabilities for unconstrained visual tasks \cite{ghodrati2014feedforward}, and even for relatively constrained object categorization tasks, surprising failures can be induced through subtle stimulus manipulation \cite{szegedy2013intriguing}.
+More fundamentally, we lack a comprehensive functional understanding of how these system work, at a theoretical level.
 Here, we take the success of artificial ``deep learning'' architectures as an opportunity to develop tools for studying and gaining insight about the function of deep networks of nonlinear units---tools that can in principle be used to study both artificial and real neuronal systems.

### 2015-03-27 paper_for_review.tex
v1.0

\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{iccv}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

% Include other packages here, before hyperref.
\usepackage[labelfont=bf]{caption}
\captionsetup{labelsep=period}

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}

% \iccvfinalcopy % *** Uncomment this line for the final submission

\def\iccvPaperID{****} % *** Enter the ICCV Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

% Pages are numbered in submission mode, and unnumbered in camera-ready
\ificcvfinal\pagestyle{empty}\fi
\begin{document}

%%%%%%%%% TITLE
\title{What Makes Deep Network's Representations Good for Visual Recognition?}

\author{First Author\\
Institution1\\
Institution1 address\\
{\tt\small firstauthor@i1.org}
% For a paper whose authors are all at the same institution,
% omit the following lines up until the closing ``}''.
% Additional authors and addresses can be added with ``\and'',
% just like the second author.
% To save space, use either the email address or home page, not both
\and
Second Author\\
Institution2\\
First line of institution2 address\\
{\tt\small secondauthor@i2.org}
}

\maketitle
%\thispagestyle{empty}

\begin{abstract}

Representations of artificial neurons in deep neural networks have been demonstrated highly effective and robust in various visual recognition tasks \cite{krizhevsky2012imagenet, sermanet2013overfeat, donahue2014decaf}.
Nevertheless, our understanding of them arguably remains limited.
In this paper, we address two important and practical questions: (1) why deep networks are better than shallow networks, and (2) among networks of the same depth, why certain networks are better than the others.
Using methods which support characterizing both first-order and second-order properties of scalar and vector representations, we conducted a large-scale experiment, with 200 randomly generated shallow and deep networks, or totally 6,400 neurons characterized, and the results analyzed through a series of measures designed to quantify the differences in representations. 
We conclude that deep representations are significantly better than shallow representations, and overall 71\% of the variance in deep networks' performances can be explained with our measures, both of which are confirmed using statistical tests for the first time.

%(\ie optimal stimulus) (\ie invariance and selectivity paths)
%use X as example of deep networks
%concern about deep networks; how alleviated via this work

\end{abstract}

\section{Introduction}

\begin{figure}
\begin{center}
\includegraphics[width=0.35\textwidth]{Figs/e_fig7s_compact-crop.pdf} 
\end{center}
\caption{{\bf What makes a deep network good?}
In existing literatures, researchers mostly tried to come up with one (or few) ``best'' network architecture (\ie hyperparameter set, either through manual tuning or automated hyperparameter search) and then benchmarked the network's performances.
However, as exemplified in the figure, when considering a larger population of networks \cite{cox2011beyond}, it is not fully clear to us why deep (\ie two-level) networks in practice outperform shallow (\ie one-level) networks, and why certain networks are better than the others, given their broad distribution of performances on LFW-a face recognition \cite{LFWTech, wolf2011effective}.
%However, it is not fully clear that in practice (1) why deep networks are better than shallow networks, and (2) among networks of the same depth, why certain networks are better than the others.
%As exemplified in the figure, deep (\ie two-level) networks not only on average outperform shallow (\ie one-level) networks \cite{cox2011beyond}, but also exhibit a wide range of performance differences on LFW-a face recognition \cite{LFWTech, wolf2011effective}.
} % (\eg optimal stimulus). %and (sometimes) characterized its properties
\label{fig:perf}
\end{figure}

Although deep neural networks have been successfully pushing the boundaries of computer vision by breaking the records of ImageNet Challenges (ILSVRC) year after year recently \cite{krizhevsky2012imagenet, sermanet2013overfeat, szegedy2014going}, our understanding about the properties of these networks arguably remains limited, since we could not yet fully answer the question ``what makes a deep network good?''.
As demonstrated in Fig.~\ref{fig:perf}, we summarized two important subquestions to be investigated: (1) why deep networks are better than shallow networks, and (2) among networks of the same depth, why certain networks are better than the others.
Specifically, from the point of view that successful networks are forming ``better'' representations of visual inputs that are more separable to the final-stage linear classifiers, we want to identify the crucial properties that are required in good representations.

%\newcommand{\ivec}{\mathrm{vec}^{-1}}
%$f({\bf{x}}) = \left\| {\bf{K}} \otimes \ivec\left({\bf{x}}\right) \right\|_{F}$
\newcommand{\expstimdim}{For simplicity of formulas, a stimulus being 2-dimensional (\ie ${\bf{x}} \in \mathbb{R}^{\sqrt{N} \times \sqrt{N}}$) or vectorized (\ie ${\bf{x}} \in \mathbb{R}^N$) are used interchangeably.}

%first order, second order, flexibility
{\bf Existing works.} 
Given an artificial neural network $f$, the studying of how $f$ represents a high dimensional stimulus\footnote{\expstimdim} ${\bf{x}} \in \mathbb{R}^N$ as a representation 
${\bf{r}} = f\left( {\bf{x}} \right)$ in existing literatures mainly focused on the first-order (\ie linear) characterization of $f$---the analysis of optimal stimuli, stimuli which lead to the strongest responses of given neurons.
For single-level convolutional networks using squared pooling, \ie $f({\bf{x}}) = \left\| {\bf{K}} \otimes {\bf{x}} \right\|_{F}$, optimal stimulus was shown to be analytically approximable \cite{saxe2011random} using a Gabor-like filter, whether the convolution kernel $\bf{K}$ is structural or random.
For multi-level convolutional networks, optimal stimuli of neurons from various depths were numerically approximated \cite{ngiam2010tiled, le2012building, zeiler2014visualizing, simonyan2013deep} to qualitatively show neurons in deeper layers are tuned to more complex visual patterns.
The generalized concept of optimal stimulus in considering groups of neurons can also be adopted to study the representations by inverting them \cite{mahendran2014understanding}.

Beyond the first-order characterization, the analyses of invariance and selectivity properties of neurons was adopted in a few works, which can be viewed as the second-order characterization of $f$, and parametric deformation in stimuli was the most commonly used setup.
For example, translation, rotation, scaling, \etc in stimulus were used in \cite{goodfellow2009measuring, zeiler2014visualizing} to test the invariance and selectivity of neurons.
Similar deformations were also adopted in \cite{lenc2014understanding}, but extended to analyze more advanced properties (\ie equivariance and equivalence) of groups of neurons.
More generalized setups that, \eg, utilize local quadratic information (\ie the Hessian) or ``global'' numerical searches in the stimulus space, even though do exist, were less explored in the context of artificial neural network studies.
For quadratic networks, \ie $f\left({\bf{x}}\right) = \frac{1}{2}{\bf{x}}^{T}{\bf{Qx}}+{\bf{L}}^{T}{\bf{x}}+c$, through eigendecomposing the quadratic term $\bf{Q}$ (\ie the Hessian), local invariance and selectivity directions can be analytically derived \cite{berkes2006analysis}, which correspond to eigenvectors of the least and most negative eigenvalues.
This method was also generalized onto multi-layer convolutional networks through numerically estimating the Hessian \cite{ngiam2010tiled}.
For multi-layer networks, characterization of the ``global'' invariance of single neurons can also be achieved through numerically searching for non-local (\ie farther to the optimal stimuli) solutions \cite{erhan2010understanding}.

%with its frequency component collocated with the peak of the Fourier spectrum of convolution kernel $\bf{K}$
%(\ie farther to the optimal stimuli)

%{\bf Contributions of this work.} Comparatively, the characterization of representations in this work is designed to be general and unbiased such that both first- and second-order methods can be supported for single neurons or groups of neurons, without assuming any \emph{de facto} property in the solutions (\eg parametric deformations as used for studying invariance in previous works).
%Additionally, with the strong statistical significance of results obtained from the large-scale experiments on 100 shallow and 100 deep networks, totally 6,400 artificial neurons, we argue that our observations toward answering the differences between (shallow \vs deep) and among (deep) representations are highly evident and comprehensive, which cab be summarized as ***

\section{Methods}

\begin{figure}
\begin{center}
\includegraphics[width=0.35\textwidth]{Figs/methods.pdf} 
\end{center}
\caption{{\bf Representation characterization methods.}
On the spherical (\ie constant energy) constraint of the $N$ dimensional stimulus space, similar to previous works, the first-order property (\ie optimal stimulus $\hat{\bf{x}}$, as shown in the figure's center) of a representation $\bf{r}$ is first iteratively searched (as the evolution of light gray eclipses shown in the figure) and analyzed.
However, unlike most previous works, the second-order properties (\ie invariant path $\lbrace \bf{x}^{+}_{\delta} \rbrace$ as dashed red curve and selectivity path $\lbrace \bf{x}^{-}_{\delta} \rbrace$ as dashed blue curve) at distances $0.1\pi$ to $0.5\pi$ from the optimal stimulus are further searched and analyzed.
Additionally, these methods are capable of characterizing both scalar and vector representations, where the ``fitnesses'' to be optimized correspond to Eq.~\ref{eq:O1}, \ref{eq:I1}, \ref{eq:S1} and Eq.~\ref{eq:O2}, \ref{eq:I2}, \ref{eq:S2}, respectively.
In this example, the scalar representation (\ie response) of the given neuron is tuned to a $45^{\circ}$ Gabor filter and invariant to phase changes, but selective to orientation changes.
}
\label{fig:methods}
\end{figure}

% DEFINE X & f????
%representational property

To study the representations deep inside an artificial neural network, methods adopted in this work include: (1) first-order (\ie linear) characterization, or search and analysis of optimal stimuli, and (2) second-order (\ie quadratic) characterization, or search and analysis of invariance and selectivity paths (\ie eigenvectors of the largest and smallest eigenvalues), which are generalized and improved from previous works.
$\bf{r}$ is defined as a scalar representation (\ie $\bf{r} \in \mathbb{R}$) when considering single neurons and a vector representation (\ie ${\bf{r}} \in \mathbb{R}^C$) when considering groups of neurons (with group size $C$).
Our methods support characterizing both scalar (using Eq.~\ref{eq:O1}, \ref{eq:I1}, \ref{eq:S1}, though Eq.~\ref{eq:O2}, \ref{eq:I2}, \ref{eq:S2} can be used as well) and vector (using Eq.~\ref{eq:O2}, \ref{eq:I2}, \ref{eq:S2}) representations in a unified way.
The search phases defined via these equations will be introduced in the following subsections, as the analyses of search results using representation measures will be covered in Sec.~\ref{sec:exp} when applied.

% SEARCH HERE, ANALTSIS MEASURES LATER
%$\bf{r}$, in this work, is equivalent to the response(s) of artificial neuron(s), and
%Arguably, for studying neural networks, especially when regarding task-related (\eg recognition) performances, the importance of characterizing vector representations can be greater than of characterizing scalar representations, considering the differences in representational powers.

While no specific knowledge of the network itself is assumed, we do restrict the space of stimuli considered with a constant energy constraint $\left\| \bf{x} \right\| = E$, where $E$ is set to $1$ in all mathematical derivations for simplicity, but valued as the average energy of task-related images in our experiments.
Since neurons can often be modulated by stimulus contrasts, which is less interesting when considering pattern selectivities, limiting the stimulus space in this manner effectively avoids degenerate solutions that simply maximize stimulus contrasts, and reduces the range of possible stimuli to consider.

%, the setting of $E=1$ is used for the rest of the paper, while  $E$ {was} set 
%, based on prior knowledge of (artificial) neuronal response properties.

\subsection{First-Order Property Characterization}

In this work, optimal stimulus is defined and numerically derived as 
\begin{align}
\hat{\bf{x}} &= \underset{\bf{x}} {\arg\max} f\left(\bf{x}\right) \label{eq:O1} \\
&= \underset{\bf{x}}{\arg\max} \left( e^{-\left\|f\left(\bf{x}\right)-f\left(\tilde{\bf{x}}\right)\right\|} \right); \label{eq:O2}
\end{align}
subject to the spherical constraint $\left\| \bf{x} \right\| = 1$. 
As visualized in Fig.~\ref{fig:methods}, this process involves an iterative search on the non-convex fitness landscape, which in fact can be handled efficiently by most of the modern numerical solvers, especially when analytical derivatives (\ie backpropagation gradients) are available.
In this work, we adopted the CMA-ES solver \cite{hansen2001completely}, which though do not rely on analytical gradients and instead stochastically estimates numerical derivatives (\ie inverse Hessian), {worked} reasonably well in our experiments.
When considering a vector representation (using Eq.~\ref{eq:O2}), this process is equivalent to optimizing the response of an ``imaginary neuron'' tuned to a reference stimulus $\tilde{\bf{x}}$.

% $\exp\left( -\left\|f\left(\bf{x}\right)-f\left(\tilde{\bf{x}}\right)\right\| \right)$
%Or equivalently, optimal stimulus of an imaginary neuron tuned to the

\subsection{Second-Order Property Characterization}

\newcommand{\expdiff}{The way the simple linear constraint is constructed to enforce the exploration of a larger extent of the fitness landscape is also one of the main differences compared to \cite{erhan2010understanding}.}

With respect to the optimal stimulus $\hat{\bf{x}}$, the searches of invariance and selectivity paths, which consist of sets of invariant stimuli $\left\lbrace \bf{x}^{+}_{\delta} \right\rbrace$ and selective stimuli $\left\lbrace \bf{x}^{-}_{\delta} \right\rbrace$ respectively, are formulated as
\begin{align}
\bf{x}^{+}_{\delta} &= \underset{\bf{x}_{\delta}}{\arg\max} f\left(\bf{x}_{\delta}\right) \label{eq:I1} \\
&= \underset{\bf{x}_{\delta}}{\arg\max} \left( e^{-\left\|f\left(\bf{x}_{\delta}\right)-f\left(\hat{\bf{x}}\right)\right\|} \right); \label{eq:I2} \\
\bf{x}^{-}_{\delta} &= \underset{\bf{x}_{\delta}}{\arg\min} f\left(\bf{x}_{\delta}\right) \label{eq:S1} \\
&= \underset{\bf{x}_{\delta}}{\arg\min} \left( e^{-\left\|f\left(\bf{x}_{\delta}\right)-f\left(\hat{\bf{x}}\right)\right\|} \right); \label{eq:S2}
\end{align}
in which $0 < \delta \le \frac{\pi}{2}$, subject to $\left\| \bf{x}_{\delta} \right\| = 1$ and $\langle \bf{x}_{\delta} , \hat{\bf{x}} \rangle = \cos\left(\delta\right)$. 
As visualized in Fig.~\ref{fig:methods}, the invariance path characterizes the $N$ dimensional curve that leads toward (or ideally, maintains at) fitness as high as possible while moving away from the optimal stimulus, and the selectivity path characterizes the curve that leads toward fitness as low as possible while moving away.
The search process is implemented as multiple runs of maximization and minimization on discretized $\delta \in \left\lbrace 0.1\pi, 0.2\pi, 0.3\pi, 0.4\pi, 0.5\pi\right\rbrace$ as the circular distance constraints shown in Fig.~\ref{fig:methods}, where each run is initialized with the result from previous run (and the $0.1\pi$ run directly with optimal stimulus $\hat{\bf{x}}$) to increase the path continuity and searching speed\footnote{\expdiff}.
In this work, the distance constraint $\delta$ only goes up to $\frac{\pi}{2}$, since on the $N$ dimensional sphere, stimuli fall in $\frac{\pi}{2} < \delta \le \pi$ are simply ``negatives'' of those in $0 < \delta \le \frac{\pi}{2}$, which are of less uniqueness and interest in general. 
The same process can also be performed with respect to a reference stimulus $\tilde{\bf{x}}$, especially for vector representations, where the invariance and selectivity of the ``imaginary neuron'' are again characterized equivalently.

%; nevertheless, going up to the full range $0 < \delta \le \pi$ is numerically straightforward.

\section{Experiments}
\label{sec:exp}
%\subsection{Experimental Setup}
% \cite{fukushima1980neocognitron, lecun1998gradient, riesenhuber1999hierarchical, krizhevsky2012imagenet}

For our experiments, a simplified variety of deep convolutional neural networks---the Spatio-Temporal Hierarchical Object Representation (STHOR) model \cite{pinto2009high, sthor}---{was} adopted.
Although like other deep networks, it also consists of the standard cascade of convolution, nonlinear activation, pooling layers, \etc (which together define a single ``level'' in this work), its convolution kernel weights are however randomly assigned rather than trained.
Though not matching the performances of deep networks trained on massive quantities of data \cite{krizhevsky2012imagenet}, it can in fact be surprisingly competitive for a variety of tasks, especially in the regime where large quantities of training samples are not easily available \cite{pinto2009high, cox2011beyond, viglarge}.
Most importantly, since it does not require training (except for the linear SVM on top of the representations extracted), a large number of networks with diverse hyperparameters can be rapidly generated and compared, which is particularly useful in this study where relative, instead of absolute task performances, are of the main interest. 
As one might concern that comparing randomly weighted networks will not tell much about representations, since they might as well be random, our results also suggest otherwise.

\newcommand{\expsettings}{In other words, 32 channels of filters in top levels' convolution layers.
Following the terminology in recent literatures, shallow and deep neurons corresponded to \texttt{pool1} and \texttt{pool2} layers respectively.
Stimulus dimensionalities of the shallow and deep neurons {were} $N=121$ and $441$ (\ie spatially overlapping $11\times11$ and $21\times21$ receptive fields). % respectively
Other minor changes included: nonlinear activations all simplified to \emph{ReLU} \cite{krizhevsky2012imagenet} mode and normalizations all conditionally subtractive mode.
Overall, the architectures were more similar to those in \cite{simonyan2014very}, except that pooling operations can be \emph{average}, \emph{squared}, or \emph{max-like} in our case.
}

In the experiments, 100 shallow (\ie one-level) networks and 100 deep (\ie two-level) networks {were} randomly generated (except all with 32 top-layer neurons\footnote{\expsettings}) and tested to see how their representations differ with networks' depths and affect networks' performances on face pair matching tasks against the LFW-a dataset \cite{LFWTech, wolf2011effective} as shown in Fig.~\ref{fig:perf}. %, using 10-fold cross validation on view 2.
The fitness landscapes of shallow and deep representations in our case are both non-convex, therefore there is no guarantee that optimality can be achieved through any optimization method.
However, by carefully adjusting the numerical solver and monitoring the quality of solutions, we argue that informative local optima can still be discovered and used to characterize the representations.
Specifically, all numerical optimizations {were} executed multiple times from different random initial stimuli, for increasing quality of the numerical solutions (of optimal stimuli and invariance and selectivity paths, with 2 runs executed and the better kept), and at the same time, for providing statistical samplings of the solution spaces (of encoding specificities and invariance and selectivity subspace properties, as detailed later).

% (as shown in Fig.~\ref{fig:perf})
%---accuracy of identifying pairs of different pictures from the same person and rejecting those from different persons---

\subsection{Results: Shallow \textbf{\textit{vs.}}~Deep Representations}

\begin{figure*}
\begin{center}
\includegraphics[width=0.70\textwidth]{Figs/pic1.pdf} 
\end{center}
\caption{{\bf Visualization of properties of shallow (A--D) and deep scalar representations (E--H).}
Color of the boarder of an image indicates the fitness of the representation (\ie response of the neuron) driven by the image, with color map following the definition in Fig.~\ref{fig:methods}.
A and E show the search trajectories of \emph{optimal stimuli} of all 32 neurons from the best performing shallow and best performing deep networks, respectively, where A.1 and E.1 show the initial $1/f$ random stimuli, A.5 and E.5 the resultant optimal stimuli, and A.2--4 and E.2--4 three intermediate steps corresponding to the three largest curvatures in the fitness histories.
B and F show the \emph{invariance paths} of shallow and deep neurons, starting from the corresponding optimal stimuli show in A.5 and E.5, moving from $0.1\pi$ (B.1 and F.1) to $0.5\pi$ (B.5 and F.5) away accordingly.
C and G show the \emph{selectivity paths} of shallow and deep neurons, with definitions following B and F.
D and H show the \emph{invariance subspaces} sampled $0.1\pi$ away from the optimal stimuli, with 5 out of 20 results randomly selected.
Comparatively, \emph{deep neurons show more complex and diverse representational properties} than shallow neurons do.
} % do visually
\label{fig:allrep}
\end{figure*}

\begin{figure}
\begin{center}
\includegraphics[width=0.35\textwidth]{Figs/e_fig3a_2-crop-nup-crop.pdf} 
\end{center}
\caption{{\bf Fitness-distance analyses of shallow (A) and deep scalar representations (B).}
Fitnesses of invariance (red dots, \ie $\left\lbrace f\left(\bf{x}^{+}_{\delta}\right) \right\rbrace$) and selectivity paths (blue dots, \ie $\left\lbrace f\left(\bf{x}^{-}_{\delta}\right) \right\rbrace$) of all 3,200 shallow and 3,200 deep neurons are shown for comparison.
A small fraction of invariance paths in fact have fitnesses higher than the optimal stimuli's due to a non-convex problem's nature, which however does not cause noticeable differences in the resultant statistics.
While both types of neurons show invariance, shallow neurons do not posses selectivity, since their selectivity paths overlap with the dashed line---the zero-selectivity baseline.
In other words, \emph{only deep neurons are selective}.
} %, and random walks (green dots, \ie $\left\lbrace f\left({\bf{x}}^{R}_{\delta}\right) \right\rbrace$)
\label{fig:fda}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=0.35\textwidth]{Figs/e_fig23b-crop.pdf} 
\end{center}
\caption{{\bf Comparison of shallow and deep scalar representation measures.}
Representation measures from left to right are, optimal stimulus spectral complexity (OSSC), invariance path potential (INPP), selectivity path potential (SLPP), and invariance subspace capacity (INSC), in which data points from shallow networks (left distributions) and deep networks (right distributions) are shown side by side.
Sensitivity indexes ($d'$) between distributions are $1.63$, $0.42$, $4.73$, and $3.67$, respectively.
}
\label{fig:pair}
\end{figure}

{\bf Optimal stimulus spectral complexity.} Figure \ref{fig:allrep}A and \ref{fig:allrep}E illustrate the trajectories of optimal stimulus searches of 32 top-layer neurons from the best performing shallow network and 32 from the best performing deep network respectively.
To further increase the searching speed, the initial stimulus {was} selected from 1000 ${1}/{f^{\alpha}}$ random stimuli with the best fitness, where $\alpha \in \left\lbrace -4,-3,-2,-1,0 \right\rbrace$, without overly sacrificing the nature of random initialization.
From the results, first, the response landscapes of deep neurons are noticeably more complex than those of shallow neurons, as a large fraction of searching trajectories for deep neurons actually go down then up before reaching the (local) optima, which also suggests the numerical solver can handle non-convexity reasonably well.
Second, the optimal stimuli of deep neurons are also relatively more complex (\ie consisting of more textural components) than those of shallow neurons (which are more uni-textured or Gabor-like).
To quantify such differences, the spectral complexity measure, which is estimated through the $L^{1}$ norm of the (2 dimensional) Fourier power spectrum of optimal stimulus, \ie $\left\| \mathcal{F}(\hat{\bf{x}}) \right\|_{1}$, is adopted, where higher value suggests higher non-sparsity (\ie more spectral components required to represent the visual stimuli).

\newcommand{\defbaseline}{Given a single inner-product neuron $f({\bf{x}}) = {\bf{w}}^{T}{\bf{x}}$ and $\left\| \bf{x} \right\| = 1$, we have $\hat{\bf{x}} = \bf{w}$ and thus $f({\bf{x}}_{\delta}) = {\hat{\bf{x}}}^{T}{\bf{x}}_{\delta} = \cos(\delta)$ by definition.}

{\bf Invariance and selectivity path potential.} Figure \ref{fig:allrep}B/C and \ref{fig:allrep}F/G illustrate the invariance and selectivity paths on the same sets of shallow and deep neurons as in \ref{fig:allrep}A and \ref{fig:allrep}D.
While invariance paths are mostly phase changes and selectivity paths are leading toward meaningless noises (all at the same falloff rate) for shallow neurons, both types of paths consist of sophisticated shape deformations for deep neurons.
As further demonstrated in the fitness-distance diagrams \cite{jones1995fitness} (Fig.~\ref{fig:fda}), on average, although shallow neurons have good invariance (\ie red curve stays high), they do not have any selectivity (\ie blue curve drops slow), while deep neurons show both good invariance and selectivity (\ie red curve stays high and blue curve drops fast).
For scalar representations, we can define the invariance (and selectivity, similarly) path potential as the area sandwiched between the invariance curve and the cosine baseline curve (\ie the invariance and selectivity curves of a single inner-product neuron\footnote{\defbaseline}, or the ``baseline'' of neural network types) on the fitness-distance diagram, \ie $\int_{0}^{\frac{\pi}{2}}{\left| \cos^{-1}\left(f\left( \bf{x}^{+}_{\delta} \right)\right) - \delta \right|}\mathrm{d}\delta \mathbin{/} {\frac{\pi}{2}}$ as normalized in the $\cos^{-1}$ domain, to measure how invariant (and selective) a neuron is compared to the baseline.

%(zero-invariance/-selectivity) 
%Also, although most shallow neurons are selective to manually rotated optimal stimuli (i.e.~Gabor filters of different orientations) as well, their fitnesses still do not drop faster then the nonparametric numerical solutions as presented in Fig.~\ref{fig:allrep}C.
%In fact, this intriguing difference generalizes across all shallow \vs deep neurons experimented in this work.
%Random walks on the fitness landscape \cite{jones1995fitness}, shown as the green curve in Fig.~\ref{fig:ind_fdd}, can be defined as $f\left(\cos\left(\delta\right)\hat{\bf{x}} + \sin\left(\delta\right)\tilde{\bf{x}}\right)$ where $\tilde{\bf{x}}$ is any random stimulus such that $\left\| \tilde{\bf{x}} \right\| = 1$ and $\left\langle \hat{\bf{x}},\tilde{\bf{x}} \right\rangle = 0$, which can be obtained through random projection of $\mathrm{Null}\left(\hat{\bf{x}}\right)$. 
%Comparing the gaps between selectivity and random-walk curves in Fig.~\ref{fig:fda}A and \ref{fig:fda}B again shows the selectivity of deep neurons is in fact way more ``selective'' (i.e.~not noise-like) than of shallow neurons.

{\bf Invariance subspace capacity.} To further characterize the invariance properties, multiple runs of invariance path searches {were} performed and the results are visualized in Fig.~\ref{fig:allrep}D and \ref{fig:allrep}H.
The diversity of the ``invariance plateau'' (\eg the central high fitness region in Fig.~\ref{fig:methods}) constitution of deep neurons is also more apparent than that of shallow neurons.
Compared to invariance path potential which is designed to characterize the best invariance path even if only one of such exists, we further adopted the invariance subspace capacity measure to estimates the ``dimensionality'' (\ie how diverse different paths can be) of the linear subspace formed by multiple path search results via the nuclear norm of concatenation of $n$ results $\left\|\left[{\bf{x}}_{\delta,1},\dots,{\bf{x}}_{\delta,n}\right]\right\|_{*}$ with $n=20$ and $\delta=0.1\pi$.

%(\ie dimensionality, or subspace capacity)

\newcommand{\defstars}{In this paper, significance levels follow the definition that *, ** and *** correspond to $p<0.05$, $p<0.01$ and $p<0.001$ respectively.}

{\bf Overall comparison.} Comparisons of spectral complexities, invariance/selectivity path potentials, and invariance subspace capacities of shallow and deep neurons (as scalar representations) are summarized in Fig.~\ref{fig:pair}.
These measures further reveal the significant\footnote{\defstars} differences between shallow and deep representations, in addition to the separability and classifiability used in previous studies \cite{donahue2014decaf, zeiler2014visualizing}.
For instance, subtle differences of visual features have better chances being distinguished when the ``gap'' between invariance and selectivity curves (\ie the dynamic range of neuronal responses, or amplification ratio of differences) is large. 
The highly structural representations of neurons from deep networks with purely random convolution kernels found in the experiments also explain why such ``random'' deep networks can still have good performances, on top of the theoretical analysis for (one-level) shallow networks \cite{saxe2011random}.

%The increase of complexities in deeper neurons' preferred/optimal stimuli also agrees with neurobiological findings in higher visual cortical areas (\eg V2 and V4 \cite{hegde2000selectivity, pasupathy2001shape}).

\subsection{Results: Good \textbf{\textit{vs.}}~Bad Representations}

\begin{figure*}
\begin{center}
\includegraphics[width=0.70\textwidth]{Figs/pic2.pdf} 
\end{center}
\caption{{\bf Representativities (\ie explanation power) of optimal stimuli of deep neurons.}
Results from deep networks of the top 3 strongest average explanation power (A--C) and the weakest average explanation power (D) are visualized.
Within each panel, the first row shows all 32 neurons' optimal stimuli from the network, the third to fifth rows show natural images with the top 3 strongest ``explainablilities'' (\ie inner-product distances between the images and the corresponding optimal stimuli), and the second row shows the average of images with top 100 explainablilities, out of 10,000 natural images.
Color of the boarder of an image indicates the explainablility of the image, with color map following the definition in Fig.~\ref{fig:methods}.
Comparing deep networks' explanation power and performances shows significant correlation (OSEP measure in Fig.~\ref{fig:corr}), which suggests \emph{``representing'' natural images can be required in first-order properties of good deep scalar representations (\ie optimal stimuli)}.
}
\label{fig:exppow}
\end{figure*}

\begin{figure}
\begin{center}
\includegraphics[width=0.35\textwidth]{Figs/pic3.pdf} 
\end{center}
\caption{{\bf Alignments of subspaces between invariance and selectivity paths and natural images.}
Using 16 natural images in A as reference stimuli, the invariance paths with $\delta = 0.1\pi$ corresponding to the best alignments against natural images (\ie $L^1$ sparsities of images represented in eigenface coordinates) among all deep networks are shown in B.1, and similarly, worst alignments of invariance paths in C.1, best alignments of selectivity paths in D.1, and worst alignments of selectivity paths in E.1.
Differences between A and B.1--E.1 are shown in B.2--E.2 for clearer visualization.
Color of the boarder of an image indicates the alignment of the image against natural images, with color map following the definition in Fig.~\ref{fig:methods}.
As good alignments (B.1 and C.1) appear more natural (like structural deformations or lighting changes), bad alignments are mostly meaningless noises.
Comparing deep networks' subspace alignment measures and performances shows significant correlations (ITSA/STSA measures in Fig.~\ref{fig:corr}), which suggest \emph{resembling natural image distributions can be required in second-order properties of good deep vector representations (\ie invariance and selectivity subspaces)}.
} %of subspaces % which are known to be important factors in visual recognition % accordingly
%Multiple runs of invariance and selectivity path searches are executed to form samples of the subspaces.
\label{fig:align}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=0.35\textwidth]{Figs/pic4.pdf} 
\end{center}
\caption{{\bf Reconstructability of deep vector representations (\ie encoding specificities of natural images).}
Using 16 natural images in A as reference stimuli, 3 examples of reconstructions from all deep networks with the best, second best, third best and worst average reconstructed image quality (SSIM) are shown in B--E respectively.
Color of the boarder of an image indicates the SSIM score between the image and reference stimulus, with color map following the definition in Fig.~\ref{fig:methods}.
Though the reconstructions are in general of lower spatial frequencies, some of the better ones do capture the spatial structures of reference stimuli decently well (\eg stimulus 1, 5, \etc).
Comparing deep networks' encoding specificities and performances shows significant correlations (ITSA/STSA measures in Fig.~\ref{fig:corr}), which suggest \emph{providing specific (\ie non-confounding) encoding of natural images can be required in first-order properties of good deep vector representations}.
} % (reconstructed)
\label{fig:encspc}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=0.95\columnwidth]{Figs/e_fig7_scatter-crop.pdf} 
\end{center}
\caption{{\bf Spearman's correlations between deep network's representation measures and performance.}
Representation measures following the raster order are, optimal stimulus explanation power (OSEP), invariance path potential (INPP), selectivity path potential (SLPP), invariance subspace capacity (INSC), invariance \vs task-related image subspace alignment (ITSA), selectivity \vs task-related image subspace alignment (STSA), task-related image encoding specificity (TSES), and all 7 measures together using linear multiple correlation analysis.
Correlations $\rho$ and significance levels are $0.60^{\ast\ast\ast}$, $-0.31^{\ast\ast}$, $-0.24^{\ast}$, $-0.39^{\ast\ast\ast}$, $-0.44^{\ast\ast\ast}$, $-0.56^{\ast\ast\ast}$, $0.64^{\ast\ast\ast}$, and $0.84^{\ast\ast\ast}$, respectively.
Overall, \emph{71\% of the variance ($\rho^2$) of deep networks' performances can be explained} using the proposed representation measures altogether.
}
\label{fig:corr}
\end{figure}

{\bf Optimal stimulus explanation power.} To understand how optimal stimulus may affect a network's performance, we {tested} how well the optimal stimulus may ``linearly explain'' natural images and how the explanation power correlates with network's performance.
Totally 10,000 image patches {were} randomly cropped out of the predefined facial regions of pictures from the LFW-a dataset \cite{wolf2011effective} and compared against the optimal stimuli of all 32 top-layer neurons from all 100 deep networks. 
Figure \ref{fig:exppow} shows the results of best 3 and worst 1 networks in terms of explanation power, which is estimated as the mean of the linear explainability of natural images, or inner-product distances between the optimal stimulus and $n$ nearest natural images $\left\lbrace {\bf{x}}^{t} \right\rbrace$, \ie $\frac{1}{n}\sum_{i=1}^{n} \langle {\bf{x}}^{t}_{i} , \hat{{\bf{x}}} \rangle$ with $n=100$, to measures how well the optimal stimulus can linearly ``approximate'' natural images (or intuitively how much neuronal response might be elicited).
It can be observed that, though none of the natural images is highly similar to any of the optimal stimuli, when alternatively comparing the averages of natural images of the top 1\% explainablilities and the optimal stimuli, the similarities (Fig.~\ref{fig:exppow}A.1 \vs \ref{fig:exppow}A.2, \etc) and dissimilarities (Fig.~\ref{fig:exppow}D.1 \vs \ref{fig:exppow}D.2) become apparent.
Such kind of ``linearities'' around task-related images out of the well-performing highly-nonlinear networks were observed in previous studies \cite{le2012building, simonyan2013deep} too, where optimal stimuli also resembled averages or combinations of visual objects.

%For simplicity and clarity of presentation, the explanation power measure is calculated using the top 1\% (i.e.~100) of the task-related stimuli with highest explainabilities, without biasing the results compared to using all task-related stimuli.

{\bf Invariance and selectivity \vs natural image subspace alignment.} It has been shown that invariance and selectivity are extremely important for biological vision \cite{desimone1991face, ito1995size}; however, how these properties directly affect the recognition performances of artificial neural networks remains unclear.
To address this question, starting from 16 randomly sampled reference stimuli, invariance and selectivity path searches {were} performed at $\delta = 0.1\pi$ using Eq.~\ref{eq:I2}, \ref{eq:S2} for vector representations. 
Figure \ref{fig:align}B--E demonstrate the results of invariance and selectivity subspace alignment analyses, which measures the alignment between subspaces formed by natural images and invariance and selectivity paths (\ie how likely the invariance and selectivity can benefit recognition) via estimating the sparsity of $n$ paths projected onto the principal component vectors $\bf{V}$ of natural images $\left\lbrace {\bf{x}}^{t} \right\rbrace$, i.e.~$\frac{1}{n}\sum_{i=1}^{n} \left\| {\bf{Vx}}_{\delta,i} \right\|_{1}$ with $n=20$ and $\delta=0.1\pi$.

%quiroga2005invariant
%It can be observed that good alignments (\ie sparser representations in the principle component vector spaces) correspond to highly structural deformations or lighting changes, which are known to be important factors in visual recognition, while bad alignments correspond to mostly meaningless noises.

\newcommand{\expregular}{Equation \ref{eq:O2} though describes the essence of our approach, was not able to produce analyzable numerical results alone, due to the fact that deep representations can be highly invariant to noises that do not resemble natural images.
Implementation-wise, it was further regularized with the energy of Laplacian filtering $\left\| \mathcal{L} \left( \bf{x} \right) \right\|$ to enforce smoothness and similarity to natural images.
Similar concept was also adopted in \cite{mahendran2014understanding} independently.
}  %, with regularization constant decided via optimizing withheld data.

{\bf Encoding specificity of natural images.} To further understand the meaning of vector representation $\bf{r}$ formed by a deep network $f$, the inverse function $f^{-1}\left( \bf{r} \right)$ {was} numerically approximated to reveal what kind of stimulus can drive the deep network to give output $\bf{r}$. 
Instead of inverting randomly generated $\bf{r}$, which may not have feasible or interpretable solutions, we {used} $\bf{r}$ from known reference stimulus $\tilde{\bf{x}}$ to study the encoding specificity and its relationship to performance.
Figure \ref{fig:encspc} shows the reference stimuli, the best 3 and worst 1 examples of encoding specificities, where the optimal stimulus search (Eq.~\ref{eq:O2}) is utilized as an inverse function of representations\footnote{\expregular}, and the average of structural similarities (SSIM) \cite{wang2004image} between the task-related reference stimulus $\tilde{\bf{x}}$ and $n$ reconstructed stimuli from random initializations, \ie $\frac{1}{n}\sum_{i=1}^{n} \mathrm{SSIM}\left(\tilde{\bf{x}} , \hat{{\bf{x}}}_i \right)$ with $n=10$, is measured to indicate how specific (or non-confounding) a representation is encoded.
While similar studies \cite{mahendran2014understanding, long2014convnets, razavian2014persistent} also employed such kind of ``reconstructability'' to study the representations of deep networks, our results suggest that, this property can actually be required for a deep network to perform well.

%SSIM measured using default settings of the standard implementation \cite{wang2004image} while different settings also yield similar correlations; other similarity measures like SNR, PSNR, and inner-product distance yield similar correlations as well

%Encoding specificity, as a single representation measure, can best explain the deep network's performance. Though bearing strong similarity to the invariance path search in terms of mathematical formulations, encoding specificity is in fact a radically different and complementary measure, since its numerical optimization is not constrained through distances to the reference stimulus (while distance constraints can induce \emph{a priori} similarities); thus, it can better function as an unconstrained global characterization of the ``encoding landscape'' (i.e.~multivariate tuning landscape of a population) and estimate if the (reconstructed) stimuli that a deep network is invariant to are ``selective'' (i.e.~visually similar to original stimuli).

\newcommand{\defvecpp}{Since no baseline can be defined for vector representations likewise, the invariance (and selectivity, similarly) path potential is redefined as path integral on the fitness landscape $\int_{0}^{\frac{\pi}{2}}{\exp\left(-\left\|f\left(\bf{x}^{+}_{\delta}\right)-f\left(\hat{\bf{x}}\right)\right\|\right)}\mathrm{d}\delta \mathbin{/} {\frac{\pi}{2}}$.}

\newcommand{\expinsc}{Vector representation's INSC is negatively correlated to network's performance, mainly due to the fact that poorer performing network's invariance paths are usually noisier (as depicted in Fig.~\ref{fig:align}D), and thus should be interpreted differently from scalar representation's INSC.}

\newcommand{\expextrinsic}{As second-order properties, INPP, SLPP, and INSC are measures of search results \emph{started} from task-related images, but not \emph{ended} compared against them, unlike ITSA and STSA.
Task-unrelated measures, \ie OSSC, INPP, SLPP, and INSC of scalar representations as used in Fig.~\ref{fig:pair}, are significantly worse at explaining network's performance (altogether with linear multiple correlation of only $0.34$), as one would expect.
}

{\bf Overall comparison.} Spearman's correlations between all representation measures and 100 deep networks' performances are summarized in Fig.~\ref{fig:corr}, and 71\% of the variance of deep networks' performances are explainable with the proposed measures altogether using linear multiple correlation analysis.
These measures further break down the differences among deep representations and suggest features like network's explanation power, invariance and selectivity subspace properties, and encoding specificities against the task-related stimuli, can be extremely important for the effectiveness of representations.
Invariance and selectivity path potentials\footnote{\defvecpp} and invariance subspace capacity\footnote{\expinsc} for deep vector representations, on the other hand, have lower correlations to network's performance (though still significant), possibly due to their nature of being less task-related\footnote{\expextrinsic} than subspace alignment measures.
Worth to note, although the first-order properties (OSEP and TSES) have stronger individual correlations, they do not explain more about the network's performance than the other second-order properties, using linear multiple correlation analyses ($0.56$ \vs $0.58$).

%Network's invariance subspace capacity (in terms of population representation), which though has an $R^2=0.15$ correlation as well, is in fact negatively correlated to network's performance, mainly due to the fact that poor performing network's invariance path search results are usually noisier (as depicted in Fig.~\ref{fig:align}D), and thus should be interpreted differently from the unit representation cases. on the other hand, are weaker in explaining networks' performances (with $R^2=0.10/0.06$), due to the fact that all deep networks have high but less ranked invariance and selectivity.

\section{Discussion}

%how to intrepret results
{\bf Contributions of this work.} Comparatively, the characterization of representations in this work is designed to be general and unbiased such that both first- and second-order methods can be supported for single neurons or groups of neurons, without assuming any \emph{de facto} property in the solutions (\eg parametric deformations as used for studying invariance in previous works).
Additionally, with the strong statistical significances of results obtained from the large-scale experiments on 100 shallow and 100 deep networks, totally 6,400 artificial neurons, we argue that our observations toward answering the differences between (shallow \vs deep) and among (deep) representations are highly conclusive, which can be summarized as:
\begin{itemize} % Main Findings
\setlength\itemsep{0.0ex}
\item Complexity of representations quantitatively increases along with network's depth.
\item Unlike deep representation which is both invariant and selective, shallow representation is only invariant, not selective, and its capacity of invariance is significantly lower than deep representation's.
\item Both first-order properties of deep representations (optimal stimulus explanation power and encoding specificity of natural images) are crucially important to deep networks' performances.
\item All second-order properties of deep representations (in particular, invariance and selectivity \vs natural image subspace alignment) together are as important to deep networks' performances, if not significantly more.
\end{itemize}

{\bf Future works.} In this work, we only studied represents from shallow networks and the ``simplest'' type of deep networks, two-level networks.
Although our results can potentially generalize if viewed as the characterization of ``adding an extra level'', it is still highly desirable to extend the study into deeper networks.
While the methods and measures proposed in this work are neutral to the depth of networks, numerical difficulties (\eg evaluation and convergence speed, quality of solutions, \etc) may need to be addressed in very deep networks.
Beyond explaining what properties make a deep representation good, it would be valuable too to incorporate these findings in learning even better representations, which are almost solely driven by bigger datasets nowadays, via certain kind of regularizations.  
We argue that this could be useful in accelerating or improving representation learning, especially when data are scarce and representations are not easily transferable.

%Secondly, statistical tests were extensively deployed to support observations reported in this paper, especially in the context of real applications (\ie LFW-a face recognition), which might seem still far-fetched to be fully explainable by mathematical theories.
%Nonetheless, we envision that combining more theoretical analyses which, \eg, have been adopted in various topics in understanding deep representations \cite{delalleau2011shallow, montufar2014number}, with statistical simulations can further empower us in answering other important questions in deep networks \cite{bengio2013deep}.

%%application
% application: where training data is scarse; actively regularize representations
% not deep in this work, but adding another layer(level), what would happen
% not only optimal stimulus; modulate the landscape
% theory works gave some general answers, we want to answer, in real apps, why.
% still, to combine with theory

%\clearpage

{\small
\bibliographystyle{ieee}
\bibliography{ref}
}

\end{document}

### 2015-04-01 paper_for_review.tex
v1.1

@@
 % Include other packages here, before hyperref.
-\usepackage[labelfont=bf]{caption}
+\usepackage[labelfont=bf, skip=5pt]{caption}
 \captionsetup{labelsep=period}
+\usepackage{footmisc}
 
@@
 Using methods which support characterizing both first-order and second-order properties of scalar and vector representations, we conducted a large-scale experiment, with 200 randomly generated shallow and deep networks, or totally 6,400 neurons characterized, and the results analyzed through a series of measures designed to quantify the differences in representations.
-We conclude that deep representations are significantly better than shallow representations, and overall 71\% of the variance in deep networks' performances can be explained with our measures, both of which are confirmed using statistical tests for the first time.
+We conclude that deep representations are quantitatively better than shallow representations, and overall 71\% of the variance in deep networks' performances can be explained with our measures, both of which are confirmed using statistical tests for the first time.
 
@@
 \begin{center}
-\includegraphics[width=0.35\textwidth]{Figs/e_fig7s_compact-crop.pdf} 
+\includegraphics[width=0.40\textwidth]{Figs/e_fig7s_compact-crop.pdf} 
 \end{center}
@@
 In existing literatures, researchers mostly tried to come up with one (or few) ``best'' network architecture (\ie hyperparameter set, either through manual tuning or automated hyperparameter search) and then benchmarked the network's performances.
-However, as exemplified in the figure, when considering a larger population of networks \cite{cox2011beyond}, it is not fully clear to us why deep (\ie two-level) networks in practice outperform shallow (\ie one-level) networks, and why certain networks are better than the others, given their broad distribution of performances on LFW-a face recognition \cite{LFWTech, wolf2011effective}.
+However, as exemplified in the figure, when considering a larger population of networks \cite{cox2011beyond}, it is not fully clear to us why deep (\eg two-level) networks in practice outperform shallow (\eg one-level) networks, and why certain networks are better than the others, given their broad distribution of performances on LFW-a face recognition \cite{LFWTech, wolf2011effective}.
 %However, it is not fully clear that in practice (1) why deep networks are better than shallow networks, and (2) among networks of the same depth, why certain networks are better than the others.
@@
 As demonstrated in Fig.~\ref{fig:perf}, we summarized two important subquestions to be investigated: (1) why deep networks are better than shallow networks, and (2) among networks of the same depth, why certain networks are better than the others.
-Specifically, from the point of view that successful networks are forming ``better'' representations of visual inputs that are more separable to the final-stage linear classifiers, we want to identify the crucial properties that are required in good representations.
+Specifically, from the point of view that successful networks are forming ``better'' representations of visual inputs that are more separable to the final-stage linear classifiers, we want to identify the crucial properties that are required in forming (or learning) better representations.
+
+%which can be potentially useful in improve representation learning
 
@@
 
-Beyond the first-order characterization, the analyses of invariance and selectivity properties of neurons was adopted in a few works, which can be viewed as the second-order characterization of $f$, and parametric deformation in stimuli was the most commonly used setup.
+In addition to the first-order characterization, the analyses of invariance and selectivity properties of neurons (\eg changes in stimuli which given neurons are least and most sensitive to, how small or large the changes are, \etc) was adopted in a few works, which can be viewed as the second-order characterization of $f$, and parametric deformation in stimuli was the most commonly used setup.
 For example, translation, rotation, scaling, \etc in stimulus were used in \cite{goodfellow2009measuring, zeiler2014visualizing} to test the invariance and selectivity of neurons.
@@
 
-\newcommand{\expsettings}{In other words, 32 channels of filters in top levels' convolution layers.
+\newcommand{\expsettings}{In other words, 32 channels of filters in top levels' convolution layers, a simple setting with reasonable performances.
 Following the terminology in recent literatures, shallow and deep neurons corresponded to \texttt{pool1} and \texttt{pool2} layers respectively.
 Stimulus dimensionalities of the shallow and deep neurons {were} $N=121$ and $441$ (\ie spatially overlapping $11\times11$ and $21\times21$ receptive fields). % respectively
-Other minor changes included: nonlinear activations all simplified to \emph{ReLU} \cite{krizhevsky2012imagenet} mode and normalizations all conditionally subtractive mode.
-Overall, the architectures were more similar to those in \cite{simonyan2014very}, except that pooling operations can be \emph{average}, \emph{squared}, or \emph{max-like} in our case.
+Other minor changes included: nonlinear activations all simplified to \emph{ReLU} \cite{krizhevsky2012imagenet} mode and normalizations all subtractive mode.
+Overall, the architectures were more similar to those in \cite{simonyan2014very}, except pooling operations can be \emph{average}, \emph{squared}, or \emph{max-like} in our case.
 }
 
-In the experiments, 100 shallow (\ie one-level) networks and 100 deep (\ie two-level) networks {were} randomly generated (except all with 32 top-layer neurons\footnote{\expsettings}) and tested to see how their representations differ with networks' depths and affect networks' performances on face pair matching tasks against the LFW-a dataset \cite{LFWTech, wolf2011effective} as shown in Fig.~\ref{fig:perf}.
+\newcommand{\exptask}{We chose face images over object images for our recognition task because facial patches can be easily aligned and compared against our results.
+However, our methods and measures are not limited to just face images.}
+
+In the experiments, 100 shallow (\ie one-level) networks and 100 deep (\ie two-level) networks {were} randomly generated (except all with 32 top-layer neurons\footnote{\expsettings}) and tested to see how their representations differ with networks' depths and affect networks' performances on face pair matching\footnote{\exptask} tasks against the LFW-a dataset \cite{LFWTech, wolf2011effective} as shown in Fig.~\ref{fig:perf}.
 %, using 10-fold cross validation on view 2.
@@
 \caption{{\bf Fitness-distance analyses of shallow (A) and deep scalar representations (B).}
-Fitnesses of invariance (red dots, \ie $\left\lbrace f\left(\bf{x}^{+}_{\delta}\right) \right\rbrace$) and selectivity paths (blue dots, \ie $\left\lbrace f\left(\bf{x}^{-}_{\delta}\right) \right\rbrace$) of all 3,200 shallow and 3,200 deep neurons are shown for comparison.
-A small fraction of invariance paths in fact have fitnesses higher than the optimal stimuli's due to a non-convex problem's nature, which however does not cause noticeable differences in the resultant statistics.
-While both types of neurons show invariance, shallow neurons do not posses selectivity, since their selectivity paths overlap with the dashed line---the zero-selectivity baseline.
+Fitnesses of invariance (red dots, \ie $\left\lbrace f\left(\bf{x}^{+}_{\delta}\right) \right\rbrace$) and selectivity paths (blue dots, \ie $\left\lbrace f\left(\bf{x}^{-}_{\delta}\right) \right\rbrace$) of all 3,200 shallow and 3,200 deep neurons are shown for comparison\protect\footnotemark.
+Although both types of neurons show invariance, shallow neurons do not posses selectivity, since their selectivity paths overlap with the dashed line, the zero-selectivity baseline.
 In other words, \emph{only deep neurons are selective}.
@@
 
+\footnotetext{A small fraction of invariance paths in fact have fitnesses higher than the optimal stimuli's due to a non-convex problem's nature, which however does not cause noticeable differences in the resultant statistics.}
+
 \newcommand{\defbaseline}{Given a single inner-product neuron $f({\bf{x}}) = {\bf{w}}^{T}{\bf{x}}$ and $\left\| \bf{x} \right\| = 1$, we have $\hat{\bf{x}} = \bf{w}$ and thus $f({\bf{x}}_{\delta}) = {\hat{\bf{x}}}^{T}{\bf{x}}_{\delta} = \cos(\delta)$ by definition.}
@@
 
-\newcommand{\defstars}{In this paper, significance levels follow the definition that *, ** and *** correspond to $p<0.05$, $p<0.01$ and $p<0.001$ respectively.}
+\newcommand{\defstars}{In this paper, significance levels follow the definition that *, ** and *** correspond to $p<0.05$, $p<0.01$ and $p<0.001$ respectively.\label{fnote:defstars}}
 
@@
 Results from deep networks of the top 3 strongest average explanation power (A--C) and the weakest average explanation power (D) are visualized.
-Within each panel, the first row shows all 32 neurons' optimal stimuli from the network, the third to fifth rows show natural images with the top 3 strongest ``explainablilities'' (\ie inner-product distances between the images and the corresponding optimal stimuli), and the second row shows the average of images with top 100 explainablilities, out of 10,000 natural images.
-Color of the boarder of an image indicates the explainablility of the image, with color map following the definition in Fig.~\ref{fig:methods}.
+Within each panel, the first row shows all 32 neurons' optimal stimuli from the network, the third to fifth rows show natural images with the top 3 strongest ``explainabilities'' (\ie inner-product distances between the images and the corresponding optimal stimuli), and the second row shows the average of images with top 100 explainabilities, out of 10,000 natural images.
+Color of the boarder of an image indicates the explainability of the image, with color map following the definition in Fig.~\ref{fig:methods}.
 Comparing deep networks' explanation power and performances shows significant correlation (OSEP measure in Fig.~\ref{fig:corr}), which suggests \emph{``representing'' natural images can be required in first-order properties of good deep scalar representations (\ie optimal stimuli)}.
@@
 Representation measures following the raster order are, optimal stimulus explanation power (OSEP), invariance path potential (INPP), selectivity path potential (SLPP), invariance subspace capacity (INSC), invariance \vs task-related image subspace alignment (ITSA), selectivity \vs task-related image subspace alignment (STSA), task-related image encoding specificity (TSES), and all 7 measures together using linear multiple correlation analysis.
-Correlations $\rho$ and significance levels are $0.60^{\ast\ast\ast}$, $-0.31^{\ast\ast}$, $-0.24^{\ast}$, $-0.39^{\ast\ast\ast}$, $-0.44^{\ast\ast\ast}$, $-0.56^{\ast\ast\ast}$, $0.64^{\ast\ast\ast}$, and $0.84^{\ast\ast\ast}$, respectively.
+Correlations $\rho$ and significance levels\footref{fnote:defstars} are $0.60^{\ast\ast\ast}$, $-0.31^{\ast\ast}$, $-0.24^{\ast}$, $-0.39^{\ast\ast\ast}$, $-0.44^{\ast\ast\ast}$, $-0.56^{\ast\ast\ast}$, $0.64^{\ast\ast\ast}$, and $0.84^{\ast\ast\ast}$, respectively.
 Overall, \emph{71\% of the variance ($\rho^2$) of deep networks' performances can be explained} using the proposed representation measures altogether.
@@
 Figure \ref{fig:exppow} shows the results of best 3 and worst 1 networks in terms of explanation power, which is estimated as the mean of the linear explainability of natural images, or inner-product distances between the optimal stimulus and $n$ nearest natural images $\left\lbrace {\bf{x}}^{t} \right\rbrace$, \ie $\frac{1}{n}\sum_{i=1}^{n} \langle {\bf{x}}^{t}_{i} , \hat{{\bf{x}}} \rangle$ with $n=100$, to measures how well the optimal stimulus can linearly ``approximate'' natural images (or intuitively how much neuronal response might be elicited).
-It can be observed that, though none of the natural images is highly similar to any of the optimal stimuli, when alternatively comparing the averages of natural images of the top 1\% explainablilities and the optimal stimuli, the similarities (Fig.~\ref{fig:exppow}A.1 \vs \ref{fig:exppow}A.2, \etc) and dissimilarities (Fig.~\ref{fig:exppow}D.1 \vs \ref{fig:exppow}D.2) become apparent.
+It can be observed that, though none of the natural images is highly similar to any of the optimal stimuli, when alternatively comparing the averages of natural images of the top 1\% explainabilities and the optimal stimuli, the similarities (Fig.~\ref{fig:exppow}A.1 \vs \ref{fig:exppow}A.2, \etc) and dissimilarities (Fig.~\ref{fig:exppow}D.1 \vs \ref{fig:exppow}D.2) become apparent.
 Such kind of ``linearities'' around task-related images out of the well-performing highly-nonlinear networks were observed in previous studies \cite{le2012building, simonyan2013deep} too, where optimal stimuli also resembled averages or combinations of visual objects.
@@
 We argue that this could be useful in accelerating or improving representation learning, especially when data are scarce and representations are not easily transferable.
+Finally, even though a decently high fraction of the variance in deep networks' performances has been explained, how to further increase this number, if there is an upper bound (\eg limits from dataset itself), \etc, need to be further investigated.
 

### 2015-04-03 paper_for_review.tex
Corrections throughout paper.

@@
 %%%%%%%%% TITLE
-\title{What Makes Deep Network's Representations Good for Visual Recognition?}
+\title{What Makes Deep Network Representations Good for Visual Recognition?}
 
@@
 
-Representations of artificial neurons in deep neural networks have been demonstrated highly effective and robust in various visual recognition tasks \cite{krizhevsky2012imagenet, sermanet2013overfeat, donahue2014decaf}.
+Representations of artificial neurons in deep neural networks have been demonstrated to be highly effective and robust for various visual recognition tasks~\cite{krizhevsky2012imagenet, sermanet2013overfeat, donahue2014decaf}.
 Nevertheless, our understanding of them arguably remains limited.
 In this paper, we address two important and practical questions: (1) why deep networks are better than shallow networks, and (2) among networks of the same depth, why certain networks are better than the others.
-Using methods which support characterizing both first-order and second-order properties of scalar and vector representations, we conducted a large-scale experiment, with 200 randomly generated shallow and deep networks, or totally 6,400 neurons characterized, and the results analyzed through a series of measures designed to quantify the differences in representations.
-We conclude that deep representations are quantitatively better than shallow representations, and overall 71\% of the variance in deep networks' performances can be explained with our measures, both of which are confirmed using statistical tests for the first time.
+Using methods which support characterizing both first-order and second-order properties of scalar and vector representations, we considered the case study of face pair matching and conducted a large-scale experiment.
+This experiment examined 200 randomly generated shallow and deep networks, with the results analyzed through a series of measures designed to quantify the differences in representations.
+We conclude though this principled analysis that deep representations are quantitatively better than shallow representations, and overall 71\% of the variance in the performance of deep networks can be explained with our measures, both of which are confirmed using statistical tests for the first time.
 
@@
 \caption{{\bf What makes a deep network good?}
-In existing literatures, researchers mostly tried to come up with one (or few) ``best'' network architecture (\ie hyperparameter set, either through manual tuning or automated hyperparameter search) and then benchmarked the network's performances.
-However, as exemplified in the figure, when considering a larger population of networks \cite{cox2011beyond}, it is not fully clear to us why deep (\eg two-level) networks in practice outperform shallow (\eg one-level) networks, and why certain networks are better than the others, given their broad distribution of performances on LFW-a face recognition \cite{LFWTech, wolf2011effective}.
+In existing work, researchers have mostly tried to produce one ``best'' network architecture (\ie a hyperparameter set, either through manual tuning or automated hyperparameter search), which is subsequently evaluated on a particular benchmark.
+However, as exemplified in this figure, when considering a larger population of networks \cite{cox2011beyond}, it is not fully clear to us why in practice deep networks (\eg two or more levels) outperform shallow networks (\eg one-level), and why certain networks are better than the others, given their broad distribution of performance on a task like face pair matching for the LFW-a data set (shown above)~\cite{LFWTech, wolf2011effective}.
 %However, it is not fully clear that in practice (1) why deep networks are better than shallow networks, and (2) among networks of the same depth, why certain networks are better than the others.
@@
 
-Although deep neural networks have been successfully pushing the boundaries of computer vision by breaking the records of ImageNet Challenges (ILSVRC) year after year recently \cite{krizhevsky2012imagenet, sermanet2013overfeat, szegedy2014going}, our understanding about the properties of these networks arguably remains limited, since we could not yet fully answer the question ``what makes a deep network good?''.
-As demonstrated in Fig.~\ref{fig:perf}, we summarized two important subquestions to be investigated: (1) why deep networks are better than shallow networks, and (2) among networks of the same depth, why certain networks are better than the others.
-Specifically, from the point of view that successful networks are forming ``better'' representations of visual inputs that are more separable to the final-stage linear classifiers, we want to identify the crucial properties that are required in forming (or learning) better representations.
+Although deep neural networks have been successfully pushing the boundaries of computer vision by breaking the records of ImageNet Challenges (ILSVRC) year after year~\cite{krizhevsky2012imagenet, sermanet2013overfeat, szegedy2014going}, our understanding of the properties of these networks arguably remains limited, since we cannot yet fully answer the question ``what makes a deep network good?'' In this paper, we investigate two important facets of this question: (1) why deep networks are better than shallow networks, and (2) among networks of the same depth, why certain networks are better than others (Fig.~\ref{fig:perf}).
+Specifically, from the point of view that successful networks are forming ``better'' representations of visual inputs that are more separable to the final-stage linear classifiers, we want to identify the crucial properties that are required for forming (or learning) better representations.
 
@@
 %$f({\bf{x}}) = \left\| {\bf{K}} \otimes \ivec\left({\bf{x}}\right) \right\|_{F}$
-\newcommand{\expstimdim}{For simplicity of formulas, a stimulus being 2-dimensional (\ie ${\bf{x}} \in \mathbb{R}^{\sqrt{N} \times \sqrt{N}}$) or vectorized (\ie ${\bf{x}} \in \mathbb{R}^N$) are used interchangeably.}
+\newcommand{\expstimdim}{For simplicity, a stimulus being 2-dimensional (\ie ${\bf{x}} \in \mathbb{R}^{\sqrt{N} \times \sqrt{N}}$) or vectorized (\ie ${\bf{x}} \in \mathbb{R}^N$) are used interchangeably.}
 
 %first order, second order, flexibility
-{\bf Existing works.} 
-Given an artificial neural network $f$, the studying of how $f$ represents a high dimensional stimulus\footnote{\expstimdim} ${\bf{x}} \in \mathbb{R}^N$ as a representation 
-${\bf{r}} = f\left( {\bf{x}} \right)$ in existing literatures mainly focused on the first-order (\ie linear) characterization of $f$---the analysis of optimal stimuli, stimuli which lead to the strongest responses of given neurons.
-For single-level convolutional networks using squared pooling, \ie $f({\bf{x}}) = \left\| {\bf{K}} \otimes {\bf{x}} \right\|_{F}$, optimal stimulus was shown to be analytically approximable \cite{saxe2011random} using a Gabor-like filter, whether the convolution kernel $\bf{K}$ is structural or random.
-For multi-level convolutional networks, optimal stimuli of neurons from various depths were numerically approximated \cite{ngiam2010tiled, le2012building, zeiler2014visualizing, simonyan2013deep} to qualitatively show neurons in deeper layers are tuned to more complex visual patterns.
+{\bf Existing work.} 
+Given an artificial neural network $f$, the study of how $f$ represents a high dimensional stimulus\footnote{\expstimdim} ${\bf{x}} \in \mathbb{R}^N$ as a representation 
+${\bf{r}} = f\left( {\bf{x}} \right)$ in existing work mainly focused on the first-order (\ie linear) characterization of $f$ --- the analysis of optimal stimuli, stimuli which lead to the strongest responses of given neurons.
+For single-level convolutional networks using squared pooling, \ie $f({\bf{x}}) = \left\| {\bf{K}} \otimes {\bf{x}} \right\|_{F}$, the optimal stimulus was shown to be analytically approximable \cite{saxe2011random} using a Gabor-like filter, whether the convolution kernel $\bf{K}$ is structural or random.
+For multi-level convolutional networks, the optimal stimuli of neurons from various depths were numerically approximated \cite{ngiam2010tiled, le2012building, zeiler2014visualizing, simonyan2013deep} to qualitatively show neurons in deeper layers are tuned to more complex visual patterns.
 The generalized concept of optimal stimulus in considering groups of neurons can also be adopted to study the representations by inverting them \cite{mahendran2014understanding}.
 
-In addition to the first-order characterization, the analyses of invariance and selectivity properties of neurons (\eg changes in stimuli which given neurons are least and most sensitive to, how small or large the changes are, \etc) was adopted in a few works, which can be viewed as the second-order characterization of $f$, and parametric deformation in stimuli was the most commonly used setup.
-For example, translation, rotation, scaling, \etc in stimulus were used in \cite{goodfellow2009measuring, zeiler2014visualizing} to test the invariance and selectivity of neurons.
+In addition to the first-order characterization, the analysis of invariance and selectivity properties of neurons (\eg changes in stimuli which given neurons are least and most sensitive to, how small or large the changes are, \etc) was adopted in a few works.
+This can be viewed as the second-order characterization of $f$, and parametric deformation in stimuli was the most commonly used setup.
+For example, translation, rotation, scaling, \etc of the stimulus were used in~\cite{goodfellow2009measuring, zeiler2014visualizing} to test the invariance and selectivity of neurons.
 Similar deformations were also adopted in \cite{lenc2014understanding}, but extended to analyze more advanced properties (\ie equivariance and equivalence) of groups of neurons.
-More generalized setups that, \eg, utilize local quadratic information (\ie the Hessian) or ``global'' numerical searches in the stimulus space, even though do exist, were less explored in the context of artificial neural network studies.
+More generalized setups that, \eg, utilize local quadratic information (\ie the Hessian) or ``global'' numerical searches in the stimulus space, even though they do exist, are less explored in the context of artificial neural network studies.
 For quadratic networks, \ie $f\left({\bf{x}}\right) = \frac{1}{2}{\bf{x}}^{T}{\bf{Qx}}+{\bf{L}}^{T}{\bf{x}}+c$, through eigendecomposing the quadratic term $\bf{Q}$ (\ie the Hessian), local invariance and selectivity directions can be analytically derived \cite{berkes2006analysis}, which correspond to eigenvectors of the least and most negative eigenvalues.
-This method was also generalized onto multi-layer convolutional networks through numerically estimating the Hessian \cite{ngiam2010tiled}.
-For multi-layer networks, characterization of the ``global'' invariance of single neurons can also be achieved through numerically searching for non-local (\ie farther to the optimal stimuli) solutions \cite{erhan2010understanding}.
+This method was also generalized onto multi-layer convolutional networks by numerically estimating the Hessian \cite{ngiam2010tiled}.
+For multi-layer networks, characterization of the ``global'' invariance of single neurons can also be achieved by numerically searching for non-local (\ie farther to the optimal stimuli) solutions \cite{erhan2010understanding}.
 
@@
 
-\section{Methods}
+\section{Representation Characterization}
 
@@
 \caption{{\bf Representation characterization methods.}
-On the spherical (\ie constant energy) constraint of the $N$ dimensional stimulus space, similar to previous works, the first-order property (\ie optimal stimulus $\hat{\bf{x}}$, as shown in the figure's center) of a representation $\bf{r}$ is first iteratively searched (as the evolution of light gray eclipses shown in the figure) and analyzed.
-However, unlike most previous works, the second-order properties (\ie invariant path $\lbrace \bf{x}^{+}_{\delta} \rbrace$ as dashed red curve and selectivity path $\lbrace \bf{x}^{-}_{\delta} \rbrace$ as dashed blue curve) at distances $0.1\pi$ to $0.5\pi$ from the optimal stimulus are further searched and analyzed.
+On the spherical (\ie constant energy) constraint of the $N$ dimensional stimulus space, similar to previous work, the first-order property (\ie optimal stimulus $\hat{\bf{x}}$, as shown in the figure's center) of a representation $\bf{r}$ is first iteratively searched (as the evolution of light gray eclipses shown in the figure) and analyzed.
+However, unlike most previous work, the second-order properties (\ie invariant path $\lbrace \bf{x}^{+}_{\delta} \rbrace$ as dashed red curve and selectivity path $\lbrace \bf{x}^{-}_{\delta} \rbrace$ as dashed blue curve) at distances $0.1\pi$ to $0.5\pi$ from the optimal stimulus are also searched and analyzed.
 Additionally, these methods are capable of characterizing both scalar and vector representations, where the ``fitnesses'' to be optimized correspond to Eq.~\ref{eq:O1}, \ref{eq:I1}, \ref{eq:S1} and Eq.~\ref{eq:O2}, \ref{eq:I2}, \ref{eq:S2}, respectively.
@@
 
-To study the representations deep inside an artificial neural network, methods adopted in this work include: (1) first-order (\ie linear) characterization, or search and analysis of optimal stimuli, and (2) second-order (\ie quadratic) characterization, or search and analysis of invariance and selectivity paths (\ie eigenvectors of the largest and smallest eigenvalues), which are generalized and improved from previous works.
+To study the representations inside an artificial neural network, methods adopted in this work include: (1) first-order (\ie linear) characterization, or search and analysis of optimal stimuli, and (2) second-order (\ie quadratic) characterization, or search and analysis of invariance and selectivity paths (\ie eigenvectors of the largest and smallest eigenvalues), which are generalized and improved from previous work.
 $\bf{r}$ is defined as a scalar representation (\ie $\bf{r} \in \mathbb{R}$) when considering single neurons and a vector representation (\ie ${\bf{r}} \in \mathbb{R}^C$) when considering groups of neurons (with group size $C$).
 Our methods support characterizing both scalar (using Eq.~\ref{eq:O1}, \ref{eq:I1}, \ref{eq:S1}, though Eq.~\ref{eq:O2}, \ref{eq:I2}, \ref{eq:S2} can be used as well) and vector (using Eq.~\ref{eq:O2}, \ref{eq:I2}, \ref{eq:S2}) representations in a unified way.
-The search phases defined via these equations will be introduced in the following subsections, as the analyses of search results using representation measures will be covered in Sec.~\ref{sec:exp} when applied.
+The search phases defined via these equations will be introduced in the following subsections, and the analyses of search results using representation measures will be covered in Sec.~\ref{sec:exp}.
 
@@
 
-While no specific knowledge of the network itself is assumed, we do restrict the space of stimuli considered with a constant energy constraint $\left\| \bf{x} \right\| = E$, where $E$ is set to $1$ in all mathematical derivations for simplicity, but valued as the average energy of task-related images in our experiments.
+While no specific knowledge of the network itself is assumed, we do restrict the space of stimuli considered with a constant energy constraint $\left\| \bf{x} \right\| = E$, where $E$ is set to $1$ in all mathematical derivations for simplicity, but set to the average energy of task-related images in our experiments.
 Since neurons can often be modulated by stimulus contrasts, which is less interesting when considering pattern selectivities, limiting the stimulus space in this manner effectively avoids degenerate solutions that simply maximize stimulus contrasts, and reduces the range of possible stimuli to consider.
@@
 
-In this work, optimal stimulus is defined and numerically derived as 
+In this work, the optimal stimulus is defined and numerically derived as 
 \begin{align}
@@
 As visualized in Fig.~\ref{fig:methods}, this process involves an iterative search on the non-convex fitness landscape, which in fact can be handled efficiently by most of the modern numerical solvers, especially when analytical derivatives (\ie backpropagation gradients) are available.
-In this work, we adopted the CMA-ES solver \cite{hansen2001completely}, which though do not rely on analytical gradients and instead stochastically estimates numerical derivatives (\ie inverse Hessian), {worked} reasonably well in our experiments.
+In this work, we adopted the CMA-ES solver \cite{hansen2001completely}, which does not rely on analytical gradients and instead stochastically estimates numerical derivatives (\ie inverse Hessian), and worked reasonably well in our experiments.
 When considering a vector representation (using Eq.~\ref{eq:O2}), this process is equivalent to optimizing the response of an ``imaginary neuron'' tuned to a reference stimulus $\tilde{\bf{x}}$.
@@
 in which $0 < \delta \le \frac{\pi}{2}$, subject to $\left\| \bf{x}_{\delta} \right\| = 1$ and $\langle \bf{x}_{\delta} , \hat{\bf{x}} \rangle = \cos\left(\delta\right)$. 
-As visualized in Fig.~\ref{fig:methods}, the invariance path characterizes the $N$ dimensional curve that leads toward (or ideally, maintains at) fitness as high as possible while moving away from the optimal stimulus, and the selectivity path characterizes the curve that leads toward fitness as low as possible while moving away.
-The search process is implemented as multiple runs of maximization and minimization on discretized $\delta \in \left\lbrace 0.1\pi, 0.2\pi, 0.3\pi, 0.4\pi, 0.5\pi\right\rbrace$ as the circular distance constraints shown in Fig.~\ref{fig:methods}, where each run is initialized with the result from previous run (and the $0.1\pi$ run directly with optimal stimulus $\hat{\bf{x}}$) to increase the path continuity and searching speed\footnote{\expdiff}.
-In this work, the distance constraint $\delta$ only goes up to $\frac{\pi}{2}$, since on the $N$ dimensional sphere, stimuli fall in $\frac{\pi}{2} < \delta \le \pi$ are simply ``negatives'' of those in $0 < \delta \le \frac{\pi}{2}$, which are of less uniqueness and interest in general.
+As visualized in Fig.~\ref{fig:methods}, the invariance path characterizes the $N$ dimensional curve that leads toward (or ideally, maintains at) a fitness as high as possible while moving away from the optimal stimulus, and the selectivity path characterizes the curve that leads toward fitness as low as possible while moving away.
+The search process is implemented as multiple runs of maximization and minimization on discretized $\delta \in \left\lbrace 0.1\pi, 0.2\pi, 0.3\pi, 0.4\pi, 0.5\pi\right\rbrace$ as the circular distance constraints shown in Fig.~\ref{fig:methods}, where each run is initialized with the result from the previous run (and the $0.1\pi$ run directly with optimal stimulus $\hat{\bf{x}}$) to increase the path continuity and searching speed\footnote{\expdiff}.
+In this work, the distance constraint $\delta$ only goes up to $\frac{\pi}{2}$, since on the $N$ dimensional sphere, stimuli that fall in $\frac{\pi}{2} < \delta \le \pi$ are simply ``negatives'' of those in $0 < \delta \le \frac{\pi}{2}$, which are of less uniqueness and interest in general.
 The same process can also be performed with respect to a reference stimulus $\tilde{\bf{x}}$, especially for vector representations, where the invariance and selectivity of the ``imaginary neuron'' are again characterized equivalently.
@@
 For our experiments, a simplified variety of deep convolutional neural networks---the Spatio-Temporal Hierarchical Object Representation (STHOR) model \cite{pinto2009high, sthor}---{was} adopted.
-Although like other deep networks, it also consists of the standard cascade of convolution, nonlinear activation, pooling layers, \etc (which together define a single ``level'' in this work), its convolution kernel weights are however randomly assigned rather than trained.
-Though not matching the performances of deep networks trained on massive quantities of data \cite{krizhevsky2012imagenet}, it can in fact be surprisingly competitive for a variety of tasks, especially in the regime where large quantities of training samples are not easily available \cite{pinto2009high, cox2011beyond, viglarge}.
-Most importantly, since it does not require training (except for the linear SVM on top of the representations extracted), a large number of networks with diverse hyperparameters can be rapidly generated and compared, which is particularly useful in this study where relative, instead of absolute task performances, are of the main interest.
-As one might concern that comparing randomly weighted networks will not tell much about representations, since they might as well be random, our results also suggest otherwise.
+Like other deep networks, it consists of the standard cascade of convolution, nonlinear activation, pooling layers, \etc (which together define a single ``level'' in this work).
+However, its convolution kernel weights are  randomly assigned rather than trained.
+Though not matching the performances of deep networks trained on massive quantities of data \cite{krizhevsky2012imagenet}, it can in fact be surprisingly competitive for a variety of tasks, especially in the regime where large quantities of training samples are not readily available \cite{pinto2009high, cox2011beyond, viglarge}.
+Most importantly, since it does not require training (except for the linear SVM on top of the representations extracted), a large number of networks with diverse hyperparameters can be rapidly generated and compared, which is particularly useful in this study where relative, instead of absolute task performances, are of primary interest.
+As one might be concerned that comparing randomly weighted networks will not tell much about representations, since they might as well be random, our results will suggest otherwise.
 
-\newcommand{\expsettings}{In other words, 32 channels of filters in top levels' convolution layers, a simple setting with reasonable performances.
-Following the terminology in recent literatures, shallow and deep neurons corresponded to \texttt{pool1} and \texttt{pool2} layers respectively.
+\newcommand{\expsettings}{In other words, 32 channels of filters in the top level's convolution layers, a simple setting with reasonable performance.
+Following the terminology in recent work, shallow and deep neurons corresponded to \texttt{pool1} and \texttt{pool2} layers respectively.
 Stimulus dimensionalities of the shallow and deep neurons {were} $N=121$ and $441$ (\ie spatially overlapping $11\times11$ and $21\times21$ receptive fields). % respectively
-Other minor changes included: nonlinear activations all simplified to \emph{ReLU} \cite{krizhevsky2012imagenet} mode and normalizations all subtractive mode.
+Other minor changes included: nonlinear activations all simplified to \emph{ReLU} \cite{krizhevsky2012imagenet} mode and normalizations all in subtractive mode.
 Overall, the architectures were more similar to those in \cite{simonyan2014very}, except pooling operations can be \emph{average}, \emph{squared}, or \emph{max-like} in our case.
@@
 
-In the experiments, 100 shallow (\ie one-level) networks and 100 deep (\ie two-level) networks {were} randomly generated (except all with 32 top-layer neurons\footnote{\expsettings}) and tested to see how their representations differ with networks' depths and affect networks' performances on face pair matching\footnote{\exptask} tasks against the LFW-a dataset \cite{LFWTech, wolf2011effective} as shown in Fig.~\ref{fig:perf}.
+In the experiments, 100 shallow (\ie one-level) networks and 100 deep (\ie two-level) networks {were} randomly generated (though all have 32 top-layer neurons\footnote{\expsettings}) and tested to see how their representations differ with networks' depths and affect networks' performances on face pair matching\footnote{\exptask} tasks against the LFW-a dataset \cite{LFWTech, wolf2011effective} as shown in Fig.~\ref{fig:perf}.
 %, using 10-fold cross validation on view 2.
@@
 Second, the optimal stimuli of deep neurons are also relatively more complex (\ie consisting of more textural components) than those of shallow neurons (which are more uni-textured or Gabor-like).
-To quantify such differences, the spectral complexity measure, which is estimated through the $L^{1}$ norm of the (2 dimensional) Fourier power spectrum of optimal stimulus, \ie $\left\| \mathcal{F}(\hat{\bf{x}}) \right\|_{1}$, is adopted, where higher value suggests higher non-sparsity (\ie more spectral components required to represent the visual stimuli).
+To quantify such differences, the spectral complexity measure, which is estimated through the $L^{1}$ norm of the (2 dimensional) Fourier power spectrum of optimal stimulus, \ie $\left\| \mathcal{F}(\hat{\bf{x}}) \right\|_{1}$, is adopted, where a higher value suggests higher non-sparsity (\ie more spectral components required to represent the visual stimuli).
 
@@
 
-{\bf Invariance subspace capacity.} To further characterize the invariance properties, multiple runs of invariance path searches {were} performed and the results are visualized in Fig.~\ref{fig:allrep}D and \ref{fig:allrep}H.
+{\bf Invariance subspace capacity.} To further characterize the invariance properties, multiple runs of invariance path searches {were} performed and the results are visualized in Figs.~\ref{fig:allrep}D and \ref{fig:allrep}H.
 The diversity of the ``invariance plateau'' (\eg the central high fitness region in Fig.~\ref{fig:methods}) constitution of deep neurons is also more apparent than that of shallow neurons.
-Compared to invariance path potential which is designed to characterize the best invariance path even if only one of such exists, we further adopted the invariance subspace capacity measure to estimates the ``dimensionality'' (\ie how diverse different paths can be) of the linear subspace formed by multiple path search results via the nuclear norm of concatenation of $n$ results $\left\|\left[{\bf{x}}_{\delta,1},\dots,{\bf{x}}_{\delta,n}\right]\right\|_{*}$ with $n=20$ and $\delta=0.1\pi$.
+Compared to the invariance path potential, which is designed to characterize the best invariance path even if only one of such exists, we further adopted the invariance subspace capacity measure to estimates the ``dimensionality'' (\ie how diverse different paths can be) of the linear subspace formed by multiple path search results via the nuclear norm of concatenation of $n$ results $\left\|\left[{\bf{x}}_{\delta,1},\dots,{\bf{x}}_{\delta,n}\right]\right\|_{*}$ with $n=20$ and $\delta=0.1\pi$.
 
@@
 \end{center}
-\caption{{\bf Representativities (\ie explanation power) of optimal stimuli of deep neurons.}
-Results from deep networks of the top 3 strongest average explanation power (A--C) and the weakest average explanation power (D) are visualized.
+\caption{{\bf Explanatory power of optimal stimuli of deep neurons.}
+Results from deep networks of the top 3 strongest average explanatory power (A--C) and the weakest average explanatory power (D) are visualized.
 Within each panel, the first row shows all 32 neurons' optimal stimuli from the network, the third to fifth rows show natural images with the top 3 strongest ``explainabilities'' (\ie inner-product distances between the images and the corresponding optimal stimuli), and the second row shows the average of images with top 100 explainabilities, out of 10,000 natural images.
 Color of the boarder of an image indicates the explainability of the image, with color map following the definition in Fig.~\ref{fig:methods}.
-Comparing deep networks' explanation power and performances shows significant correlation (OSEP measure in Fig.~\ref{fig:corr}), which suggests \emph{``representing'' natural images can be required in first-order properties of good deep scalar representations (\ie optimal stimuli)}.
+Comparing deep networks' explanatory power and performances shows significant correlation (OSEP measure in Fig.~\ref{fig:corr}), which suggests \emph{``representing'' natural images can be required in first-order properties of good deep scalar representations (\ie optimal stimuli)}.
 }
@@
 As good alignments (B.1 and C.1) appear more natural (like structural deformations or lighting changes), bad alignments are mostly meaningless noises.
-Comparing deep networks' subspace alignment measures and performances shows significant correlations (ITSA/STSA measures in Fig.~\ref{fig:corr}), which suggest \emph{resembling natural image distributions can be required in second-order properties of good deep vector representations (\ie invariance and selectivity subspaces)}.
+Comparing deep networks' subspace alignment measures and performances shows significant correlations (ITSA/STSA measures in Fig.~\ref{fig:corr}), which suggest \emph{a resemblance to natural image distributions can be required in second-order properties of good deep vector representations (\ie invariance and selectivity subspaces)}.
 } %of subspaces % which are known to be important factors in visual recognition % accordingly
@@
 \caption{{\bf Spearman's correlations between deep network's representation measures and performance.}
-Representation measures following the raster order are, optimal stimulus explanation power (OSEP), invariance path potential (INPP), selectivity path potential (SLPP), invariance subspace capacity (INSC), invariance \vs task-related image subspace alignment (ITSA), selectivity \vs task-related image subspace alignment (STSA), task-related image encoding specificity (TSES), and all 7 measures together using linear multiple correlation analysis.
+Representation measures following the raster order are, optimal stimulus explanatory power (OSEP), invariance path potential (INPP), selectivity path potential (SLPP), invariance subspace capacity (INSC), invariance \vs task-related image subspace alignment (ITSA), selectivity \vs task-related image subspace alignment (STSA), task-related image encoding specificity (TSES), and all 7 measures together using linear multiple correlation analysis.
 Correlations $\rho$ and significance levels\footref{fnote:defstars} are $0.60^{\ast\ast\ast}$, $-0.31^{\ast\ast}$, $-0.24^{\ast}$, $-0.39^{\ast\ast\ast}$, $-0.44^{\ast\ast\ast}$, $-0.56^{\ast\ast\ast}$, $0.64^{\ast\ast\ast}$, and $0.84^{\ast\ast\ast}$, respectively.
@@
 
-{\bf Optimal stimulus explanation power.} To understand how optimal stimulus may affect a network's performance, we {tested} how well the optimal stimulus may ``linearly explain'' natural images and how the explanation power correlates with network's performance.
-Totally 10,000 image patches {were} randomly cropped out of the predefined facial regions of pictures from the LFW-a dataset \cite{wolf2011effective} and compared against the optimal stimuli of all 32 top-layer neurons from all 100 deep networks.
-Figure \ref{fig:exppow} shows the results of best 3 and worst 1 networks in terms of explanation power, which is estimated as the mean of the linear explainability of natural images, or inner-product distances between the optimal stimulus and $n$ nearest natural images $\left\lbrace {\bf{x}}^{t} \right\rbrace$, \ie $\frac{1}{n}\sum_{i=1}^{n} \langle {\bf{x}}^{t}_{i} , \hat{{\bf{x}}} \rangle$ with $n=100$, to measures how well the optimal stimulus can linearly ``approximate'' natural images (or intuitively how much neuronal response might be elicited).
+{\bf Optimal stimulus explanatory power.} To understand how the optimal stimulus may affect a network's performance, we {tested} how well the optimal stimulus may ``linearly explain'' natural images and how the explanatory power correlates with a network's performance.
+In total, 10,000 image patches {were} randomly cropped out of the predefined facial regions of pictures from the LFW-a dataset \cite{wolf2011effective} and compared against the optimal stimuli of all 32 top-layer neurons from all 100 deep networks.
+Figure \ref{fig:exppow} shows the results of best 3 and worst 1 networks in terms of explanatory power, which is estimated as the mean of the linear explainability of natural images, or inner-product distances between the optimal stimulus and $n$ nearest natural images $\left\lbrace {\bf{x}}^{t} \right\rbrace$, \ie $\frac{1}{n}\sum_{i=1}^{n} \langle {\bf{x}}^{t}_{i} , \hat{{\bf{x}}} \rangle$ with $n=100$, to measures how well the optimal stimulus can linearly ``approximate'' natural images (or intuitively how much neuronal response might be elicited).
 It can be observed that, though none of the natural images is highly similar to any of the optimal stimuli, when alternatively comparing the averages of natural images of the top 1\% explainabilities and the optimal stimuli, the similarities (Fig.~\ref{fig:exppow}A.1 \vs \ref{fig:exppow}A.2, \etc) and dissimilarities (Fig.~\ref{fig:exppow}D.1 \vs \ref{fig:exppow}D.2) become apparent.
@@
 
-{\bf Invariance and selectivity \vs natural image subspace alignment.} It has been shown that invariance and selectivity are extremely important for biological vision \cite{desimone1991face, ito1995size}; however, how these properties directly affect the recognition performances of artificial neural networks remains unclear.
+{\bf Invariance and selectivity \vs natural image subspace alignment.} It has been shown that invariance and selectivity are extremely important for biological vision \cite{desimone1991face, ito1995size}; however, how these properties directly affect the recognition performance of artificial neural networks remains unclear.
 To address this question, starting from 16 randomly sampled reference stimuli, invariance and selectivity path searches {were} performed at $\delta = 0.1\pi$ using Eq.~\ref{eq:I2}, \ref{eq:S2} for vector representations.
@@
 
-\newcommand{\expregular}{Equation \ref{eq:O2} though describes the essence of our approach, was not able to produce analyzable numerical results alone, due to the fact that deep representations can be highly invariant to noises that do not resemble natural images.
+\newcommand{\expregular}{While Equation \ref{eq:O2} describes the essence of our approach, it was not able to produce analyzable numerical results alone, due to the fact that deep representations can be highly invariant to noises that do not resemble natural images.
 Implementation-wise, it was further regularized with the energy of Laplacian filtering $\left\| \mathcal{L} \left( \bf{x} \right) \right\|$ to enforce smoothness and similarity to natural images.
-Similar concept was also adopted in \cite{mahendran2014understanding} independently.
+A similar concept was also adopted in~\cite{mahendran2014understanding}.
 }  %, with regularization constant decided via optimizing withheld data.
 
-{\bf Encoding specificity of natural images.} To further understand the meaning of vector representation $\bf{r}$ formed by a deep network $f$, the inverse function $f^{-1}\left( \bf{r} \right)$ {was} numerically approximated to reveal what kind of stimulus can drive the deep network to give output $\bf{r}$.
-Instead of inverting randomly generated $\bf{r}$, which may not have feasible or interpretable solutions, we {used} $\bf{r}$ from known reference stimulus $\tilde{\bf{x}}$ to study the encoding specificity and its relationship to performance.
+{\bf Encoding specificity of natural images.} To further understand the meaning of a vector representation $\bf{r}$ formed by a deep network $f$, the inverse function $f^{-1}\left( \bf{r} \right)$ {was} numerically approximated to reveal what kind of stimulus can drive the deep network to give output $\bf{r}$.
+Instead of inverting a randomly generated $\bf{r}$, which may not have feasible or interpretable solutions, we {used} $\bf{r}$ from a known reference stimulus $\tilde{\bf{x}}$ to study the encoding specificity and its relationship to performance.
 Figure \ref{fig:encspc} shows the reference stimuli, the best 3 and worst 1 examples of encoding specificities, where the optimal stimulus search (Eq.~\ref{eq:O2}) is utilized as an inverse function of representations\footnote{\expregular}, and the average of structural similarities (SSIM) \cite{wang2004image} between the task-related reference stimulus $\tilde{\bf{x}}$ and $n$ reconstructed stimuli from random initializations, \ie $\frac{1}{n}\sum_{i=1}^{n} \mathrm{SSIM}\left(\tilde{\bf{x}} , \hat{{\bf{x}}}_i \right)$ with $n=10$, is measured to indicate how specific (or non-confounding) a representation is encoded.
@@
 \newcommand{\expextrinsic}{As second-order properties, INPP, SLPP, and INSC are measures of search results \emph{started} from task-related images, but not \emph{ended} compared against them, unlike ITSA and STSA.
-Task-unrelated measures, \ie OSSC, INPP, SLPP, and INSC of scalar representations as used in Fig.~\ref{fig:pair}, are significantly worse at explaining network's performance (altogether with linear multiple correlation of only $0.34$), as one would expect.
+Task-unrelated measures, \ie OSSC, INPP, SLPP, and INSC of scalar representations as used in Fig.~\ref{fig:pair}, are significantly worse at explaining a network's performance (altogether with linear multiple correlation of only $0.34$), as one would expect.
 }
@@
 {\bf Overall comparison.} Spearman's correlations between all representation measures and 100 deep networks' performances are summarized in Fig.~\ref{fig:corr}, and 71\% of the variance of deep networks' performances are explainable with the proposed measures altogether using linear multiple correlation analysis.
-These measures further break down the differences among deep representations and suggest features like network's explanation power, invariance and selectivity subspace properties, and encoding specificities against the task-related stimuli, can be extremely important for the effectiveness of representations.
+These measures further break down the differences among deep representations and suggest features like network's explanatory power, invariance and selectivity subspace properties, and encoding specificities against the task-related stimuli, can be extremely important for the effectiveness of representations.
 Invariance and selectivity path potentials\footnote{\defvecpp} and invariance subspace capacity\footnote{\expinsc} for deep vector representations, on the other hand, have lower correlations to network's performance (though still significant), possibly due to their nature of being less task-related\footnote{\expextrinsic} than subspace alignment measures.
@@
 {\bf Contributions of this work.} Comparatively, the characterization of representations in this work is designed to be general and unbiased such that both first- and second-order methods can be supported for single neurons or groups of neurons, without assuming any \emph{de facto} property in the solutions (\eg parametric deformations as used for studying invariance in previous works).
-Additionally, with the strong statistical significances of results obtained from the large-scale experiments on 100 shallow and 100 deep networks, totally 6,400 artificial neurons, we argue that our observations toward answering the differences between (shallow \vs deep) and among (deep) representations are highly conclusive, which can be summarized as:
+Additionally, with the strong statistical significances of results obtained from the large-scale experiments on 100 shallow and 100 deep networks, totalling 6,400 artificial neurons, we argue that our observations toward answering the differences between (shallow \vs deep) and among (deep) representations are highly conclusive, which can be summarized as:
 \begin{itemize} % Main Findings
@@
 \item Complexity of representations quantitatively increases along with network's depth.
-\item Unlike deep representation which is both invariant and selective, shallow representation is only invariant, not selective, and its capacity of invariance is significantly lower than deep representation's.
-\item Both first-order properties of deep representations (optimal stimulus explanation power and encoding specificity of natural images) are crucially important to deep networks' performances.
-\item All second-order properties of deep representations (in particular, invariance and selectivity \vs natural image subspace alignment) together are as important to deep networks' performances, if not significantly more.
+\item Unlike a deep representation which is both invariant and selective, shallow representation is only invariant, not selective, and its capacity of invariance is significantly lower than a deep representation's.
+\item Both first-order properties of deep representations (optimal stimulus explanatory power and encoding specificity of natural images) are crucially important to deep network performance.
+\item All second-order properties of deep representations (in particular, invariance and selectivity \vs natural image subspace alignment) together are as important as first-order properties, if not significantly more.
 \end{itemize}
 
-{\bf Future works.} In this work, we only studied represents from shallow networks and the ``simplest'' type of deep networks, two-level networks.
+{\bf Future work.} In this work, we only studied shallow networks and the ``simplest'' type of deep networks, two-level networks.
 Although our results can potentially generalize if viewed as the characterization of ``adding an extra level'', it is still highly desirable to extend the study into deeper networks.

### 2015-04-06 paper_for_review.tex
v1.2

@@
 % Include other packages here, before hyperref.
-\usepackage[labelfont=bf, skip=5pt]{caption}
+\usepackage[labelfont=bf, skip=3pt]{caption}
 \captionsetup{labelsep=period}
@@
 This experiment examined 200 randomly generated shallow and deep networks, with the results analyzed through a series of measures designed to quantify the differences in representations.
-We conclude though this principled analysis that deep representations are quantitatively better than shallow representations, and overall 71\% of the variance in the performance of deep networks can be explained with our measures, both of which are confirmed using statistical tests for the first time.
+We conclude through this principled analysis that deep representations are quantitatively better than shallow representations, and overall 71\% of the variance in the performance of deep networks can be explained with our measures, both of which are confirmed using statistical tests for the first time.
 
@@
 \begin{center}
-\includegraphics[width=0.40\textwidth]{Figs/e_fig7s_compact-crop.pdf} 
+\includegraphics[width=0.90\columnwidth]{Figs/e_fig7s_compact-crop.pdf} 
 \end{center}
@@
 Given an artificial neural network $f$, the study of how $f$ represents a high dimensional stimulus\footnote{\expstimdim} ${\bf{x}} \in \mathbb{R}^N$ as a representation 
-${\bf{r}} = f\left( {\bf{x}} \right)$ in existing work mainly focused on the first-order (\ie linear) characterization of $f$ --- the analysis of optimal stimuli, stimuli which lead to the strongest responses of given neurons.
+${\bf{r}} = f\left( {\bf{x}} \right)$ in existing work mainly focused on the first-order (\ie linear) characterization of $f$---the analysis of optimal stimuli, stimuli which lead to the strongest responses of given neurons.
 For single-level convolutional networks using squared pooling, \ie $f({\bf{x}}) = \left\| {\bf{K}} \otimes {\bf{x}} \right\|_{F}$, the optimal stimulus was shown to be analytically approximable \cite{saxe2011random} using a Gabor-like filter, whether the convolution kernel $\bf{K}$ is structural or random.
@@
 \begin{center}
-\includegraphics[width=0.35\textwidth]{Figs/methods.pdf} 
+\includegraphics[width=1.0\columnwidth]{Figs/methods.jpg} 
 \end{center}
 \caption{{\bf Representation characterization methods.}
-On the spherical (\ie constant energy) constraint of the $N$ dimensional stimulus space, similar to previous work, the first-order property (\ie optimal stimulus $\hat{\bf{x}}$, as shown in the figure's center) of a representation $\bf{r}$ is first iteratively searched (as the evolution of light gray eclipses shown in the figure) and analyzed.
-However, unlike most previous work, the second-order properties (\ie invariant path $\lbrace \bf{x}^{+}_{\delta} \rbrace$ as dashed red curve and selectivity path $\lbrace \bf{x}^{-}_{\delta} \rbrace$ as dashed blue curve) at distances $0.1\pi$ to $0.5\pi$ from the optimal stimulus are also searched and analyzed.
+On the spherical (\ie constant energy) constraint of the $N$ dimensional stimulus space, similar to previous work, the first-order property (\ie optimal stimulus $\hat{\bf{x}}$) of a representation $\bf{r}$ is first iteratively searched (as the evolution of light gray eclipses) and analyzed.
+However, unlike most previous work, the second-order properties (\ie invariant path $\lbrace \bf{x}^{+}_{\delta} \rbrace$ and selectivity path $\lbrace \bf{x}^{-}_{\delta} \rbrace$) at distances $0.1\pi$ to $0.5\pi$ from the optimal stimulus are also searched and analyzed.
 Additionally, these methods are capable of characterizing both scalar and vector representations, where the ``fitnesses'' to be optimized correspond to Eq.~\ref{eq:O1}, \ref{eq:I1}, \ref{eq:S1} and Eq.~\ref{eq:O2}, \ref{eq:I2}, \ref{eq:S2}, respectively.
@@
 
-\newcommand{\expsettings}{In other words, 32 channels of filters in the top level's convolution layers, a simple setting with reasonable performance.
+\newcommand{\expsettings}{\Ie 32 channels of filters in the top level's convolution layers, a simple setting with reasonable performance.
 Following the terminology in recent work, shallow and deep neurons corresponded to \texttt{pool1} and \texttt{pool2} layers respectively.
@@
 \begin{center}
-\includegraphics[width=0.70\textwidth]{Figs/pic1.pdf} 
+\includegraphics[width=0.80\textwidth]{Figs/pic1.jpg} 
 \end{center}
@@
 \begin{center}
-\includegraphics[width=0.35\textwidth]{Figs/e_fig3a_2-crop-nup-crop.pdf} 
+\includegraphics[width=0.75\columnwidth]{Figs/fig4.pdf} 
 \end{center}
@@
 Fitnesses of invariance (red dots, \ie $\left\lbrace f\left(\bf{x}^{+}_{\delta}\right) \right\rbrace$) and selectivity paths (blue dots, \ie $\left\lbrace f\left(\bf{x}^{-}_{\delta}\right) \right\rbrace$) of all 3,200 shallow and 3,200 deep neurons are shown for comparison\protect\footnotemark.
-Although both types of neurons show invariance, shallow neurons do not posses selectivity, since their selectivity paths overlap with the dashed line, the zero-selectivity baseline.
+Although both types of neurons show invariance, shallow neurons do not posses selectivity, since their selectivity paths overlap with the zero-selectivity baseline.
 In other words, \emph{only deep neurons are selective}.
 } %, and random walks (green dots, \ie $\left\lbrace f\left({\bf{x}}^{R}_{\delta}\right) \right\rbrace$)
+%the dashed line, 
 \label{fig:fda}
@@
 \begin{center}
-\includegraphics[width=0.35\textwidth]{Figs/e_fig23b-crop.pdf} 
+\includegraphics[width=0.80\columnwidth]{Figs/fig5.pdf} 
 \end{center}
 \caption{{\bf Comparison of shallow and deep scalar representation measures.}
-Representation measures from left to right are, optimal stimulus spectral complexity (OSSC), invariance path potential (INPP), selectivity path potential (SLPP), and invariance subspace capacity (INSC), in which data points from shallow networks (left distributions) and deep networks (right distributions) are shown side by side.
-Sensitivity indexes ($d'$) between distributions are $1.63$, $0.42$, $4.73$, and $3.67$, respectively.
-}
+Representation measures from left to right are, optimal stimulus spectral complexity (OSSC), invariance path potential (INPP), selectivity path potential (SLPP), and invariance subspace capacity (INSC).
+Sensitivity indexes ($d'$) between distributions are $1.63$, $0.42$, $4.73$, and $3.67$, and differences between means of distributions are all significant (***), which suggest \emph{deep representations are quantitatively and significantly better than shallow representations}.
+} %, in which data points from shallow networks (left distributions) and deep networks (right distributions) are shown side by side
 \label{fig:pair}
@@
 {\bf Overall comparison.} Comparisons of spectral complexities, invariance/selectivity path potentials, and invariance subspace capacities of shallow and deep neurons (as scalar representations) are summarized in Fig.~\ref{fig:pair}.
-These measures further reveal the significant\footnote{\defstars} differences between shallow and deep representations, in addition to the separability and classifiability used in previous studies \cite{donahue2014decaf, zeiler2014visualizing}.
+These measures further quantify the significant\footnote{\defstars} differences between shallow and deep representations, in addition to the separability and classifiability used in previous studies \cite{donahue2014decaf, zeiler2014visualizing}.
 For instance, subtle differences of visual features have better chances being distinguished when the ``gap'' between invariance and selectivity curves (\ie the dynamic range of neuronal responses, or amplification ratio of differences) is large.
@@
 \begin{center}
-\includegraphics[width=0.70\textwidth]{Figs/pic2.pdf} 
+\includegraphics[width=0.85\textwidth]{Figs/pic2.jpg} 
 \end{center}
@@
 \begin{center}
-\includegraphics[width=0.35\textwidth]{Figs/pic3.pdf} 
+\includegraphics[width=\columnwidth, trim=1em 0 1em 0]{Figs/pic3.jpg} 
 \end{center}
@@
 \begin{center}
-\includegraphics[width=0.35\textwidth]{Figs/pic4.pdf} 
+\includegraphics[width=\columnwidth, trim=1em 0 1em 0]{Figs/pic4.jpg} 
 \end{center}

### 2015-04-07 paper_for_review.tex
v1.3

@@
 % Include other packages here, before hyperref.
-\usepackage[labelfont=bf, skip=3pt]{caption}
+%\usepackage{float}
+\usepackage[font=small, labelfont=bf, skip=3pt]{caption} %
 \captionsetup{labelsep=period}
@@
 
-Representations of artificial neurons in deep neural networks have been demonstrated to be highly effective and robust for various visual recognition tasks~\cite{krizhevsky2012imagenet, sermanet2013overfeat, donahue2014decaf}.
+Representations of artificial neurons in deep neural networks have been demonstrated to be highly effective and robust for various visual recognition tasks \cite{krizhevsky2012imagenet, sermanet2013overfeat, donahue2014decaf}.
 Nevertheless, our understanding of them arguably remains limited.
 In this paper, we address two important and practical questions: (1) why deep networks are better than shallow networks, and (2) among networks of the same depth, why certain networks are better than the others.
-Using methods which support characterizing both first-order and second-order properties of scalar and vector representations, we considered the case study of face pair matching and conducted a large-scale experiment.
-This experiment examined 200 randomly generated shallow and deep networks, with the results analyzed through a series of measures designed to quantify the differences in representations.
+Using methods which support characterizing both first- and second-order properties of scalar and vector representations, we considered the case study of face pair matching and conducted a large-scale experiment.
+The experiment examined 200 randomly generated shallow and deep networks, with the results analyzed via a series of measures designed to quantify the differences in representations. 
 We conclude through this principled analysis that deep representations are quantitatively better than shallow representations, and overall 71\% of the variance in the performance of deep networks can be explained with our measures, both of which are confirmed using statistical tests for the first time.
@@
 %concern about deep networks; how alleviated via this work
-
 \end{abstract}
@@
 \begin{center}
-\includegraphics[width=0.90\columnwidth]{Figs/e_fig7s_compact-crop.pdf} 
+\includegraphics[width=0.85\columnwidth]{Figs/e_fig7s_compact-crop.pdf} 
 \end{center}
 \caption{{\bf What makes a deep network good?}
-In existing work, researchers have mostly tried to produce one ``best'' network architecture (\ie a hyperparameter set, either through manual tuning or automated hyperparameter search), which is subsequently evaluated on a particular benchmark.
-However, as exemplified in this figure, when considering a larger population of networks \cite{cox2011beyond}, it is not fully clear to us why in practice deep networks (\eg two or more levels) outperform shallow networks (\eg one-level), and why certain networks are better than the others, given their broad distribution of performance on a task like face pair matching for the LFW-a data set (shown above)~\cite{LFWTech, wolf2011effective}.
+In existing work, researchers have mostly tried to produce one ``best'' network architecture (\ie a hyperparameter set, through manual tuning or automated hyperparameter search), which is subsequently evaluated on a particular benchmark.
+However, when considering a larger population of networks \cite{cox2011beyond}, it is not fully clear why in practice deep networks (\eg two or more levels) outperform shallow networks (\eg one level), and why certain networks are better than the others, given their broad distribution of performance on a task like face pair matching for the LFW-a dataset \cite{LFWTech, wolf2011effective} as shown above.
 %However, it is not fully clear that in practice (1) why deep networks are better than shallow networks, and (2) among networks of the same depth, why certain networks are better than the others.
 %As exemplified in the figure, deep (\ie two-level) networks not only on average outperform shallow (\ie one-level) networks \cite{cox2011beyond}, but also exhibit a wide range of performance differences on LFW-a face recognition \cite{LFWTech, wolf2011effective}.
+%as exemplified in this figure,
 } % (\eg optimal stimulus). %and (sometimes) characterized its properties
@@
 
-Although deep neural networks have been successfully pushing the boundaries of computer vision by breaking the records of ImageNet Challenges (ILSVRC) year after year~\cite{krizhevsky2012imagenet, sermanet2013overfeat, szegedy2014going}, our understanding of the properties of these networks arguably remains limited, since we cannot yet fully answer the question ``what makes a deep network good?'' In this paper, we investigate two important facets of this question: (1) why deep networks are better than shallow networks, and (2) among networks of the same depth, why certain networks are better than others (Fig.~\ref{fig:perf}).
+Although deep neural networks have been successfully pushing the boundaries of computer vision by breaking the records of ImageNet Challenges (ILSVRC) \cite{russakovsky} year after year \cite{krizhevsky2012imagenet, sermanet2013overfeat, szegedy2014going}, our understanding of the properties of these networks arguably remains limited, since we cannot yet fully answer the question ``what makes a deep network good?'' In this paper, we investigate two important facets of this question: (1) why deep networks are better than shallow networks, and (2) among networks of the same depth, why certain networks are better than others (Fig.~\ref{fig:perf}).
 Specifically, from the point of view that successful networks are forming ``better'' representations of visual inputs that are more separable to the final-stage linear classifiers, we want to identify the crucial properties that are required for forming (or learning) better representations.
@@
 {\bf Existing work.} 
-Given an artificial neural network $f$, the study of how $f$ represents a high dimensional stimulus\footnote{\expstimdim} ${\bf{x}} \in \mathbb{R}^N$ as a representation 
+When given an artificial neural network $f$, the study of how $f$ represents a high dimensional stimulus\footnote{\expstimdim} ${\bf{x}} \in \mathbb{R}^N$ as a representation 
 ${\bf{r}} = f\left( {\bf{x}} \right)$ in existing work mainly focused on the first-order (\ie linear) characterization of $f$---the analysis of optimal stimuli, stimuli which lead to the strongest responses of given neurons.
@@
 This can be viewed as the second-order characterization of $f$, and parametric deformation in stimuli was the most commonly used setup.
-For example, translation, rotation, scaling, \etc of the stimulus were used in~\cite{goodfellow2009measuring, zeiler2014visualizing} to test the invariance and selectivity of neurons.
+For example, translation, rotation, scaling, \etc of the stimulus were used in \cite{goodfellow2009measuring, zeiler2014visualizing} to test the invariance and selectivity of neurons.
 Similar deformations were also adopted in \cite{lenc2014understanding}, but extended to analyze more advanced properties (\ie equivariance and equivalence) of groups of neurons.
@@
 \begin{center}
-\includegraphics[width=1.0\columnwidth]{Figs/methods.jpg} 
+\includegraphics[width=0.90\columnwidth]{Figs/methods.jpg} 
 \end{center}
-\caption{{\bf Representation characterization methods.}
-On the spherical (\ie constant energy) constraint of the $N$ dimensional stimulus space, similar to previous work, the first-order property (\ie optimal stimulus $\hat{\bf{x}}$) of a representation $\bf{r}$ is first iteratively searched (as the evolution of light gray eclipses) and analyzed.
+\caption{{\bf Methods of representation characterization.}
+On the spherical (\ie constant energy) constraint of the $N$ dimensional stimulus space, similar to previous work, the first-order property (\ie optimal stimulus $\hat{\bf{x}}$) of a representation $\bf{r}$ is first iteratively searched (as the evolution of gray eclipses) and analyzed.
 However, unlike most previous work, the second-order properties (\ie invariant path $\lbrace \bf{x}^{+}_{\delta} \rbrace$ and selectivity path $\lbrace \bf{x}^{-}_{\delta} \rbrace$) at distances $0.1\pi$ to $0.5\pi$ from the optimal stimulus are also searched and analyzed.
-Additionally, these methods are capable of characterizing both scalar and vector representations, where the ``fitnesses'' to be optimized correspond to Eq.~\ref{eq:O1}, \ref{eq:I1}, \ref{eq:S1} and Eq.~\ref{eq:O2}, \ref{eq:I2}, \ref{eq:S2}, respectively.
+Additionally, these methods can characterize both scalar and vector representations, in which the ``fitnesses'' to be optimized correspond to Eq.~\ref{eq:O1}, \ref{eq:I1}, \ref{eq:S1} and Eq.~\ref{eq:O2}, \ref{eq:I2}, \ref{eq:S2}, respectively.
 In this example, the scalar representation (\ie response) of the given neuron is tuned to a $45^{\circ}$ Gabor filter and invariant to phase changes, but selective to orientation changes.
@@
 To study the representations inside an artificial neural network, methods adopted in this work include: (1) first-order (\ie linear) characterization, or search and analysis of optimal stimuli, and (2) second-order (\ie quadratic) characterization, or search and analysis of invariance and selectivity paths (\ie eigenvectors of the largest and smallest eigenvalues), which are generalized and improved from previous work.
-$\bf{r}$ is defined as a scalar representation (\ie $\bf{r} \in \mathbb{R}$) when considering single neurons and a vector representation (\ie ${\bf{r}} \in \mathbb{R}^C$) when considering groups of neurons (with group size $C$).
+In this paper, $\bf{r}$ is defined as a scalar representation (\ie $\bf{r} \in \mathbb{R}$) when considering single neurons and a vector representation (\ie ${\bf{r}} \in \mathbb{R}^C$) when considering groups of neurons (with group size $C$).
 Our methods support characterizing both scalar (using Eq.~\ref{eq:O1}, \ref{eq:I1}, \ref{eq:S1}, though Eq.~\ref{eq:O2}, \ref{eq:I2}, \ref{eq:S2} can be used as well) and vector (using Eq.~\ref{eq:O2}, \ref{eq:I2}, \ref{eq:S2}) representations in a unified way.
@@
 
-While no specific knowledge of the network itself is assumed, we do restrict the space of stimuli considered with a constant energy constraint $\left\| \bf{x} \right\| = E$, where $E$ is set to $1$ in all mathematical derivations for simplicity, but set to the average energy of task-related images in our experiments.
+While no specific knowledge of the network itself is assumed, we do restrict the space of stimuli considered with a constant energy constraint $\left\| \bf{x} \right\| = E$, where $E$ in all mathematical derivations is set to $1$ for simplicity, but in our experiments, set to the average energy of task-related images.
 Since neurons can often be modulated by stimulus contrasts, which is less interesting when considering pattern selectivities, limiting the stimulus space in this manner effectively avoids degenerate solutions that simply maximize stimulus contrasts, and reduces the range of possible stimuli to consider.
@@
 
-\newcommand{\expdiff}{The way the simple linear constraint is constructed to enforce the exploration of a larger extent of the fitness landscape is also one of the main differences compared to \cite{erhan2010understanding}.}
+\newcommand{\expdiff}{The way we constructed the simple linear constraints to enforce the exploration of the landscape is also one of the main differences compared to methods in \cite{erhan2010understanding}.}
+%to enforce the exploration of a larger extent of the fitness landscape
 
@@
 \end{align}
-in which $0 < \delta \le \frac{\pi}{2}$, subject to $\left\| \bf{x}_{\delta} \right\| = 1$ and $\langle \bf{x}_{\delta} , \hat{\bf{x}} \rangle = \cos\left(\delta\right)$. 
-As visualized in Fig.~\ref{fig:methods}, the invariance path characterizes the $N$ dimensional curve that leads toward (or ideally, maintains at) a fitness as high as possible while moving away from the optimal stimulus, and the selectivity path characterizes the curve that leads toward fitness as low as possible while moving away.
+in which $0 < \delta \le \frac{\pi}{2}$, subject to $\left\| \bf{x}_{\delta} \right\| = 1$ and $\langle {\bf{x}}_{\delta} , \hat{{\bf{x}}} \rangle = \cos\left(\delta\right)$. 
+As visualized in Fig.~\ref{fig:methods}, the invariance path characterizes the $N$ dimensional curve that leads toward (or ideally, maintains at a) fitness as high as possible while moving away from the optimal stimulus, and the selectivity path characterizes the curve that leads toward fitness as low as possible while moving away.
 The search process is implemented as multiple runs of maximization and minimization on discretized $\delta \in \left\lbrace 0.1\pi, 0.2\pi, 0.3\pi, 0.4\pi, 0.5\pi\right\rbrace$ as the circular distance constraints shown in Fig.~\ref{fig:methods}, where each run is initialized with the result from the previous run (and the $0.1\pi$ run directly with optimal stimulus $\hat{\bf{x}}$) to increase the path continuity and searching speed\footnote{\expdiff}.
 In this work, the distance constraint $\delta$ only goes up to $\frac{\pi}{2}$, since on the $N$ dimensional sphere, stimuli that fall in $\frac{\pi}{2} < \delta \le \pi$ are simply ``negatives'' of those in $0 < \delta \le \frac{\pi}{2}$, which are of less uniqueness and interest in general.
-The same process can also be performed with respect to a reference stimulus $\tilde{\bf{x}}$, especially for vector representations, where the invariance and selectivity of the ``imaginary neuron'' are again characterized equivalently.
+The same process can also be performed with respect to a reference stimulus $\tilde{\bf{x}}$, especially for vector representations, where the invariance and selectivity of the aforementioned ``imaginary neuron'' are again characterized equivalently.
 
@@
 Most importantly, since it does not require training (except for the linear SVM on top of the representations extracted), a large number of networks with diverse hyperparameters can be rapidly generated and compared, which is particularly useful in this study where relative, instead of absolute task performances, are of primary interest.
-As one might be concerned that comparing randomly weighted networks will not tell much about representations, since they might as well be random, our results will suggest otherwise.
+One might be concerned that comparing randomly weighted networks will not tell much about representations; however, our results will suggest otherwise.
+
+%(since they might as well be random)
 
@@
 \begin{center}
-\includegraphics[width=0.80\textwidth]{Figs/pic1.jpg} 
+\includegraphics[width=0.845\textwidth]{Figs/pic1.jpg} 
 \end{center}
@@
 
+{\bf Optimal stimulus spectral complexity.} Figure \ref{fig:allrep}A and \ref{fig:allrep}E illustrate the trajectories of optimal stimulus searches of 32 top-layer neurons from the best performing shallow network and 32 from the best performing deep network respectively.
+To further increase the searching speed, the initial stimulus {was} selected from 1000 ${1}/{f^{\alpha}}$ random stimuli with the best fitness, where $\alpha \in \left\lbrace -4,-3,-2,-1,0 \right\rbrace$, without overly sacrificing the nature of random initialization.
+From the results, first, the response landscapes of deep neurons are noticeably more complex than those of shallow neurons, as a large fraction of search trajectories for deep neurons actually go down then up before reaching the (local) optima, which also suggests the numerical solver can handle non-convexity reasonably well.
+Second, the optimal stimuli of deep neurons are also relatively more complex (\ie consisting of more textural components) than those of shallow neurons (which are more uni-textured or Gabor-like).
+To quantify such differences, the spectral complexity measure, which is estimated through the $L^{1}$ norm of the two dimensional Fourier power spectrum of optimal stimulus, \ie $\left\| \mathcal{F}(\hat{\bf{x}}) \right\|_{1}$, is adopted, where a higher value suggests higher non-sparsity (\ie more spectral components required to represent the visual stimuli).
+
 \begin{figure}
@@
 Fitnesses of invariance (red dots, \ie $\left\lbrace f\left(\bf{x}^{+}_{\delta}\right) \right\rbrace$) and selectivity paths (blue dots, \ie $\left\lbrace f\left(\bf{x}^{-}_{\delta}\right) \right\rbrace$) of all 3,200 shallow and 3,200 deep neurons are shown for comparison\protect\footnotemark.
-Although both types of neurons show invariance, shallow neurons do not posses selectivity, since their selectivity paths overlap with the zero-selectivity baseline.
-In other words, \emph{only deep neurons are selective}.
+Although both types of neurons show invariance, shallow neurons do not posses selectivity, since their selectivity paths overlap with the zero-selectivity baseline, \ie \emph{only deep neurons are selective}.
 } %, and random walks (green dots, \ie $\left\lbrace f\left({\bf{x}}^{R}_{\delta}\right) \right\rbrace$)
@@
 \end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=0.80\columnwidth]{Figs/fig5.pdf} 
-\end{center}
-\caption{{\bf Comparison of shallow and deep scalar representation measures.}
-Representation measures from left to right are, optimal stimulus spectral complexity (OSSC), invariance path potential (INPP), selectivity path potential (SLPP), and invariance subspace capacity (INSC).
-Sensitivity indexes ($d'$) between distributions are $1.63$, $0.42$, $4.73$, and $3.67$, and differences between means of distributions are all significant (***), which suggest \emph{deep representations are quantitatively and significantly better than shallow representations}.
-} %, in which data points from shallow networks (left distributions) and deep networks (right distributions) are shown side by side
-\label{fig:pair}
-\end{figure}
-
-{\bf Optimal stimulus spectral complexity.} Figure \ref{fig:allrep}A and \ref{fig:allrep}E illustrate the trajectories of optimal stimulus searches of 32 top-layer neurons from the best performing shallow network and 32 from the best performing deep network respectively.
-To further increase the searching speed, the initial stimulus {was} selected from 1000 ${1}/{f^{\alpha}}$ random stimuli with the best fitness, where $\alpha \in \left\lbrace -4,-3,-2,-1,0 \right\rbrace$, without overly sacrificing the nature of random initialization.
-From the results, first, the response landscapes of deep neurons are noticeably more complex than those of shallow neurons, as a large fraction of searching trajectories for deep neurons actually go down then up before reaching the (local) optima, which also suggests the numerical solver can handle non-convexity reasonably well.
-Second, the optimal stimuli of deep neurons are also relatively more complex (\ie consisting of more textural components) than those of shallow neurons (which are more uni-textured or Gabor-like).
-To quantify such differences, the spectral complexity measure, which is estimated through the $L^{1}$ norm of the (2 dimensional) Fourier power spectrum of optimal stimulus, \ie $\left\| \mathcal{F}(\hat{\bf{x}}) \right\|_{1}$, is adopted, where a higher value suggests higher non-sparsity (\ie more spectral components required to represent the visual stimuli).
 
@@
 As further demonstrated in the fitness-distance diagrams \cite{jones1995fitness} (Fig.~\ref{fig:fda}), on average, although shallow neurons have good invariance (\ie red curve stays high), they do not have any selectivity (\ie blue curve drops slow), while deep neurons show both good invariance and selectivity (\ie red curve stays high and blue curve drops fast).
-For scalar representations, we can define the invariance (and selectivity, similarly) path potential as the area sandwiched between the invariance curve and the cosine baseline curve (\ie the invariance and selectivity curves of a single inner-product neuron\footnote{\defbaseline}, or the ``baseline'' of neural network types) on the fitness-distance diagram, \ie $\int_{0}^{\frac{\pi}{2}}{\left| \cos^{-1}\left(f\left( \bf{x}^{+}_{\delta} \right)\right) - \delta \right|}\mathrm{d}\delta \mathbin{/} {\frac{\pi}{2}}$ as normalized in the $\cos^{-1}$ domain, to measure how invariant (and selective) a neuron is compared to the baseline.
+For scalar representations, we can define the invariance (and selectivity, similarly) path potential as the area sandwiched between the invariance curve and the baseline curve (\ie the invariance and selectivity curves of a single inner-product neuron\footnote{\defbaseline}, or the ``baseline'' of all neural network types) on the fitness-distance diagram, \ie $\int_{0}^{\frac{\pi}{2}}{\left| \cos^{-1}\left(f\left( \bf{x}^{+}_{\delta} \right)\right) - \delta \right|}\mathrm{d}\delta \mathbin{/} {\frac{\pi}{2}}$ as normalized in the $\cos^{-1}$ domain, to measure how invariant (and selective) a neuron is compared to the baseline.
 
@@
 
-{\bf Invariance subspace capacity.} To further characterize the invariance properties, multiple runs of invariance path searches {were} performed and the results are visualized in Figs.~\ref{fig:allrep}D and \ref{fig:allrep}H.
+\newcommand{\expnoslsc}{Selectivity subspace capacity, on the other hand, is not a fair measure and thus not included, since shallow neurons do not even have selectivity close to deep neurons', and in a sense have more diverse ``most selective'' paths, namely noises.}
+
+{\bf Invariance subspace capacity.} To further characterize the properties of invariance, multiple runs of invariance path searches {were} performed, and the results are visualized in Figs.~\ref{fig:allrep}D and \ref{fig:allrep}H.
 The diversity of the ``invariance plateau'' (\eg the central high fitness region in Fig.~\ref{fig:methods}) constitution of deep neurons is also more apparent than that of shallow neurons.
-Compared to the invariance path potential, which is designed to characterize the best invariance path even if only one of such exists, we further adopted the invariance subspace capacity measure to estimates the ``dimensionality'' (\ie how diverse different paths can be) of the linear subspace formed by multiple path search results via the nuclear norm of concatenation of $n$ results $\left\|\left[{\bf{x}}_{\delta,1},\dots,{\bf{x}}_{\delta,n}\right]\right\|_{*}$ with $n=20$ and $\delta=0.1\pi$.
+In addition to the invariance path potential, which is designed to characterize the ``best invariance path'' even if only one of such exists, we further adopted the invariance subspace capacity measure\footnote{\expnoslsc} to estimates the ``dimensionality'' (\ie how diverse different paths can be) of the linear subspace formed by multiple path search results via the nuclear norm of concatenation of $n$ results $\left\|\left[{\bf{x}}_{\delta,1},\dots,{\bf{x}}_{\delta,n}\right]\right\|_{*}$ with $n=20$ and $\delta=0.1\pi$.
 
@@
 
-\newcommand{\defstars}{In this paper, significance levels follow the definition that *, ** and *** correspond to $p<0.05$, $p<0.01$ and $p<0.001$ respectively.\label{fnote:defstars}}
+\begin{figure}
+\begin{center}
+\includegraphics[width=0.80\columnwidth]{Figs/fig5.pdf} 
+\end{center}
+\caption{{\bf Comparisons of shallow and deep scalar representation measures.}
+Representation measures from left to right are, optimal stimulus spectral complexity (OSSC), invariance path potential (INPP), selectivity path potential (SLPP), and invariance subspace capacity (INSC).
+Sensitivity indexes ($d'$) between shallow and deep measures are $1.63$, $0.42$, $4.73$, and $3.67$ respectively, and the differences of means between shallow and deep measures are all significant\protect\footnotemark, which suggest \emph{deep representations are quantitatively and significantly better than shallow representations}.
+} %, in which data points from shallow networks (left distributions) and deep networks (right distributions) are shown side by side
+\label{fig:pair}
+\end{figure}
+
+\newcommand{\defstars}{In this paper, significance levels follow the definition that *, ** and *** correspond to $p<0.05$, $p<0.01$ and $p<0.001$ respectively, where all $p$-values {are} calculated using permutation tests.}
+
+\footnotetext{\defstars \label{fnote:defstars}}
+
+\begin{figure*}
+\begin{center}
+\includegraphics[width=0.86\textwidth]{Figs/pic2.jpg} 
+\end{center}
+\caption{{\bf Explanatory power of optimal stimuli of deep neurons.}
+Results from deep networks of the top 3 strongest average explanatory power (A--C) and the weakest average explanatory power (D) are visualized.
+Within each panel, the first row shows all 32 neurons' optimal stimuli from the network, the third to fifth rows show natural images with the top 3 strongest ``explainabilities'' (\ie inner-product distances between the images and the corresponding optimal stimuli), and the second row shows the average of images with top 100 explainabilities, out of 10,000 natural images.
+Color of the boarder of an image indicates the linear explainability of the image, with color map following the definition in Fig.~\ref{fig:methods}.
+Comparing deep networks' explanatory power and performances shows significant correlation (OSEP measure in Fig.~\ref{fig:corr}), which suggests \emph{``representing'' natural images can be required in first-order properties of good deep scalar representations (\ie optimal stimuli)}.
+}
+\label{fig:exppow}
+\end{figure*}
 
 {\bf Overall comparison.} Comparisons of spectral complexities, invariance/selectivity path potentials, and invariance subspace capacities of shallow and deep neurons (as scalar representations) are summarized in Fig.~\ref{fig:pair}.
-These measures further quantify the significant\footnote{\defstars} differences between shallow and deep representations, in addition to the separability and classifiability used in previous studies \cite{donahue2014decaf, zeiler2014visualizing}.
+These measures further quantify the significant differences between shallow and deep representations, in addition to the separability and classifiability used in previous studies \cite{donahue2014decaf, zeiler2014visualizing}.
 For instance, subtle differences of visual features have better chances being distinguished when the ``gap'' between invariance and selectivity curves (\ie the dynamic range of neuronal responses, or amplification ratio of differences) is large.
-The highly structural representations of neurons from deep networks with purely random convolution kernels found in the experiments also explain why such ``random'' deep networks can still have good performances, on top of the theoretical analysis for (one-level) shallow networks \cite{saxe2011random}.
+The highly structural representations of neurons from deep networks using purely random convolution kernels found in the experiments also explain why such ``random'' deep networks can still have good performances, on top of the theoretical analysis for (one-level) shallow networks \cite{saxe2011random}.
 
@@
 
-\begin{figure*}
-\begin{center}
-\includegraphics[width=0.85\textwidth]{Figs/pic2.jpg} 
-\end{center}
-\caption{{\bf Explanatory power of optimal stimuli of deep neurons.}
-Results from deep networks of the top 3 strongest average explanatory power (A--C) and the weakest average explanatory power (D) are visualized.
-Within each panel, the first row shows all 32 neurons' optimal stimuli from the network, the third to fifth rows show natural images with the top 3 strongest ``explainabilities'' (\ie inner-product distances between the images and the corresponding optimal stimuli), and the second row shows the average of images with top 100 explainabilities, out of 10,000 natural images.
-Color of the boarder of an image indicates the explainability of the image, with color map following the definition in Fig.~\ref{fig:methods}.
-Comparing deep networks' explanatory power and performances shows significant correlation (OSEP measure in Fig.~\ref{fig:corr}), which suggests \emph{``representing'' natural images can be required in first-order properties of good deep scalar representations (\ie optimal stimuli)}.
-}
-\label{fig:exppow}
-\end{figure*}
+{\bf Optimal stimulus explanatory power.} To understand how the optimal stimulus may affect a deep network's performance, we {measured} how well the optimal stimulus can ``linearly explain'' natural images and how the explanatory power correlates with a network's performance.
+In specific, 10,000 image patches in total {were} randomly cropped out of the predefined facial regions of pictures from the LFW-a dataset \cite{wolf2011effective} and compared against the optimal stimuli of all 32 top-layer neurons from all 100 deep networks.
+Figure \ref{fig:exppow} shows the results of best 3 and worst 1 networks in terms of explanatory power, which is estimated as the mean of the linear explainability of natural images, or inner-product distances between the optimal stimulus and $n$ nearest natural images $\left\lbrace {\bf{x}}^{t} \right\rbrace$, \ie $\frac{1}{n}\sum_{i=1}^{n} \langle {\bf{x}}^{t}_{i} , \hat{{\bf{x}}} \rangle$ with $n=100$, to measure how well the optimal stimulus can linearly ``approximate'' natural images (or intuitively how much neuronal response might be elicited).
+It can be observed that, though none of the natural images is apparently similar to any of the optimal stimuli, when alternatively comparing the averages of natural images of the top 1\% explainabilities and the optimal stimuli, the similarities (Fig.~\ref{fig:exppow}A.1 \vs \ref{fig:exppow}A.2, \etc) and dissimilarities (Fig.~\ref{fig:exppow}D.1 \vs \ref{fig:exppow}D.2) become apparent.
+Such kind of ``linearities'' around task-related images out of the well-performing highly-nonlinear networks were observed in previous studies \cite{le2012building, simonyan2013deep} too, where optimal stimuli also resembled averages or combinations of visual objects.
+
+%For simplicity and clarity of presentation, the explanation power measure is calculated using the top 1\% (i.e.~100) of the task-related stimuli with highest explainabilities, without biasing the results compared to using all task-related stimuli.
 
@@
 \begin{center}
-\includegraphics[width=\columnwidth, trim=1em 0 1em 0]{Figs/pic3.jpg} 
+\includegraphics[width=1.025\columnwidth, trim=1em 0 1em 0]{Figs/pic3.jpg} 
 \end{center}
 \caption{{\bf Alignments of subspaces between invariance and selectivity paths and natural images.}
-Using 16 natural images in A as reference stimuli, the invariance paths with $\delta = 0.1\pi$ corresponding to the best alignments against natural images (\ie $L^1$ sparsities of images represented in eigenface coordinates) among all deep networks are shown in B.1, and similarly, worst alignments of invariance paths in C.1, best alignments of selectivity paths in D.1, and worst alignments of selectivity paths in E.1.
+Using 16 natural images in A as reference stimuli, the invariance paths at $\delta = 0.1\pi$ corresponding to the best alignments against natural images (\ie high $L^1$ sparsities of images represented in eigenface coordinates) among all deep networks are shown in B.1, and similarly, best alignments of selectivity paths in C.1, worst alignments of invariance paths in D.1, and worst alignments of selectivity paths in E.1.
 Differences between A and B.1--E.1 are shown in B.2--E.2 for clearer visualization.
 Color of the boarder of an image indicates the alignment of the image against natural images, with color map following the definition in Fig.~\ref{fig:methods}.
-As good alignments (B.1 and C.1) appear more natural (like structural deformations or lighting changes), bad alignments are mostly meaningless noises.
+While good alignments (B.1 and C.1) appear more natural (like structural deformations, lighting changes, \etc), bad alignments are mostly meaningless noises.
 Comparing deep networks' subspace alignment measures and performances shows significant correlations (ITSA/STSA measures in Fig.~\ref{fig:corr}), which suggest \emph{a resemblance to natural image distributions can be required in second-order properties of good deep vector representations (\ie invariance and selectivity subspaces)}.
@@
 
+{\bf Invariance and selectivity \vs natural image subspace alignment.} It has been demonstrated that invariance and selectivity are extremely important for biological vision \cite{desimone1991face, ito1995size}; however, how these properties directly affect the recognition performance of artificial neural networks remains unclear.
+To address this question, starting from 16 randomly sampled reference stimuli, invariance and selectivity path searches {were} performed at $\delta = 0.1\pi$ using Eq.~\ref{eq:I2}, \ref{eq:S2} for vector representations.
+Figure \ref{fig:align}B--E demonstrate the results of invariance and selectivity subspace alignment analyses, which measure the alignments between subspaces formed by natural images and invariance and selectivity paths (\ie how likely the invariance and selectivity can benefit recognition) via estimating the sparsity of $n$ paths projected onto the principal component vectors $\bf{V}$ of natural images $\left\lbrace {\bf{x}}^{t} \right\rbrace$, \ie $\frac{1}{n}\sum_{i=1}^{n} \left\| {\bf{Vx}}_{\delta,i} \right\|_{1}$ with $n=20$ and $\delta=0.1\pi$.
+
+%quiroga2005invariant
+%It can be observed that good alignments (\ie sparser representations in the principle component vector spaces) correspond to highly structural deformations or lighting changes, which are known to be important factors in visual recognition, while bad alignments correspond to mostly meaningless noises.
+
 \begin{figure}
 \begin{center}
-\includegraphics[width=\columnwidth, trim=1em 0 1em 0]{Figs/pic4.jpg} 
+\includegraphics[width=1.025\columnwidth, trim=1em 0 1em 0]{Figs/pic4.jpg} 
 \end{center}
-\caption{{\bf Reconstructability of deep vector representations (\ie encoding specificities of natural images).}
-Using 16 natural images in A as reference stimuli, 3 examples of reconstructions from all deep networks with the best, second best, third best and worst average reconstructed image quality (SSIM) are shown in B--E respectively.
-Color of the boarder of an image indicates the SSIM score between the image and reference stimulus, with color map following the definition in Fig.~\ref{fig:methods}.
+\caption{{\bf Encoding specificities for natural images (\ie reconstructabilities of deep vector representations).}
+Utilizing 16 natural images in A as reference stimuli, 3 examples of reconstructions from deep networks corresponding to the best, second best, third best and worst average qualities (SSIM) of reconstructions are shown in B--E respectively.
+Color of the boarder of an image indicates the SSIM between the image and reference stimulus, with color map following the definition in Fig.~\ref{fig:methods}.
 Though the reconstructions are in general of lower spatial frequencies, some of the better ones do capture the spatial structures of reference stimuli decently well (\eg stimulus 1, 5, \etc).
-Comparing deep networks' encoding specificities and performances shows significant correlations (ITSA/STSA measures in Fig.~\ref{fig:corr}), which suggest \emph{providing specific (\ie non-confounding) encoding of natural images can be required in first-order properties of good deep vector representations}.
+Comparing deep networks' encoding specificities and performances shows significant correlations (ITSA/STSA measures in Fig.~\ref{fig:corr}), which suggest \emph{providing specific (\ie non-confounding) encoding for natural images can be required in first-order properties of good deep vector representations}.
 } % (reconstructed)
@@
 
-\begin{figure}
-\begin{center}
-\includegraphics[width=0.95\columnwidth]{Figs/e_fig7_scatter-crop.pdf} 
-\end{center}
-\caption{{\bf Spearman's correlations between deep network's representation measures and performance.}
-Representation measures following the raster order are, optimal stimulus explanatory power (OSEP), invariance path potential (INPP), selectivity path potential (SLPP), invariance subspace capacity (INSC), invariance \vs task-related image subspace alignment (ITSA), selectivity \vs task-related image subspace alignment (STSA), task-related image encoding specificity (TSES), and all 7 measures together using linear multiple correlation analysis.
-Correlations $\rho$ and significance levels\footref{fnote:defstars} are $0.60^{\ast\ast\ast}$, $-0.31^{\ast\ast}$, $-0.24^{\ast}$, $-0.39^{\ast\ast\ast}$, $-0.44^{\ast\ast\ast}$, $-0.56^{\ast\ast\ast}$, $0.64^{\ast\ast\ast}$, and $0.84^{\ast\ast\ast}$, respectively.
-Overall, \emph{71\% of the variance ($\rho^2$) of deep networks' performances can be explained} using the proposed representation measures altogether.
-}
-\label{fig:corr}
-\end{figure}
-
-{\bf Optimal stimulus explanatory power.} To understand how the optimal stimulus may affect a network's performance, we {tested} how well the optimal stimulus may ``linearly explain'' natural images and how the explanatory power correlates with a network's performance.
-In total, 10,000 image patches {were} randomly cropped out of the predefined facial regions of pictures from the LFW-a dataset \cite{wolf2011effective} and compared against the optimal stimuli of all 32 top-layer neurons from all 100 deep networks.
-Figure \ref{fig:exppow} shows the results of best 3 and worst 1 networks in terms of explanatory power, which is estimated as the mean of the linear explainability of natural images, or inner-product distances between the optimal stimulus and $n$ nearest natural images $\left\lbrace {\bf{x}}^{t} \right\rbrace$, \ie $\frac{1}{n}\sum_{i=1}^{n} \langle {\bf{x}}^{t}_{i} , \hat{{\bf{x}}} \rangle$ with $n=100$, to measures how well the optimal stimulus can linearly ``approximate'' natural images (or intuitively how much neuronal response might be elicited).
-It can be observed that, though none of the natural images is highly similar to any of the optimal stimuli, when alternatively comparing the averages of natural images of the top 1\% explainabilities and the optimal stimuli, the similarities (Fig.~\ref{fig:exppow}A.1 \vs \ref{fig:exppow}A.2, \etc) and dissimilarities (Fig.~\ref{fig:exppow}D.1 \vs \ref{fig:exppow}D.2) become apparent.
-Such kind of ``linearities'' around task-related images out of the well-performing highly-nonlinear networks were observed in previous studies \cite{le2012building, simonyan2013deep} too, where optimal stimuli also resembled averages or combinations of visual objects.
-
-%For simplicity and clarity of presentation, the explanation power measure is calculated using the top 1\% (i.e.~100) of the task-related stimuli with highest explainabilities, without biasing the results compared to using all task-related stimuli.
-
-{\bf Invariance and selectivity \vs natural image subspace alignment.} It has been shown that invariance and selectivity are extremely important for biological vision \cite{desimone1991face, ito1995size}; however, how these properties directly affect the recognition performance of artificial neural networks remains unclear.
-To address this question, starting from 16 randomly sampled reference stimuli, invariance and selectivity path searches {were} performed at $\delta = 0.1\pi$ using Eq.~\ref{eq:I2}, \ref{eq:S2} for vector representations.
-Figure \ref{fig:align}B--E demonstrate the results of invariance and selectivity subspace alignment analyses, which measures the alignment between subspaces formed by natural images and invariance and selectivity paths (\ie how likely the invariance and selectivity can benefit recognition) via estimating the sparsity of $n$ paths projected onto the principal component vectors $\bf{V}$ of natural images $\left\lbrace {\bf{x}}^{t} \right\rbrace$, i.e.~$\frac{1}{n}\sum_{i=1}^{n} \left\| {\bf{Vx}}_{\delta,i} \right\|_{1}$ with $n=20$ and $\delta=0.1\pi$.
-
-%quiroga2005invariant
-%It can be observed that good alignments (\ie sparser representations in the principle component vector spaces) correspond to highly structural deformations or lighting changes, which are known to be important factors in visual recognition, while bad alignments correspond to mostly meaningless noises.
-
 \newcommand{\expregular}{While Equation \ref{eq:O2} describes the essence of our approach, it was not able to produce analyzable numerical results alone, due to the fact that deep representations can be highly invariant to noises that do not resemble natural images.
 Implementation-wise, it was further regularized with the energy of Laplacian filtering $\left\| \mathcal{L} \left( \bf{x} \right) \right\|$ to enforce smoothness and similarity to natural images.
-A similar concept was also adopted in~\cite{mahendran2014understanding}.
+A similar concept was also adopted in \cite{mahendran2014understanding}.
 }  %, with regularization constant decided via optimizing withheld data.
 
-{\bf Encoding specificity of natural images.} To further understand the meaning of a vector representation $\bf{r}$ formed by a deep network $f$, the inverse function $f^{-1}\left( \bf{r} \right)$ {was} numerically approximated to reveal what kind of stimulus can drive the deep network to give output $\bf{r}$.
-Instead of inverting a randomly generated $\bf{r}$, which may not have feasible or interpretable solutions, we {used} $\bf{r}$ from a known reference stimulus $\tilde{\bf{x}}$ to study the encoding specificity and its relationship to performance.
+{\bf Encoding specificity for natural images.} To further understand the meaning of a vector representation $\bf{r}$ formed by a deep network $f$, the inverse function $f^{-1}\left( \bf{r} \right)$ {was} numerically approximated to reveal what kind of stimulus can drive the deep network to give output $\bf{r}$.
+Instead of inverting a randomly generated $\bf{r}$, which may not have feasible or interpretable solutions, we {used} $\bf{r}$ from known reference stimulus $\tilde{\bf{x}}$ to study the encoding specificity and its relationship to performance.
 Figure \ref{fig:encspc} shows the reference stimuli, the best 3 and worst 1 examples of encoding specificities, where the optimal stimulus search (Eq.~\ref{eq:O2}) is utilized as an inverse function of representations\footnote{\expregular}, and the average of structural similarities (SSIM) \cite{wang2004image} between the task-related reference stimulus $\tilde{\bf{x}}$ and $n$ reconstructed stimuli from random initializations, \ie $\frac{1}{n}\sum_{i=1}^{n} \mathrm{SSIM}\left(\tilde{\bf{x}} , \hat{{\bf{x}}}_i \right)$ with $n=10$, is measured to indicate how specific (or non-confounding) a representation is encoded.
@@
 Though bearing strong similarity to the invariance path search in terms of mathematical formulations, encoding specificity is in fact a radically different and complementary measure, since its numerical optimization is not constrained through distances to the reference stimulus (while distance constraints can induce \emph{a priori} similarities); thus, it can better function as an unconstrained global characterization of the ``encoding landscape'' (i.e.~multivariate tuning landscape of a population) and estimate if the (reconstructed) stimuli that a deep network is invariant to are ``selective'' (i.e.~visually similar to original stimuli).
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=0.95\columnwidth]{Figs/e_fig7_scatter-crop.pdf} 
+\end{center}
+\caption{{\bf Spearman's rank correlations between deep network's representation measures and performance.}
+Representation measures following the raster order are, optimal stimulus explanatory power (OSEP), invariance path potential (INPP), selectivity path potential (SLPP), invariance subspace capacity (INSC), invariance \vs task-related image subspace alignment (ITSA), selectivity \vs task-related image subspace alignment (STSA), task-related image encoding specificity (TSES), and all 7 measures together using linear multiple correlation analysis.
+Correlations ($\rho$) and significance levels\footref{fnote:defstars} are $0.60^{\ast\ast\ast}$, $-0.31^{\ast\ast}$, $-0.24^{\ast}$, $-0.39^{\ast\ast\ast}$, $-0.44^{\ast\ast\ast}$, $-0.56^{\ast\ast\ast}$, $0.64^{\ast\ast\ast}$, and $0.84^{\ast\ast\ast}$, respectively.
+Overall, \emph{71\% of the variance ($\rho^2$) of deep networks' performances can be explained} using the proposed representation measures altogether.
+}
+\label{fig:corr}
+\end{figure}
 
@@
 
-{\bf Overall comparison.} Spearman's correlations between all representation measures and 100 deep networks' performances are summarized in Fig.~\ref{fig:corr}, and 71\% of the variance of deep networks' performances are explainable with the proposed measures altogether using linear multiple correlation analysis.
-These measures further break down the differences among deep representations and suggest features like network's explanatory power, invariance and selectivity subspace properties, and encoding specificities against the task-related stimuli, can be extremely important for the effectiveness of representations.
-Invariance and selectivity path potentials\footnote{\defvecpp} and invariance subspace capacity\footnote{\expinsc} for deep vector representations, on the other hand, have lower correlations to network's performance (though still significant), possibly due to their nature of being less task-related\footnote{\expextrinsic} than subspace alignment measures.
+{\bf Overall comparison.} Spearman's correlations between all representation measures and 100 deep networks' performances are summarized in Fig.~\ref{fig:corr}, and 71\% of the variance of deep networks' performances is explainable with the 7 proposed measures altogether using linear multiple correlation analysis.
+These measures further break down the differences among deep representations and suggest features like network's explanatory power, invariance and selectivity subspace properties, and encoding specificities against the task-related stimuli, can be extremely important for the effectiveness of deep representations.
+Invariance and selectivity path potentials\footnote{\defvecpp} and invariance subspace capacity\footnote{\expinsc} for deep vector representations, on the contrary, have lower correlations to network's performance (though still significant), possibly due to their nature of being less task-related\footnote{\expextrinsic} than subspace alignment measures.
 Worth to note, although the first-order properties (OSEP and TSES) have stronger individual correlations, they do not explain more about the network's performance than the other second-order properties, using linear multiple correlation analyses ($0.56$ \vs $0.58$).
@@
 {\bf Contributions of this work.} Comparatively, the characterization of representations in this work is designed to be general and unbiased such that both first- and second-order methods can be supported for single neurons or groups of neurons, without assuming any \emph{de facto} property in the solutions (\eg parametric deformations as used for studying invariance in previous works).
-Additionally, with the strong statistical significances of results obtained from the large-scale experiments on 100 shallow and 100 deep networks, totalling 6,400 artificial neurons, we argue that our observations toward answering the differences between (shallow \vs deep) and among (deep) representations are highly conclusive, which can be summarized as:
+Additionally, with the strong statistical significances of results obtained from the large-scale experiments on 100 shallow and 100 deep networks, totaling 6,400 artificial neurons, we argue that our observations toward answering the differences between (shallow \vs deep) and among (deep) representations are highly conclusive, which can be summarized as:
 \begin{itemize} % Main Findings
@@
 \item Complexity of representations quantitatively increases along with network's depth.
-\item Unlike a deep representation which is both invariant and selective, shallow representation is only invariant, not selective, and its capacity of invariance is significantly lower than a deep representation's.
-\item Both first-order properties of deep representations (optimal stimulus explanatory power and encoding specificity of natural images) are crucially important to deep network performance.
+\item Unlike a deep representation, which is both invariant and selective, a shallow representation is only invariant, not selective, and its capacity of invariance is significantly lower than a deep representation's.
+\item Both first-order properties of deep representations (optimal stimulus explanatory power and encoding specificity for natural images) are highly important to deep network's performance.
 \item All second-order properties of deep representations (in particular, invariance and selectivity \vs natural image subspace alignment) together are as important as first-order properties, if not significantly more.
@@
 {\bf Future work.} In this work, we only studied shallow networks and the ``simplest'' type of deep networks, two-level networks.
-Although our results can potentially generalize if viewed as the characterization of ``adding an extra level'', it is still highly desirable to extend the study into deeper networks.
-While the methods and measures proposed in this work are neutral to the depth of networks, numerical difficulties (\eg evaluation and convergence speed, quality of solutions, \etc) may need to be addressed in very deep networks.
-Beyond explaining what properties make a deep representation good, it would be valuable too to incorporate these findings in learning even better representations, which are almost solely driven by bigger datasets nowadays, via certain kind of regularizations.
-We argue that this could be useful in accelerating or improving representation learning, especially when data are scarce and representations are not easily transferable.
+Our results can potentially generalize if viewed as the characterization of ``adding an extra level''.
+However, it is still highly desirable to extend the study into deeper networks.
+Even though the methods and measures proposed in this work are neutral to the depth of networks, numerical difficulties (\eg evaluation and convergence speed, quality of solutions, \etc) may need to be addressed in very deep networks.
+Beyond explaining what properties make a deep representation good, it would be valuable too to incorporate these findings in learning even better representations, which are almost solely driven by bigger datasets nowadays, \eg, via regularizing the representation measures.
+We argue that this could be useful in accelerating or improving representation learning, particularly when data are scarce and representations are not easily transferable.
 Finally, even though a decently high fraction of the variance in deep networks' performances has been explained, how to further increase this number, if there is an upper bound (\eg limits from dataset itself), \etc, need to be further investigated.

### 2015-04-07 paper_for_review.tex
v1.3

@@
 
-Although deep neural networks have been successfully pushing the boundaries of computer vision by breaking the records of ImageNet Challenges (ILSVRC) \cite{russakovsky} year after year \cite{krizhevsky2012imagenet, sermanet2013overfeat, szegedy2014going}, our understanding of the properties of these networks arguably remains limited, since we cannot yet fully answer the question ``what makes a deep network good?'' In this paper, we investigate two important facets of this question: (1) why deep networks are better than shallow networks, and (2) among networks of the same depth, why certain networks are better than others (Fig.~\ref{fig:perf}).
+Although deep neural networks have been successfully pushing the boundaries of computer vision by breaking the records of ImageNet Challenges (ILSVRC) \cite{russakovsky2014imagenet} year after year \cite{krizhevsky2012imagenet, sermanet2013overfeat, szegedy2014going}, our understanding of the properties of these networks arguably remains limited, since we cannot yet fully answer the question ``what makes a deep network good?'' In this paper, we investigate two important facets of this question: (1) why deep networks are better than shallow networks, and (2) among networks of the same depth, why certain networks are better than others (Fig.~\ref{fig:perf}).
 Specifically, from the point of view that successful networks are forming ``better'' representations of visual inputs that are more separable to the final-stage linear classifiers, we want to identify the crucial properties that are required for forming (or learning) better representations.

### 2015-06-03 nips2015.tex
nips

\documentclass[fleqn]{article} % For LaTeX2e
\usepackage{nips15submit_e,times}
\usepackage[hyperfootnotes=false]{hyperref}
\usepackage{url}
\usepackage{graphicx}
\usepackage[labelfont=bf]{caption}
\usepackage[outercaption]{sidecap} 
\usepackage{setspace}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{footmisc}
\usepackage{enumitem}
\usepackage{xspace}

\makeatletter
\DeclareRobustCommand\onedot{\futurelet\@let@token\@onedot}
\def\@onedot{\ifx\@let@token.\else.\null\fi\xspace}
\def\eg{\emph{e.g}\onedot} \def\Eg{\emph{E.g}\onedot}
\def\ie{\emph{i.e}\onedot} \def\Ie{\emph{I.e}\onedot}
\def\cf{\emph{c.f}\onedot} \def\Cf{\emph{C.f}\onedot}
\def\etc{\emph{etc}\onedot} \def\vs{\emph{vs}\onedot}
\def\wrt{w.r.t\onedot} \def\Wrt{W.r.t\onedot} \def\dof{d.o.f\onedot}
\def\etal{\emph{et al}\onedot}
\makeatother

%\title{How Representations Affect Recognition Performances of Deep Networks?}
\title{What Makes Deep Convolutional Representations Good for Visual Recognition?} %

\author{
David S.~Hippocampus\\
Department of Computer Science\\
Cranberry-Lemon University\\
\texttt{hippo@cs.cranberry-lemon.edu} \\
}

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

%\nipsfinalcopy % Uncomment for camera-ready version

\begin{document}

\maketitle

\begin{abstract}
Representations of artificial neurons in deep convolutional networks have been demonstrated to be highly effective for various visual recognition tasks \cite{szegedy2014going, schroff2015facenet, donahue2014decaf}.
However, our understanding of them arguably remains limited.
In this paper, we address two important questions: (1) why deep networks outperform shallow networks, and (2) among networks of the same depth, why certain networks outperform the others.
We considered the case study of face pair matching and conducted a large-scale experiment.
The experiment examined 200 randomly generated shallow and deep networks, with the results analyzed via a series of measures designed to quantify the differences in representations. 
We conclude through this principled analysis that deep representations are quantitatively better than shallow representations, and overall 71\% of the variance in the performance of deep networks can be explained with our measures, both of which are confirmed using statistical tests for the first time.
\end{abstract}

\begin{SCfigure}[1.5][h] % force here
\centering \includegraphics[width=0.40\textwidth]{Figs/perf.pdf} %\vspace{-1.5ex}
\caption{{\bf What makes a deep network good?}
Among existing work, researchers mostly tried to produce one ``best'' network architecture (\ie a hyperparameter set, either through manual tuning or automated hyperparameter search), which is subsequently evaluated on a particular benchmark.
Nevertheless, it is not fully understood why in practice deep networks (\eg two or more levels) on average are better shallow networks (\eg one level), and why certain networks are better than the others, \eg when considering a larger population of networks \cite{cox2011beyond} and their distribution of performance (as shown here) on task like LFW-a face pair matching \cite{wolf2011effective} (see Sec.~\ref{sec:methods} for details of the experimental setup). \vspace{-0.5ex}}
\label{fig:perf}
\end{SCfigure}

\section{Introduction}

\newcommand{\expwhyface}{Nevertheless, the methods and most observations presented in this work are in general not limited to just face pair matching, and may be generalized onto other domains in visual recognition.}

Although deep convolutional networks have been successfully pushing the boundaries of computer vision by breaking the records of datasets like ImageNet \cite{russakovsky2014imagenet} and LFW \cite{LFWTech} year after year \cite{krizhevsky2012imagenet, sermanet2013overfeat, szegedy2014going, taigman2014deepface, sun2014deep, schroff2015facenet}, our understanding of these networks' properties arguably remains limited, since we cannot yet fully answer the question ``what makes a deep network good?''
In this paper, we address two practical aspects of this important question: (1) why deep networks perform better than shallow networks, and (2) among networks of the same depth, why certain networks perform better than others (as illustrated in Fig.~\ref{fig:perf}).
More specifically, from the point of view that successful networks are forming ``better'' representations of visual inputs that are more separable to the final-stage linear classifiers, we want to identify the crucial properties that are required for forming (or learning) better representations. 
We adopted face pair matching (\ie identifying pairs of different pictures from the same person and rejecting those from different persons) as the benchmark task in our experiments for two main reasons: (1) representations of face recognition networks are far less studied than those of object recognition networks, and (2) face images can be easily aligned and extracted from irreverent backgrounds, which can serve as a better source of visual stimuli for studying representations that fundamentally facilitates recognition\footnote{\expwhyface}.

%\newcommand{\ivec}{\mathrm{vec}^{-1}}
\newcommand{\expstimdim}{For simplicity, a stimulus being 2-dimensional (\ie ${\bf{x}} \in \mathbb{R}^{\sqrt{N} \times \sqrt{N}}$) or vectorized (\ie ${\bf{x}} \in \mathbb{R}^N$) are used interchangeably.}

%first order, second order, flexibility
{\bf Existing work.} 
Given an artificial neural network $f$, the study of how $f$ represents a high dimensional stimulus\footnote{\expstimdim} ${\bf{x}} \in \mathbb{R}^N$ as a representation 
${\bf{r}}$ in existing work mainly focused on the first-order (\ie linear) characterization of $f$---the analysis of optimal stimuli, stimuli which lead to the strongest responses of given neurons.
For single-level convolutional networks using squared pooling, \ie $f_c({\bf{x}}) = \left\| {\bf{K}} \otimes {\bf{x}} \right\|_{F}$, the optimal stimulus was shown to be analytically approximable \cite{saxe2011random} using a Gabor-like filter, whether the convolution kernel $\bf{K}$ is structural or random.
For multi-level convolutional networks, the optimal stimuli of neurons from various depths were numerically approximated \cite{erhan2010understanding, ngiam2010tiled, le2012building, zeiler2014visualizing, simonyan2013deep} to qualitatively show neurons in deeper layers are tuned to more complex visual patterns.
The generalized concept of optimal stimulus when considering groups of neurons can also be adopted to study the representations by inverting them \cite{mahendran2014understanding}.

Beyond the first-order characterization, the analysis of invariance and selectivity properties of neurons (\eg changes in stimuli which given neurons are least and most sensitive to, how small or large the changes are, \etc) was also adopted in a few works.
This can be viewed as the second-order characterization of $f$, and parametric deformation in stimuli was the most commonly used setup.
For example, translation, rotation, scaling, \etc of the stimulus were used in \cite{goodfellow2009measuring, zeiler2014visualizing} to test the invariance and selectivity of neurons.
Similar deformations were also adopted in \cite{lenc2014understanding}, but extended to analyze more advanced properties (equivariance and equivalence) of groups of neurons.
More generalized setups that, \eg, utilize local quadratic information (\ie the Hessian) or ``global'' numerical searches in the stimulus space, are less explored in the context of artificial neural network studies. %even though they do exist, 
For quadratic networks, \ie $f_q\left({\bf{x}}\right) = \frac{1}{2}{\bf{x}}^{T}{\bf{Qx}}+{\bf{L}}^{T}{\bf{x}}+c$, through eigendecomposing the quadratic term $\bf{Q}$ (\ie the Hessian), local invariance and selectivity directions can be analytically derived \cite{berkes2006analysis}, which correspond to eigenvectors of the least and most negative eigenvalues.
This method also can be generalized onto multi-layer convolutional networks by numerically estimating the Hessian \cite{ngiam2010tiled}.
For multi-layer networks, characterization of the ``global'' invariance of single neurons can also be achieved by numerically searching for non-local (\ie farther to the optimal stimuli) solutions \cite{erhan2010understanding}.

While it is fully justifiable to qualitatively characterize and verify the effectiveness of representations in individual well-performing deep networks to be deployed in real applications, we however argue that this approach may not either teach us much about what really makes them good, and how to further improve them when even bigger training datasets are not easily available.
Since theoretical tools capable of fully harnessing these highly nonlinear networks still await discovery, we took a different route by studying a large population of networks and using statistical tools to help us sort out the crucial properties of better representations, through directly comparing representations from the shallow \vs deep networks, and the good \vs bad networks.

%{\bf Contributions of this work.} Comparatively, the characterization of representations in this work is designed to be general and unbiased such that both first- and second-order methods can be supported for single neurons or groups of neurons, without assuming any \emph{de facto} property in the solutions (\eg parametric deformations as used for studying invariance in previous works).
%Additionally, with the strong statistical significance of results obtained from the large-scale experiments on 100 shallow and 100 deep networks, totally 6,400 artificial neurons, we argue that our observations toward answering the differences between (shallow \vs deep) and among (deep) representations are highly evident and comprehensive, which cab be summarized as ***

%\section{Representation Characterization}
\section{Methods}
\label{sec:methods}

\begin{SCfigure}[1.5][t]
\includegraphics[width=0.40\textwidth, clip]{Figs/methods.pdf}
\caption{{\bf Outline of representation characterization.}
On the spherical (\ie constant energy) constraint of the $N$-dimensional stimulus space, the first-order property, \ie optimal stimulus $\hat{\bf{x}}$, of a representation $\bf{r}$ is first iteratively searched \cite{hansen2001completely} (as the evolution of gray eclipses) and analyzed.
Then, the second-order properties, \ie invariance path $\lbrace \bf{x}^{+}_{\delta} \rbrace$ and selectivity path $\lbrace \bf{x}^{-}_{\delta} \rbrace$ (as dashed red and blue curves) at distances $0.1\pi$ to $0.5\pi$ away from the optimal stimulus, are both searched and analyzed too.
In this example, the scalar representation (\ie response) of the given neuron is optimally tuned to the $45^{\circ}$ Gabor filter and invariant to phase changes, but selective to orientation changes.
\vspace{1.0ex}}
\label{fig:methods}
\end{SCfigure}

%representational property

%similar to previous work,
%However, unlike most previous work
%These methods can also characterize both scalar and vector representations, in which the ``fitnesses'' to be optimized correspond to Eq.~\ref{eq:O1}, \ref{eq:I1}, \ref{eq:S1} and Eq.~\ref{eq:O2}, \ref{eq:I2}, \ref{eq:S2}, respectively.
% (linear) (quadratic-like) (\ie eigenvectors of the largest and smallest eigenvalues)

To study the representations inside an artificial neural network, methods adopted in this work include: (1) first-order characterization, or search and analysis of optimal stimuli, and (2) second-order characterization, or search and analysis of invariance and selectivity paths, which are generalized and improved from previous work.
Throughout the paper, $\bf{r}$ is defined as a scalar representation (\ie ${\bf{r}} = f_s\left({\bf{x}}\right) \in \mathbb{R}$) when considering single neurons and a vector representation (\ie ${\bf{r}} = f_v\left({\bf{x}}\right) \in \mathbb{R}^C$) when considering groups of neurons (with group size $C$).
The search (\ie numerical optimization) phases are defined via the following equations, and the analyses of search results using representation measures will be covered in Sec.~\ref{sec:results1} and \ref{sec:results2}. 
We denote the following functions being optimized as \emph{fitness functions} and their values as \emph{fitnesses}.

\setlength{\mathindent}{1.0ex}
\begin{minipage}[c]{0.33\textwidth}
\begin{align}
\hat{\bf{x}} &= \underset{\bf{x}} {\arg\max} \ f_s(\bf{x}) \label{eq:O1} \\
&= \underset{\bf{x}}{\arg\max} \ e^{-\left\|f_v({\bf{x}})-f_v(\tilde{\bf{x}})\right\|} \label{eq:O2}
\end{align}
\end{minipage}
\begin{minipage}[c]{0.33\textwidth}
\begin{align}
\bf{x}^{+}_{\delta} &= \underset{\bf{x}_{\delta}}{\arg\max} \ f_s(\bf{x}_{\delta}) \label{eq:I1} \\
&= \underset{\bf{x}_{\delta}}{\arg\max} \ e^{-\left\|f_v({\bf{x}}_{\delta})-f_v(\hat{\bf{x}})\right\|} \label{eq:I2}
\end{align}
\end{minipage}
\begin{minipage}[c]{0.33\textwidth}
\begin{align}
\bf{x}^{-}_{\delta} &= \underset{\bf{x}_{\delta}}{\arg\min} \ f_s(\bf{x}_{\delta}) \label{eq:S1} \\
&= \underset{\bf{x}_{\delta}}{\arg\min} \ e^{-\left\|f_v({\bf{x}}_{\delta})-f_v(\hat{\bf{x}})\right\|} \label{eq:S2}
\end{align}
\end{minipage}

\newcommand{\expconst}{We restrict the space of stimuli considered with a constant energy constraint $\left\| \bf{x} \right\| = E$, where $E$ in all mathematical derivations is set to $1$ for simplicity, but in our experiments, set to the average energy of task-related images. This avoids degenerate solutions that simply maximize stimulus contrasts.}
%While no specific knowledge of the network itself is assumed

The optimal stimulus of a scalar representation (\ie a single neuron) can be defined as Eq.~\ref{eq:O1} subject to $\left\| \bf{x} \right\| = 1$\footnote{\expconst}.
When considering a vector representation (using Eq.~\ref{eq:O2}), it is equivalent to optimizing the response of an ``auxiliary neuron'' tuned to a certain reference stimulus $\tilde{\bf{x}}$.
With respect to the optimal stimulus $\hat{\bf{x}}$, the invariant and selective stimuli can be defined using Eq.~\ref{eq:I1} and \ref{eq:S1} respectively, where $0 < \delta \le \frac{\pi}{2}$, subject to $\left\| \bf{x}_{\delta} \right\| = 1$ and $\langle {\bf{x}}_{\delta} , \hat{{\bf{x}}} \rangle = \cos\left(\delta\right)$. 
The distance constraint, while being simple and linear, enforces the exploration of the landscape, which is one of the main improvements too compared to \cite{erhan2010understanding}.
The invariance path $\left\lbrace \bf{x}^{+}_{\delta} \right\rbrace$ and selectivity path $\left\lbrace \bf{x}^{-}_{\delta} \right\rbrace$ are then derived through multiple runs of maximization and minimization on discretized $\delta \in \left\lbrace 0.1\pi, 0.2\pi, 0.3\pi, 0.4\pi, 0.5\pi\right\rbrace$ as the distance constraints shown in Fig.~\ref{fig:methods}, where each run is initialized with the result from the previous run (and the $0.1\pi$ run directly with optimal stimulus $\hat{\bf{x}}$) to increase the path continuity and searching speed.
This can also be performed with respect to certain reference stimulus $\tilde{\bf{x}}$, especially for vector representations (using Eq.~\ref{eq:I2} and \ref{eq:S2}), where the invariance and selectivity of the auxiliary neuron are equivalently characterized again.

\newcommand{\expsettings}{
\Ie 32 channels of filters in the top level's convolution layers, a simple setting yielding reasonable performances.
Following the terminology in recent papers, shallow and deep neurons corresponded to \texttt{pool1} and \texttt{pool2} layers respectively.
Stimulus dimensionalities of the shallow and deep neurons {were} $N=121$ and $441$ (\ie spatially overlapping $11\times11$ and $21\times21$ receptive fields). % respectively
Other minor changes included: nonlinear activations all simplified to \emph{ReLU} \cite{krizhevsky2012imagenet} mode and normalizations all in subtractive mode.
Overall, the architectures were more similar to those in \cite{simonyan2014very}, except pooling operations can be \emph{average}, \emph{squared}, or \emph{max-like} in our case.
}

% \cite{fukushima1980neocognitron, lecun1998gradient, riesenhuber1999hierarchical, krizhevsky2012imagenet}
\newcommand{\expperf}{10-fold cross validation accuracies using the LFW view 2 splits (\cite{LFWTech}, totally 6,000 face pairs) are reported.
Although it is highly desirable to characterize representations from the state-of-the-art face pair matching networks \cite{taigman2014deepface, sun2014deep, schroff2015facenet}, neither those trained networks nor their datasets are publicly available.
Nonetheless, under the hard protocol adopted here (\ie image-restricted, no outside data, data augmentation, or multiple-crop evaluation), the (well-performing) random networks remain competitive against trained networks and do provide good representations in terms of accuracies attained.
} 
%State-of-the-art networks trained on massive private datasets \cite{taigman2014deepface, sun2014deep, schroff2015facenet} do have significantly higher accuracies. %absolute
%However, under the hard protocol adopted here (\ie image-restricted, no outside data), this model remains competitive among neural network-based algorithms.
%accuracy of identifying pairs of different pictures from the same person and rejecting those from different persons

{\bf Experimental setup.}
A simplified variety of deep convolutional neural networks \cite{krizhevsky2012imagenet, sermanet2013overfeat, szegedy2014going}---the Spatio-Temporal Hierarchical Object Representation (STHOR) model \cite{cox2011beyond, sthor}---{was} adopted in our experiments.
Like other deep networks, it consists of the standard cascade of convolution, nonlinear activation, pooling layers, \etc (which together define a single ``level'' in this work).
However, its convolution kernel weights are randomly assigned rather than trained.
Though not fully matching the performances of deep networks trained on massive quantities of data, it can in fact still be quite competitive for a variety of tasks, especially in the regime where large quantities of training samples are not publicly available. %\eg the LFW-a setting we used \cite{krizhevsky2012imagenet} \cite{pinto2009high, cox2011beyond, viglarge}.
Most importantly, since it does not require training---except for the linear SVM on top of the representations (\ie feature vectors) extracted---a large number of networks with diverse hyperparameters can be rapidly generated and compared, which is particularly useful in this study where relative, instead of absolute task performances, are of primary interest.
In the experiments, 100 shallow (\ie one-level) networks and 100 deep (\ie two-level) networks {were} randomly generated (except all with 32 top-layer neurons\footnote{\expsettings}) and tested to see how their representations differ with the networks' depths and affect the networks' performances on face pair matching against the LFW-a dataset \cite{wolf2011effective} as shown in Fig.~\ref{fig:perf}\footnote{\expperf}.

\section{Results: Shallow \vs Deep Representations}
\label{sec:results1}

\begin{figure}[t]
\centering \includegraphics[width=\textwidth]{Figs/s_vs_d_repr.pdf} 
\caption{{\bf Visualization of shallow and deep representations.} %scalar
Representations of 4 neurons randomly selected (out of all 32 neurons) from the best performing shallow and best performing deep networks are shown in the left and right panels respectively. 
The optimal stimuli, invariance paths, selectivity paths, and invariance subspaces (sampled using invariant stimuli $0.1\pi$ away from the optimal stimuli; 5 out of 20 results randomly selected) of the 4 shallow and deep neurons are ordered and shown accordingly, where color of the boarder of an image indicates the fitness of the representation (\ie response of the neuron) elicited by the image.
See supplementary material for more examples.
Comparatively, \emph{deep neurons show more complex and diverse representational properties} than shallow neurons do.
} % do visually
\label{fig:allrep1}
\end{figure}

{\bf Optimal stimulus spectral complexity.}
Figure \ref{fig:allrep1} demonstrates examples of the optimal stimuli of neurons from the best performing (\ie best LFW-a accuracy) shallow and the best performing deep network.
The optimal stimuli of deep neurons are visually much more complex (\ie consisting of more textural components) than those of shallow neurons (which are more uni-textured or Gabor-like).
To quantify such differences, the spectral complexity measure defined as the $L^{1}$ norm of the 2-dimensional Fourier power spectrum of optimal stimulus, \ie $\left\| \mathcal{F}(\hat{\bf{x}}) \right\|_{1}$, is adopted, where a higher value suggests higher non-sparsity (\ie more spectral components required to represent the visual stimuli).

% trajectories -> SI

\begin{figure}[t]
\centering \includegraphics[width=0.80\textwidth]{Figs/fda.pdf} 
\caption{{\bf Fitness-distance analyses of shallow and deep representations.} %scalar
Fitnesses of invariance (red dots, \ie $\left\lbrace f_s(\bf{x}^{+}_{\delta}) \right\rbrace$) and selectivity paths (blue dots, \ie $\left\lbrace f_s(\bf{x}^{-}_{\delta}) \right\rbrace$) of all 3,200 shallow (panel A) and 3,200 deep neurons (panel B) are shown for comparison\protect\footnotemark.
Although both types of neurons show invariance, shallow neurons do not posses selectivity, since their selectivity paths overlap with the zero-selectivity baseline (see text for definition), \ie \emph{only deep neurons are selective}.
}
\label{fig:fda}
\end{figure}

\footnotetext{A small fraction of invariance paths actually have fitnesses higher than the optimal stimuli's, due to the non-convexity of these networks, which however does not lead to noticeable differences in the results in Fig.~\ref{fig:pair}.}

\newcommand{\defbaseline}{Given a single inner-product neuron $f_n({\bf{x}}) = {\bf{w}}^{T}{\bf{x}}$ and $\left\| \bf{x} \right\| = 1$, we have $\hat{\bf{x}} = \bf{w}$ and thus the invariance (and selectivity) curve $f_n({\bf{x}}_{\delta}) = \langle \hat{\bf{x}} ,{\bf{x}}_{\delta} \rangle = \cos(\delta)$ by definition.}

{\bf Invariance and selectivity path potential.}
Figure \ref{fig:allrep1} also illustrates the invariance and selectivity paths of the same sets of shallow and deep neurons as aforementioned.
While invariance paths are mostly phase changes and selectivity paths are leading toward meaningless noises (all at the same falloff rate) in shallow neurons, both types of paths consist of sophisticated shape deformations in deep neurons.
Interestingly, although most shallow neurons are also selective to manually rotated optimal stimuli (\ie Gabor filters of different orientations) as well, their fitnesses still do not drop faster then the nonparametric numerical solutions as presented in Fig.~\ref{eq:I1}.
As further demonstrated in the fitness-distance diagrams \cite{jones1995fitness} (Fig.~\ref{fig:fda}), on average, although shallow neurons have good invariance (\ie red curve stays high), they do not have any selectivity (\ie blue curve drops slow), while deep neurons show both good invariance and selectivity (\ie red curve stays high and blue curve drops fast).
For scalar representations, we can define the invariance (and selectivity, similarly) path potential as the area sandwiched between the invariance curve and the baseline curve (\ie the invariance curves of a single inner-product neuron\footnote{\defbaseline}---the ``baseline'' of all types of neural networks) on the diagram, \ie $\int_{0}^{0.5\pi} \left| \cos^{-1}(f_s(\bf{x}^{+}_{\delta} )) - \delta \right| \mathrm{d}\delta$ (in the $\cos^{-1}$ domain for equalizing both types of path potentials), which measures how invariant (and selective) a neuron is compared to the baseline. %/ {\frac{\pi}{2}}

%(zero-invariance/-selectivity)
%In fact, this intriguing difference generalizes across all shallow \vs deep neurons experimented in this work.

\newcommand{\expnoslsc}{Selectivity subspace capacity, on the other hand, is not a fair measure and thus not included, since shallow neurons do not even have selectivity close to deep neurons', and in a sense have more diverse ``most selective'' paths---noises.}

{\bf Invariance subspace capacity.}
To further characterize the properties of invariance, multiple runs of invariance path searches {were} performed, and the results are also visualized in Fig.~\ref{fig:allrep1}.
The constitution of the ``invariance plateau'' (\ie the central high fitness region in Fig.~\ref{fig:methods}) of deep neurons is much more diverse than that of shallow neurons.
In addition to the invariance path potential, which is designed to characterize the ``best invariance path'' even when only one of such exists, we further adopted the invariance subspace capacity measure\footnote{\expnoslsc} to estimates the ``dimensionality'' (\ie how diverse different paths can be) of the linear subspace formed by multiple path search results using the nuclear norm of the concatenation of $n$ results $\left\|\left[{\bf{x}}_{\delta,1},\dots,{\bf{x}}_{\delta,n}\right]\right\|_{*}$ with $n=20$ and $\delta=0.1\pi$.

%(\ie dimensionality, or subspace capacity)

\begin{SCfigure}[1.5][t]
\includegraphics[width=0.40\textwidth]{Figs/s_vs_d.pdf} %\vspace{2.5ex}
%\begin{minipage}{0.40\textwidth}
%\includegraphics[width=\linewidth]{Figs/s_vs_d.pdf} %\vspace{-1.0ex}
%\scalebox{0.59}{\textsf{\emph{OSSC: Optimal Stimulus Spectral Complexity, INPP: INvariance Path Potential, SLPP: SeLectivity Path Potential, INSC: INvariance Subspace Capacity.}}}
%\end{minipage} 
\caption{{\bf Representation measures \vs depth of networks.} %scalar
Bhattacharyya distances and their significance levels\protect\footnotemark~between the 4 shallow and deep measures from left to right are $0.39^{\ast\ast\ast}$, $0.07^{\ast\ast\ast}$, $6.42^{\ast\ast\ast}$ and $1.27^{\ast\ast\ast}$ respectively, which suggest \emph{deep representations are quantitatively and significantly better than shallow representations}. \vspace{2.0ex} \\
\tiny{\textsf{\emph{OSSC: Optimal Stimulus Spectral Complexity, INPP: INvariance Path Potential, SLPP: SeLectivity Path Potential, INSC: INvariance Subspace Capacity.}}}
\vspace{1.5ex}}
\label{fig:pair}
\end{SCfigure}

\footnotetext{Significance levels *, **, and *** indicate $p<0.05$, $p<0.01$, and $p<0.001$, respectively, where all $p$-values {are} calculated using permutation tests (see supplementary material for algorithmic details). \label{fnote:defstars}}

{\bf Overall comparison.}
Comparisons of spectral complexities, invariance/selectivity path potentials, and invariance subspace capacities of all shallow and deep neurons (\ie scalar representations) are summarized in Fig.~\ref{fig:pair}.
These measures more precisely quantify the significant differences between shallow and deep representations, in comparison to the qualitative visualizations \cite{erhan2010understanding, ngiam2010tiled, zeiler2014visualizing} and separability/classifiability analysis \cite{donahue2014decaf, zeiler2014visualizing} adopted in previous studies.
For example, subtle differences of visual features have better chances being distinguished when the ``gap'' between the invariance and selectivity curves (\ie the dynamic range of neuronal responses, or amplification ratio of differences) is larger.
The highly structural representations of neurons from deep networks using purely random convolution kernels observed in the experiments also explain why such ``random'' deep networks can still have good performances, on top of the theoretical analysis for the one-level shallow networks \cite{saxe2011random}. %ADD MORE?

%The increase of complexities in deeper neurons' preferred/optimal stimuli also agrees with neurobiological findings in higher visual cortical areas (\eg V2 and V4 \cite{hegde2000selectivity, pasupathy2001shape}).

\section{Results: Good \vs Bad Representations}
\label{sec:results2}

\begin{figure}[t]
\centering \includegraphics[width=\textwidth, trim=1.0ex 0.5ex 0 0, clip]{Figs/g_vs_b_repr.pdf} 
\caption{{\bf Visualization of good and bad representations.}
Good and bad examples from 3 types of the representation measures among all deep networks are shown in the left and right panels respectively, where color of the boarder of an image indicates its value of the corresponding representation measure (see text for definitions). %representation characterizations
(1) \Wrt the TSEA measure, good representations can be inverted with much stronger similarities to the original reference stimuli.
(2) \Wrt the ITSSA/STSSA measures, the ``intrinsic'' invariant and selective directions ($\Delta$'s, differences between the reference and invariant/selective stimuli $0.1\pi$ away) of good representations are much more natural (like structural deformations, lighting changes, \etc), and the invariant and selective stimuli align better against the eigenface subspace, whereas those of bad representations are highly noisy and align poorly against the eigenface subspace.
(3) \Wrt the OTSA measure, the optimal stimuli of good representations are comparatively more correlated to face images, especially to the linear averages of the most similar (top 100) face images, whereas those of bad representations do not.
Comparing these measures against the performances of deep networks (see Fig.~\ref{fig:corr}) shows strong correlations, which imply \emph{these first- and second-order properties of good representations must well ``align with'' the distribution of task-related stimuli}, even if they are not explicitly engineered to do so (unlike generative models).
}
\label{fig:allrep2}
\end{figure}

{\bf Task-related stimulus encoding accuracy.}
To understand the meaning of a vector representation ${\bf{r}}$ formed by a deep network $f$, the inverse function $f^{-1}\left( \bf{r} \right)$ {was} numerically approximated (using Eq.~\ref{eq:O2}) to reveal what kind of stimulus can drive the deep network to give output $\bf{r}$.
Instead of inverting a randomly generated $\bf{r}$, which may not have feasible or interpretable solutions, we {used} $\bf{r}$ from known reference stimuli $\lbrace\tilde{\bf{x}}\rbrace$ (specifically, task-related stimulus, \ie randomly sampled face patches from the LFW-a dataset \cite{wolf2011effective}) to quantify the encoding accuracy and its relationship to performance.
Figure \ref{fig:allrep2} shows the reference stimuli and good \vs bad examples of encoding accuracies, where the average of structural similarities (SSIM) \cite{wang2004image} between the task-related reference stimulus $\tilde{\bf{x}}$ and $n$ inverted stimuli from multiple random initializations, \ie $\frac{1}{n}\sum_{i=1}^{n} \mathrm{SSIM}\left(\tilde{\bf{x}} , \hat{{\bf{x}}}_i \right)$ with $n=10$, is measured to estimate how accurate (\ie non-confounding) a representation is encoded.
This can also be interpreted as the first-oder alignment between the ``encoding landscape'' (\ie fitness landscape of the auxiliary neuron) and the task-related stimuli.
Our analysis predicts that strong encoding accuracies should also be observed in those even deeper and more powerful state-of-the-art networks. 
Interestingly, this in fact has been independently verified in recent studies \cite{mahendran2014understanding, long2014convnets, razavian2014persistent}, which also employed such kind of ``invertibility'' to study the representations. % of individual well-performing deep networks.
In addition, even though these deep convolutional representations are not formulated to be invertible (as opposed to, \eg, autoencoder's), our results further suggest that this property can actually be \emph{required} for a deep network to perform well.

%SSIM measured using default settings of the standard implementation \cite{wang2004image} while different settings also yield similar correlations; other similarity measures like SNR, PSNR, and inner-product distance yield similar correlations as well

{\bf Invariance and selectivity \vs task-related stimulus subspace alignment.} 
It is widely known that invariance and selectivity are essential for biological vision \cite{desimone1991face, ito1995size}; however, how these properties directly affect the recognition performances of artificial neural networks remains hypothetical.
To address this question, starting from randomly sampled reference stimuli, invariance and selectivity path searches {were} performed at $\delta = 0.1\pi$ using Eq.~\ref{eq:I2} and \ref{eq:S2} for vector representations. 
Figure \ref{fig:allrep2} demonstrates examples of good and bad alignments between the subspaces formed by the invariance/selectivity paths and the task-related stimuli (\ie how likely the invariance and selectivity can benefit recognition), which are defined as the sparsity of $n$ paths projected onto the task-related stimuli's principal component (or here, eigenface) vector space $\bf{V}$, \ie $\frac{1}{n}\sum_{i=1}^{n} \left\| {\bf{Vx}}_{\delta,i} \right\|_{1}$ with $n=20$ and $\delta=0.1\pi$.
Compared to previous work, our method does not assume that any kind of parametric deformation is important and thus should be studied.
Instead, the ``intrinsic'' invariant and selective directions of representations are directly searched and compared against the distribution of the task-related stimuli, which is more bias-free.
Surprisingly, highly natural invariance/selectivity are again observed in well-performing networks even given their nature of randomness.

%preferred
%$\left\lbrace {\bf{x}}^{t} \right\rbrace$
%quiroga2005invariant
%It can be observed that good alignments (\ie sparser representations in the principle component vector spaces) correspond to highly structural deformations or lighting changes, which are known to be important factors in visual recognition, while bad alignments correspond to mostly meaningless noises.

{\bf Optimal \vs task-related stimulus alignment.}
To understand how the optimal stimuli of individual neurons may affect a deep network's performance, we {measured} how well the optimal stimuli linearly aligns with the task-related stimuli, and how the alignment correlates with a network's performance.
Specifically, 10,000 face patches in total {were} randomly sampled from the LFW-a dataset and compared against the optimal stimuli of deep neurons.
Figure \ref{fig:allrep2} shows examples of good and bad alignments between the optimal stimuli and task-related stimuli, where the alignment measure is defined as the mean of the inner-product distances between an optimal stimulus and its $n$ nearest task-related stimuli $\left\lbrace {\bf{x}}^{t} \right\rbrace$, \ie $\frac{1}{n}\sum_{i=1}^{n} \langle {\bf{x}}^{t}_{i} , \hat{{\bf{x}}} \rangle$ with $n=100$, to estimate how well an optimal stimulus can linearly represent the task-related stimuli (or intuitively how responsive the corresponding neuron might be under the task).
Intriguingly, though none of the optimal stimuli is apparently similar to any of the task-related stimuli, when alternatively compared against the averages of the better aligned task-related stimuli (top 1\%), the similarities (in good representations) and dissimilarities (in bad representations) become apparent.
While it is unclear how to ``optimally place'' an optimal stimulus (or shape the entire landscape) in a high dimensional stimulus space under visual recognition tasks, based on the observation, we hypothesize that it should be closely surrounded by multiple visually similar task-related stimuli (but not directly on any individual task-related stimulus) such that the neuron is more likely to represent those similar stimuli (and only those stimuli) with similar values.
This may also explain why optimal stimuli often resemble the averages \cite{le2012building} or combinations \cite{simonyan2013deep} of multiple visual objects.

% (which in low dimensional cases are theoretically validated)
%Such kind of ``linearities'' around task-related images out of the well-performing highly-nonlinear networks were observed in previous studies \cite{le2012building, simonyan2013deep} too, where optimal stimuli also resembled averages or combinations of visual objects.
%much neuronal response might be elicited
%Specifically, 10,000 image patches in total {were} randomly cropped out of the predefined facial regions of pictures from the LFW-a dataset \cite{wolf2011effective} and compared against the optimal stimuli of all 32 top-layer neurons from all 100 deep networks.

%For simplicity and clarity of presentation, the explanation power measure is calculated using the top 1\% (i.e.~100) of the task-related stimuli with highest explainabilities, without biasing the results compared to using all task-related stimuli.

\begin{figure}[t]
\centering 
\begin{minipage}{0.90\textwidth}
\includegraphics[width=\linewidth]{Figs/corr.pdf} \vspace{-1.0ex}
{\tiny \begin{spacing}{1.2} \textsf{\emph{TSEA: Task-related Stimulus Encoding Accuracy, ITSSA: Invariance \vs Task-related Stimulus Subspace Alignment, STSSA: Selectivity \vs Task-related Stimulus Subspace Alignment, OTSA: Optimal \vs Task-related Stimulus Alignment.}} \end{spacing}} \vspace{-0.5ex}
\end{minipage} %\includegraphics[width=0.90\textwidth]{Figs/corr.pdf}
\caption{{\bf Representation measures \vs performance of networks.}
Spearman's rank correlations ($\rho$) and their significance levels\protect\footref{fnote:defstars}~of the 4 measures and all measures\protect\footnotemark~together (using linear multiple correlation analysis) against the network's performance are $0.64^{\ast\ast\ast}$, $-0.44^{\ast\ast\ast}$, $-0.56^{\ast\ast\ast}$, $0.60^{\ast\ast\ast}$, and $0.84^{\ast\ast\ast}$, respectively (see supplementary material for full analysis). In other words, \emph{71\% of the variance ($\rho^2$) of deep networks' performances can be explained} using the proposed representation measures altogether.
}
\label{fig:corr}
\end{figure}

\footnotetext{The rest measures which are included in the multiple correlation analysis but not individually reintroduced and shown in Fig.~\ref{fig:corr} are the invariance path potential, selectivity path potential, and invariance subspace capacity of vector representations. Their individual correlations and significance levels are $-0.31^{\ast\ast}$, $-0.24^{\ast}$, $-0.39^{\ast\ast\ast}$, respectively. See supplementary material for related details.}

{\bf Overall comparison.}
Spearman's rank correlations between all representation measures and all 100 deep networks' performances are summarized in Fig.~\ref{fig:corr}, and 71\% of the variance of deep networks' performances is explainable with the proposed measures altogether.
These measures further quantify the differences among deep representations and suggest features like network's 
encoding accuracies, invariance and selectivity subspaces' alignment, and optimal stimuli's alignment, against the task-related stimuli, can be extremely important for the effectiveness of deep representations.
Worth to note, although the first-order properties (TSEA and OTSA) do have stronger individual correlations, they cannot explain more about the network's performance than the other second-order properties ($0.56$ \vs $0.58$, again using linear multiple correlation analysis).

%Invariance and selectivity path potentials\footnote{\defvecpp} and invariance subspace capacity\footnote{\expinsc} for deep vector representations, on the contrary, have lower correlations to network's performance (though still significant), possibly due to their nature of being less task-related\footnote{\expextrinsic} than subspace alignment measures.

%Network's invariance subspace capacity (in terms of population representation), which though has an $R^2=0.15$ correlation as well, is in fact negatively correlated to network's performance, mainly due to the fact that poor performing network's invariance path search results are usually noisier (as depicted in Fig.~\ref{fig:align}D), and thus should be interpreted differently from the unit representation cases. on the other hand, are weaker in explaining networks' performances (with $R^2=0.10/0.06$), due to the fact that all deep networks have high but less ranked invariance and selectivity.

%Encoding specificity, as a single representation measure, can best explain the deep network's performance. Though bearing strong similarity to the invariance path search in terms of mathematical formulations, encoding specificity is in fact a radically different and complementary measure, since its numerical optimization is not constrained through distances to the reference stimulus (while distance constraints can induce \emph{a priori} similarities); thus, it can better function as an unconstrained global characterization of the ``encoding landscape'' (i.e.~multivariate tuning landscape of a population) and estimate if the (reconstructed) stimuli that a deep network is invariant to are ``selective'' (i.e.~visually similar to original stimuli).

\section{Discussion}

%how to intrepret results
{\bf Contributions of this work.}
Comparatively, the characterization of representations in this work is designed to be general and unbiased such that both first- and second-order methods can be supported for single neurons or groups of neurons, without assuming any \emph{de facto} property in the solutions (\eg invariance to parametric deformations).
With the strong statistical significances of results obtained from the large-scale experiments on 100 shallow and 100 deep convolutional networks, in total 6,400 artificial neurons, we argue that our observations toward answering the differences between (shallow \vs deep) and among (deep) representations are highly conclusive, which can be summarized as:
\begin{itemize}[leftmargin=*] % Main Findings
%\setlength\itemsep{0.0ex}
\item Complexity of representations quantitatively increases along with network's depth.
\item Unlike deep representations, which are invariant and selective at the same time, shallow representations are only invariant, not selective, and their capacities of invariance are significantly lower than deep representations'.
\item Both first-order properties of deep representations (TSEA and OTSA) are extremely important to deep network's performance.
\item All second-order properties of deep representations (in particular, ITSSA and STSSA) together are as important as the first-order properties, if not significantly more.
\end{itemize}

\begin{figure}[t]
\centering \includegraphics[width=0.90\textwidth]{Figs/whole_face.pdf}
\caption{{\bf Characterizing face pair matching decisions.}
The best-performing decision function $\mathcal{D}$ (\ie the SVM classifier trained based on representations from the best-performing deep network $f_v$) {was} presented with the out-of-training-set reference stimuli, and the optimal stimulus searches, \ie $\arg\max_{\bf{X}}\mathcal{D}(| f_v\left({\bf{X}}\right) - f_v(\tilde{\bf{X}}) |^\frac{1}{2})$, initialized with the LFW-a average face {were} performed.
The resultant optimal stimuli for the ``match'' decisions, though far from photorealism, actually still well capture certain facial features, especially around the eye regions\protect\footnotemark.
}
\label{fig:wface}
\end{figure}

\footnotetext{Unlike inverting vector representations, inverting the decision (\ie a single top-layer neuron) back into the entire image $\bf{X}$ is significantly harder, and using the average face (which does not ``match'' any of the reference stimuli) significantly helps with the speed and quality of convergence without biasing the results.
Both resultant optimal stimuli here surprisingly lead to very high matching scores, given that they are merely subtlety different from the average face.
Worth to note, the decision function {was} only trained on 6,000 face pairs.
}
%and visually carry similarities 
%Why blurry? not inv. rep. it thinks it's close enough. considering the fact that it is only trained on 6,000 face pairs.

In addition to the representations inside deep networks, we also tried to probe the face pair matching decisions made based upon those representations, as exemplified in Fig.~\ref{fig:wface}, which is reported for the first time too to our best knowledge. Though being qualitative, the results still positively confirm the effectiveness of these deep representations as in the complete face pair matching decisions.

{\bf Future work.}
In this work, we only studied shallow networks and the ``simplest'' type of deep networks, two-level networks.
Our results can potentially generalize if viewed as the characterization of ``adding an extra level.''
However, it is indeed highly desirable to extend the study into even deeper networks.
Even though the methods and measures proposed in this work are neutral to the depth of networks, numerical difficulties (\eg evaluation and convergence speed, quality of solutions, \etc) might need to be addressed in very deep networks.
Beyond explaining what properties make a deep representation good, it would be valuable too to incorporate these findings into learning even better representations (which are almost solely driven by bigger datasets nowadays), \eg, via regularizing the representation measures.  
We believe that this could be useful in accelerating or improving representation learning, particularly when data are scarce and representations are not easily transferable.
Finally, even though a decently high fraction of the variance in deep network's performance has been explained, how to further increase this number, if there is an upper bound (\eg limits from dataset itself), \etc, need to be further investigated.

%Secondly, statistical tests were extensively deployed to support observations reported in this paper, especially in the context of real applications (\ie LFW-a face recognition), which might seem still far-fetched to be fully explainable by mathematical theories.
%Nonetheless, we envision that combining more theoretical analyses which, \eg, have been adopted in various topics in understanding deep representations \cite{delalleau2011shallow, montufar2014number}, with statistical simulations can further empower us in answering other important questions in deep networks \cite{bengio2013deep}.

%%application
% application: where training data is scarse; actively regularize representations
% not deep in this work, but adding another layer(level), what would happen
% not only optimal stimulus; modulate the landscape
% theory works gave some general answers, we want to answer, in real apps, why.
% still, to combine with theory

%\clearpage

{\small
\bibliographystyle{ieee}
\bibliography{ref}
}

\end{document}

### 2015-06-05 nips2015.tex
Edits

@@
 %\title{How Representations Affect Recognition Performances of Deep Networks?}
-\title{What Makes Deep Convolutional Representations Good for Visual Recognition?} %
+\title{What Makes a Deep Convolutional Representations Good for Visual Recognition?} %
 
@@
 \begin{abstract}
-Representations of artificial neurons in deep convolutional networks have been demonstrated to be highly effective for various visual recognition tasks \cite{szegedy2014going, schroff2015facenet, donahue2014decaf}.
-However, our understanding of them arguably remains limited.
-In this paper, we address two important questions: (1) why deep networks outperform shallow networks, and (2) among networks of the same depth, why certain networks outperform the others.
-We considered the case study of face pair matching and conducted a large-scale experiment.
-The experiment examined 200 randomly generated shallow and deep networks, with the results analyzed via a series of measures designed to quantify the differences in representations. 
-We conclude through this principled analysis that deep representations are quantitatively better than shallow representations, and overall 71\% of the variance in the performance of deep networks can be explained with our measures, both of which are confirmed using statistical tests for the first time.
+Deep convolutional networks have been proven to be highly effective across a wide range of visual recognition tasks \cite{szegedy2014going, schroff2015facenet, donahue2014decaf}.
+However, our understanding of these networks work arguably remains limited.
+Here, we examine two fundamental questions: first, why do deep networks outperform shallow networks, and second, among networks of the same depth, why do certain network architectures outperform the others.
+We used face pair matching as a simplified recognition test-bed, and conducted a large-scale experiment in which random shallow and deep networks were generated and analyzed to measure properties of their resulting representations of visual inputs.
+We describe a suite of measures of network selectivity, invariance, and representational quality that are able to account for up to 71\% of the variance in network performance, and we explore how these measures can yield insights into the factors that govern network performance.
 \end{abstract}
@@
 \centering \includegraphics[width=0.40\textwidth]{Figs/perf.pdf} %\vspace{-1.5ex}
-\caption{{\bf What makes a deep network good?}
+\caption{{\bf What makes a given deep network good?}
 Among existing work, researchers mostly tried to produce one ``best'' network architecture (\ie a hyperparameter set, either through manual tuning or automated hyperparameter search), which is subsequently evaluated on a particular benchmark.
-Nevertheless, it is not fully understood why in practice deep networks (\eg two or more levels) on average are better shallow networks (\eg one level), and why certain networks are better than the others, \eg when considering a larger population of networks \cite{cox2011beyond} and their distribution of performance (as shown here) on task like LFW-a face pair matching \cite{wolf2011effective} (see Sec.~\ref{sec:methods} for details of the experimental setup).
+Nevertheless, it is not fully understood why in practice deep networks (\eg two or more levels) on average are better shallow networks (\eg one level), and why certain network architectures are better than the others, \eg when considering a larger population of networks \cite{cox2011beyond} and their distribution of performance (as shown here) on task like LFW-a face pair matching \cite{wolf2011effective} (see Sec.~\ref{sec:methods} for details of the experimental setup).
 \vspace{-0.5ex}}
@@
 
-Although deep convolutional networks have been successfully pushing the boundaries of computer vision by breaking the records of datasets like ImageNet \cite{russakovsky2014imagenet} and LFW \cite{LFWTech} year after year \cite{krizhevsky2012imagenet, sermanet2013overfeat, szegedy2014going, taigman2014deepface, sun2014deep, schroff2015facenet}, our understanding of these networks' properties arguably remains limited, since we cannot yet fully answer the question ``what makes a deep network good?''
-In this paper, we address two practical aspects of this important question: (1) why deep networks perform better than shallow networks, and (2) among networks of the same depth, why certain networks perform better than others (as illustrated in Fig.~\ref{fig:perf}).
-More specifically, from the point of view that successful networks are forming ``better'' representations of visual inputs that are more separable to the final-stage linear classifiers, we want to identify the crucial properties that are required for forming (or learning) better representations.
-We adopted face pair matching (\ie identifying pairs of different pictures from the same person and rejecting those from different persons) as the benchmark task in our experiments for two main reasons: (1) representations of face recognition networks are far less studied than those of object recognition networks, and (2) face images can be easily aligned and extracted from irreverent backgrounds, which can serve as a better source of visual stimuli for studying representations that fundamentally facilitates recognition\footnote{\expwhyface}.
+In recent years, deep convolutional networks have come to dominate many benchmark challenges in computer vision, including datasets such as ImageNet \cite{russakovsky2014imagenet} and LFW \cite{LFWTech} year after year \cite{krizhevsky2012imagenet, sermanet2013overfeat, szegedy2014going, taigman2014deepface, sun2014deep, schroff2015facenet}.
+However, in spite of this impressive performance, our understanding of these networks' properties arguably remains limited, since we cannot yet fully answer the question ``what makes a given deep network good?''
+In this paper, we address two practical aspects of this important question: (1) why do deep networks perform better than shallow networks, and (2) among networks of the same depth, why do certain networks perform better than others (as illustrated in Fig.~\ref{fig:perf}).
+More specifically, from the point of view that successful networks are forming better representations of visual inputs in which object categories are more separable, we want to identify the crucial properties that are required for forming (or learning) better representations.
+We adopted face pair matching (\ie identifying pairs of different pictures from the same person and rejecting those from different persons) as a benchmark task in our experiments, since previous investigations have shown that deep networks perform well in this domain \ref{taigman2014deepface}, and because face images can optionally be easily aligned and extracted from irreverent backgrounds, which can serve as a good source of visual stimuli for studying representations that fundamentally facilitates recognition.
+The techniques presented here, however, are in no way specific to face recognition.
+
+% for two main reasons: (1) representations of face recognition networks are far less studied than those of object recognition networks, and (2) face images can be easily aligned and extracted from irreverent backgrounds, which can serve as a better source of visual stimuli for studying representations that fundamentally facilitates recognition\footnote{\expwhyface}.
 
@@
 {\bf Experimental setup.}
-A simplified variety of deep convolutional neural networks \cite{krizhevsky2012imagenet, sermanet2013overfeat, szegedy2014going}---the Spatio-Temporal Hierarchical Object Representation (STHOR) model \cite{cox2011beyond, sthor}---{was} adopted in our experiments.
-Like other deep networks, it consists of the standard cascade of convolution, nonlinear activation, pooling layers, \etc (which together define a single ``level'' in this work).
-However, its convolution kernel weights are randomly assigned rather than trained.
-Though not fully matching the performances of deep networks trained on massive quantities of data, it can in fact still be quite competitive for a variety of tasks, especially in the regime where large quantities of training samples are not publicly available.
+To enable the rapid probing of a large number of candidate networks, we used simplified random (untrained) deep convolutional neural networks \cite{cox2011beyond, sthor} in our experiments.
+As with other deep networks, these networks consist of the standard cascade of convolution, nonlinear activation, pooling layers, \etc (which together define a single ``level'' in this work), however, the convolution kernel weights are randomly assigned rather than trained.
+Though not fully matching the performances of deep networks trained on massive quantities of data, such models nonetheless surprisingly effective in regimes where limited training data is available \cite{cox2011beyond}.
 %\eg the LFW-a setting we used \cite{krizhevsky2012imagenet} \cite{pinto2009high, cox2011beyond, viglarge}.
-Most importantly, since it does not require training---except for the linear SVM on top of the representations (\ie feature vectors) extracted---a large number of networks with diverse hyperparameters can be rapidly generated and compared, which is particularly useful in this study where relative, instead of absolute task performances, are of primary interest.
+Most importantly, since the networks themselves do not require training---except for a simple linear classifier at the final stage of the network---a large number of networks with diverse architectures and hyperparameters can be rapidly generated and compared, which is particularly useful in this study where relative, instead of absolute task performances, are of primary interest.
 In the experiments, 100 shallow (\ie one-level) networks and 100 deep (\ie two-level) networks {were} randomly generated (except all with 32 top-layer neurons\footnote{\expsettings}) and tested to see how their representations differ with the networks' depths and affect the networks' performances on face pair matching against the LFW-a dataset \cite{wolf2011effective} as shown in Fig.~\ref{fig:perf}\footnote{\expperf}.
@@
 
-\footnotetext{The rest measures which are included in the multiple correlation analysis but not individually reintroduced and shown in Fig.~\ref{fig:corr} are the invariance path potential, selectivity path potential, and invariance subspace capacity of vector representations.
+\footnotetext{The remainder of the measures that are included in the multiple correlation analysis but not individually reintroduced and shown in Fig.~\ref{fig:corr} are the invariance path potential, selectivity path potential, and invariance subspace capacity of vector representations.
 Their individual correlations and significance levels are $-0.31^{\ast\ast}$, $-0.24^{\ast}$, $-0.39^{\ast\ast\ast}$, respectively.
@@
 
-In addition to the representations inside deep networks, we also tried to probe the face pair matching decisions made based upon those representations, as exemplified in Fig.~\ref{fig:wface}, which is reported for the first time too to our best knowledge.
-Though being qualitative, the results still positively confirm the effectiveness of these deep representations as in the complete face pair matching decisions.
+Previous investigations have sought to visualize the optimal stimulus for the decision output of fully trained network \cite{szegendy}, and the methods presented here can also be applied to the final classifier layer in our face matching network.
+Fig.~\ref{fig:wface} shows examples of such a procedure, wherein an optimal ``matching'' stimulus was generated starting from an average face.
+The emergence of distinctive facial features in the optimized image could provide useful information about the features that the network uses to represent matches.
 
 {\bf Future work.}
-In this work, we only studied shallow networks and the ``simplest'' type of deep networks, two-level networks.
-Our results can potentially generalize if viewed as the characterization of ``adding an extra level.''
-However, it is indeed highly desirable to extend the study into even deeper networks.
-Even though the methods and measures proposed in this work are neutral to the depth of networks, numerical difficulties (\eg evaluation and convergence speed, quality of solutions, \etc) might need to be addressed in very deep networks.
+In this work, we studied the simplest forms of convolutional networks--- random ``shallow'' (one-layer) networks and the simplest ``deep'' (two-layer) network.
+However, it is obviously highly desirable to extend such investigations to deeper networks, and to networks that have been trained on large amounts of data.
+Though the methods and measures proposed here are neutral to the depth of networks, numerical challenges (\eg evaluation and convergence speed, quality of solutions, \etc) would need to be addressed in very deep networks.
+
 Beyond explaining what properties make a deep representation good, it would be valuable too to incorporate these findings into learning even better representations (which are almost solely driven by bigger datasets nowadays), \eg, via regularizing the representation measures.

### 2015-06-05 nips2015.tex
A few more tweaks

@@
 (3) \Wrt the OTSA measure, the optimal stimuli of good representations are comparatively more correlated to face images, especially to the linear averages of the most similar (top 100) face images, whereas those of bad representations do not.
-Comparing these measures against the performances of deep networks (see Fig.~\ref{fig:corr}) shows strong correlations, which imply \emph{these first- and second-order properties of good representations must well ``align with'' the distribution of task-related stimuli}, even if they are not explicitly engineered to do so (unlike generative models).
+Comparing these measures against the performances of deep networks (see Fig.~\ref{fig:corr}) shows strong correlations, which implies taht these first- and second-order properties of good representations must ``align with'' the distribution of task-related stimuli, even if the networks were not explicitly trained to have this property.
 }
@@
 % of individual well-performing deep networks.
-In addition, even though these deep convolutional representations are not formulated to be invertible (as opposed to, \eg, autoencoder's), our results further suggest that this property can actually be \emph{required} for a deep network to perform well.
+In addition, even though the deep convolutional representations used here are not formulated to be invertible (as opposed to, \eg, autoencoders), our results further suggest that this property can actually be \emph{required} for a deep network to perform well.
 
@@
 {\bf Invariance and selectivity \vs task-related stimulus subspace alignment.} 
-It is widely known that invariance and selectivity are essential for biological vision \cite{desimone1991face, ito1995size}; however, how these properties directly affect the recognition performances of artificial neural networks remains hypothetical.
+It is widely known that invariance and selectivity are essential for biological vision \cite{desimone1991face, ito1995size}; however, how these properties directly affect the recognition performances of artificial neural networks remains unclear.
 To address this question, starting from randomly sampled reference stimuli, invariance and selectivity path searches {were} performed at $\delta = 0.1\pi$ using Eq.~\ref{eq:I2} and \ref{eq:S2} for vector representations.
@@
 Compared to previous work, our method does not assume that any kind of parametric deformation is important and thus should be studied.
-Instead, the ``intrinsic'' invariant and selective directions of representations are directly searched and compared against the distribution of the task-related stimuli, which is more bias-free.
-Surprisingly, highly natural invariance/selectivity are again observed in well-performing networks even given their nature of randomness.
+Instead, the ``intrinsic'' invariant and selective directions of representations are directly searched and compared against the distribution of the task-related stimuli, which is relatively bias-free.
+Interestingly, high degrees of invariance and selectivity are observed in high-performing networks, even though these networks were not trained to have this property.
 
@@
 Intriguingly, though none of the optimal stimuli is apparently similar to any of the task-related stimuli, when alternatively compared against the averages of the better aligned task-related stimuli (top 1\%), the similarities (in good representations) and dissimilarities (in bad representations) become apparent.
-While it is unclear how to ``optimally place'' an optimal stimulus (or shape the entire landscape) in a high dimensional stimulus space under visual recognition tasks, based on the observation, we hypothesize that it should be closely surrounded by multiple visually similar task-related stimuli (but not directly on any individual task-related stimulus) such that the neuron is more likely to represent those similar stimuli (and only those stimuli) with similar values.
-This may also explain why optimal stimuli often resemble the averages \cite{le2012building} or combinations \cite{simonyan2013deep} of multiple visual objects.
+While it is unclear how to ``optimally place'' an optimal stimulus (or shape the entire landscape) in a high dimensional stimulus space under visual recognition tasks, based on our observations, we hypothesize that it should be closely surrounded by multiple visually similar task-relevant stimuli, such that the neuron is more likely to represent those similar stimuli (and only those stimuli) with similar values.
+This observation may also explain why optimal stimuli often resemble the averages \cite{le2012building} or combinations \cite{simonyan2013deep} of multiple visual objects.
 

### 2015-06-05 nips2015.tex
Fix title
supp

@@
 %\title{How Representations Affect Recognition Performances of Deep Networks?}
-\title{What Makes a Deep Convolutional Representations Good for Visual Recognition?} %
+\title{What Makes a Deep Convolutional Representation Good for Visual Recognition?} %
 
@@
 Deep convolutional networks have been proven to be highly effective across a wide range of visual recognition tasks \cite{szegedy2014going, schroff2015facenet, donahue2014decaf}.
-However, our understanding of these networks work arguably remains limited.
+However, our understanding of how these networks work arguably remains limited.
 Here, we examine two fundamental questions: first, why do deep networks outperform shallow networks, and second, among networks of the same depth, why do certain network architectures outperform the others.
@@
 More specifically, from the point of view that successful networks are forming better representations of visual inputs in which object categories are more separable, we want to identify the crucial properties that are required for forming (or learning) better representations.
-We adopted face pair matching (\ie identifying pairs of different pictures from the same person and rejecting those from different persons) as a benchmark task in our experiments, since previous investigations have shown that deep networks perform well in this domain \ref{taigman2014deepface}, and because face images can optionally be easily aligned and extracted from irreverent backgrounds, which can serve as a good source of visual stimuli for studying representations that fundamentally facilitates recognition.
+We adopted face pair matching (\ie identifying pairs of different pictures from the same person and rejecting those from different persons) as a benchmark task in our experiments, since previous investigations have shown that deep networks perform well in this domain \cite{taigman2014deepface}, and because face images can optionally be easily aligned and extracted from irreverent backgrounds, which can serve as a good source of visual stimuli for studying representations that fundamentally facilitates recognition.
 The techniques presented here, however, are in no way specific to face recognition.
@@
 \begin{figure}[t]
-\centering \includegraphics[width=\textwidth, trim=1.0ex 0.5ex 0 0, clip]{Figs/g_vs_b_repr.pdf} 
+\centering \includegraphics[width=\textwidth, trim=1.4ex 0.5ex 0 0, clip]{Figs/g_vs_b_repr.pdf} 
 \caption{{\bf Visualization of good and bad representations.}
@@
 Figure \ref{fig:allrep2} demonstrates examples of good and bad alignments between the subspaces formed by the invariance/selectivity paths and the task-related stimuli (\ie how likely the invariance and selectivity can benefit recognition), which are defined as the sparsity of $n$ paths projected onto the task-related stimuli's principal component (or here, eigenface) vector space $\bf{V}$, \ie $\frac{1}{n}\sum_{i=1}^{n} \left\| {\bf{Vx}}_{\delta,i} \right\|_{1}$ with $n=20$ and $\delta=0.1\pi$.
-Compared to previous work, our method does not assume that any kind of parametric deformation is important and thus should be studied.
+Compared to some previous work, our method does not assume that any kind of parametric deformation is important and thus should be studied.
 Instead, the ``intrinsic'' invariant and selective directions of representations are directly searched and compared against the distribution of the task-related stimuli, which is relatively bias-free.
-Interestingly, high degrees of invariance and selectivity are observed in high-performing networks, even though these networks were not trained to have this property.
+Surprisingly, highly natural invariance/selectivity are again observed in well-performing networks even given their nature of randomness.
 
@@
 Intriguingly, though none of the optimal stimuli is apparently similar to any of the task-related stimuli, when alternatively compared against the averages of the better aligned task-related stimuli (top 1\%), the similarities (in good representations) and dissimilarities (in bad representations) become apparent.
-While it is unclear how to ``optimally place'' an optimal stimulus (or shape the entire landscape) in a high dimensional stimulus space under visual recognition tasks, based on our observations, we hypothesize that it should be closely surrounded by multiple visually similar task-relevant stimuli, such that the neuron is more likely to represent those similar stimuli (and only those stimuli) with similar values.
+While it is unclear how to ``optimally place'' an optimal stimulus (or shape the entire landscape) in a high dimensional stimulus space under visual recognition tasks, based on our observations, we hypothesize that it should be closely surrounded by multiple visually similar task-related stimuli, such that the neuron is more likely to represent those similar stimuli (and only those stimuli) with similar values.
 This observation may also explain why optimal stimuli often resemble the averages \cite{le2012building} or combinations \cite{simonyan2013deep} of multiple visual objects.
@@
 
-\begin{figure}[t]
-\centering \includegraphics[width=0.90\textwidth]{Figs/whole_face.pdf}
-\caption{{\bf Characterizing face pair matching decisions.}
-The best-performing decision function $\mathcal{D}$ (\ie the SVM classifier trained based on representations from the best-performing deep network $f_v$) {was} presented with the out-of-training-set reference stimuli, and the optimal stimulus searches, \ie $\arg\max_{\bf{X}}\mathcal{D}(| f_v\left({\bf{X}}\right) - f_v(\tilde{\bf{X}}) |^\frac{1}{2})$, initialized with the LFW-a average face {were} performed.
-The resultant optimal stimuli for the ``match'' decisions, though far from photorealism, actually still well capture certain facial features, especially around the eye regions\protect\footnotemark.
-}
-\label{fig:wface}
-\end{figure}
-
-\footnotetext{Unlike inverting vector representations, inverting the decision (\ie a single top-layer neuron) back into the entire image $\bf{X}$ is significantly harder, and using the average face (which does not ``match'' any of the reference stimuli) significantly helps with the speed and quality of convergence without biasing the results.
-Both resultant optimal stimuli here surprisingly lead to very high matching scores, given that they are merely subtlety different from the average face.
-Worth to note, the decision function {was} only trained on 6,000 face pairs.
-}
-%and visually carry similarities 
-%Why blurry? not inv. rep. it thinks it's close enough. considering the fact that it is only trained on 6,000 face pairs.
-
-Previous investigations have sought to visualize the optimal stimulus for the decision output of fully trained network \cite{szegendy}, and the methods presented here can also be applied to the final classifier layer in our face matching network.
-Fig.~\ref{fig:wface} shows examples of such a procedure, wherein an optimal ``matching'' stimulus was generated starting from an average face.
-The emergence of distinctive facial features in the optimized image could provide useful information about the features that the network uses to represent matches.
-
 {\bf Future work.}

### 2015-06-05 Figs_supp/reformat.sh

@@ -1,0 +1,4 @@
+for img in `ls *crop.pdf`;do
+	echo $img
+	convert -density 200x200 -quality 90 -compress jpeg $img $img
+done

### 2015-06-05 nips2015.tex
Some tweaks

--- 2015-06-05 nips2015.tex
@@
 \newcommand{\expperf}{10-fold cross validation accuracies using the LFW view 2 splits (\cite{LFWTech}, totally 6,000 face pairs) are reported.
-Although it is highly desirable to characterize representations from the state-of-the-art face pair matching networks \cite{taigman2014deepface, sun2014deep, schroff2015facenet}, neither those trained networks nor their datasets are publicly available.
-Nonetheless, under the hard protocol adopted here (\ie image-restricted, no outside data, data augmentation, or multiple-crop evaluation), the (well-performing) random networks remain competitive against trained networks and do provide good representations in terms of accuracies attained.
+We stress that differential comparison of performance among models, rather than absolute performance was the goal of the present work.
+Although it would be desirable to also characterize representations from the state-of-the-art face pair matching networks \cite{taigman2014deepface, sun2014deep, schroff2015facenet}, neither the trained networks nor their datasets are publicly available.
+Nonetheless, under the challenging protocol adopted here (\ie image-restricted, no outside data, no data augmentation, or multiple-crop evaluation), the highest-performing among the random networks still can achieve reasonable levels of performance (esp.
+if more layers are used) and thus we argue they are useful objects of study.
 } 

### 2015-06-05 nips2015_supp.tex

\documentclass{article} % For LaTeX2e
\usepackage{nips15submit_e,times}
\usepackage{hyperref}
\usepackage{url}

\usepackage{graphicx}
\usepackage[labelfont=bf]{caption}
\usepackage{float}

\usepackage{amsmath}
\usepackage{amssymb}

\usepackage{xspace}

\makeatletter
\DeclareRobustCommand\onedot{\futurelet\@let@token\@onedot}
\def\@onedot{\ifx\@let@token.\else.\null\fi\xspace}
\def\eg{\emph{e.g}\onedot} \def\Eg{\emph{E.g}\onedot}
\def\ie{\emph{i.e}\onedot} \def\Ie{\emph{I.e}\onedot}
\def\cf{\emph{c.f}\onedot} \def\Cf{\emph{C.f}\onedot}
\def\etc{\emph{etc}\onedot} \def\vs{\emph{vs}\onedot}
\def\wrt{w.r.t\onedot} \def\Wrt{W.r.t\onedot} \def\dof{d.o.f\onedot}
\def\etal{\emph{et al}\onedot}
\makeatother

\setcounter{figure}{0}
\renewcommand{\thefigure}{S\arabic{figure}}

\title{What Makes a Deep Convolutional Representation Good for Visual Recognition?\\\emph{Supplementary Material}}

\author{
David S.~Hippocampus\\
Department of Computer Science\\
Cranberry-Lemon University\\
\texttt{hippo@cs.cranberry-lemon.edu} \\
}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to \LaTeX{} to determine where to break
% the lines. Using \AND forces a linebreak at that point. So, if \LaTeX{}
% puts 3 of 4 authors names on the first line, and the last on the second
% line, try using \AND instead of \And before the third author name.

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

%\nipsfinalcopy % Uncomment for camera-ready version

\begin{document}

\maketitle
\vspace{-2.0ex}

\tableofcontents
\clearpage

\section{Methods}

\subsection{Numerical Optimization and Regularization}

The CMA-ES (Covariance Matrix Adaptation Evolution Strategy) numerical solver \cite{hansen2001completely} adopted in this work treats networks as non-analytical black boxes---it makes no assumption about networks' structures or properties (\eg we do not assume that analytical gradients are available, even though they do exist in most artificial networks). 
It stochastically estimates numerical derivatives (\ie inverse Hessian), and worked reasonably well in our experiments.
This strategic decision {was} made in order to allow the proposed methods to transfer over to experiments with biological neurons, where only the outputs of neurons are available. While no specific knowledge of the network itself is assumed, we do restrict the space of stimuli considered, based on prior knowledge of biological and artificial neuronal response properties. In particular, all stimuli considered in this work follow a constant energy constraint $\left\| \bf{x} \right\| = E$, complying with the general observation that neurons are often modulated by stimulus contrasts \cite{albrecht1982striate, cheng1994comparison}, but this modulation is less interesting when considering pattern selectivity of neurons. Limiting the stimulus search space in this manner effectively reduces the range of possible stimuli to consider, and avoids degenerate solutions that simply maximize stimulus contrasts.
In this work, the distance constraint $\delta$ only goes up to $\frac{\pi}{2}$, since on the $N$ dimensional sphere, stimuli fall in $\frac{\pi}{2} < \delta \le \pi$ are simply ``negatives'' of those in $0 < \delta \le \frac{\pi}{2}$, which are of less uniqueness and interest; nevertheless, going up to the full range $0 < \delta \le \pi$ is numerically supported.

The fitness landscapes of both shallow and deep representations are generally non-convex, therefore there is no guarantee that optimality can be achieved through any optimization method.
However, by carefully adjusting the numerical solver and monitoring the quality of solutions, we argue that informative local optima can still be discovered and used to characterize the representations.
Specifically, all numerical optimizations {were} executed multiple times from different random initial stimuli, for increasing quality of the numerical solutions (of optimal stimuli and invariance and selectivity paths, with 2 runs executed and the better kept), and at the same time, for providing statistical samplings of the solution spaces (of encoding accuracies and invariance/selectivity subspace properties, with 10 and 20 runs executed).

In the main paper, for inverting deep vector representations, even though Eq.~2  describes the essence of the approach, it was not able to produce analyzable numerical results alone, due to the fact that deep representations can be highly invariant to noises that do not resemble natural images.
Implementation-wise, it was further regularized with the energy of Laplacian filtering $\left\| \mathcal{L} \left( \bf{x} \right) \right\|$ to enforce smoothness and similarity to natural images.
A similar concept was also adopted in \cite{mahendran2014understanding}.

\subsection{Permutation Tests}

\subsubsection{Permutation Test for Significance of Difference between Two Distributions}

Considering two distributions $A$ and $B$, we can test how likely these two distributions are significantly different, rather than actually being from the same distribution (the null hypothesis), using a permutation test as follows. The original difference $d$ between $A$ and $B$ is first calculated, then a large number of trials are performed. Within each trial, $A'$ and $B'$ (of the same size as $A$ and $B$) are formed by random sampling from $A \cup B$ (without replacement), and a new $d'$ is calculated. The $p$-value of the null hypothesis is then the probability of $d' > d$ among all trials. In this work, $A$ and $B$ are measure values from the shallow and deep networks, respectively.

\subsubsection{Permutation Test for Significance of Correlation between Two Sequences}

Considering two sequences $a$ and $b$, we can test how likely these two sequences are significantly correlated, rather than actually not being correlated (the null hypothesis), using a permutation test as follows. The original correlation $c$ between $a$ and $b$ is first calculated, then a large number of trials are performed. Within each trial, $a'$ is formed by randomly shuffling $a$, and a new $c'$ is calculated between $a'$ and $b$. The $p$-value of the null hypothesis is then the probability of $c' > c$ among all trials.
In this work, $a$ and $b$ are measure values and performances of deep networks, respectively.

%\subsubsection{Permutation Test for Significance of Correlation between Two Variables}

%\subsection{Dataset Properties} % eigenface details

\clearpage

\section{Results: Shallow \vs Deep Representations}
\label{sec2}

More data related to the optimal stimulus, invariance and selectivity paths, and invariance subspace are presented in the following figure.

\begin{figure}[H]
\centering \includegraphics[width=0.90\textwidth]{Figs_supp/pic1.pdf} 
\caption{{\bf Visualization of more shallow and deep representations.}
Color of the boarder of an image indicates the fitness of the representation (\ie response of the neuron) driven by the image, with color map following the definition in the main paper.
A and E show the search trajectories of \emph{optimal stimuli} of all 32 neurons from the best performing shallow and best performing deep networks, respectively, where A.1 and E.1 show the initial $1/f$ random stimuli, A.5 and E.5 the resultant optimal stimuli, and A.2--4 and E.2--4 three intermediate steps corresponding to the three largest curvatures in the fitness histories.
B and F show the \emph{invariance paths} of shallow and deep neurons, starting from the corresponding optimal stimuli show in A.5 and E.5, moving from $0.1\pi$ (B.1 and F.1) to $0.5\pi$ (B.5 and F.5) away accordingly.
C and G show the \emph{selectivity paths} of shallow and deep neurons, with definitions following B and F.
D and H show the \emph{invariance subspaces} sampled $0.1\pi$ away from the optimal stimuli, with 5 out of 20 results randomly selected.
} % do visually
\end{figure}

For optimal stimulus searches, in order to further increase the searching speed, the initial stimulus {was} selected from 1000 ${1}/{f^{\alpha}}$ random stimuli with the best fitness, where $\alpha \in \left\lbrace -4,-3,-2,-1,0 \right\rbrace$, without overly sacrificing the nature of random initialization.
From the results, first, the response landscapes of deep neurons are noticeably more complex than those of shallow neurons, as a large fraction of search trajectories for deep neurons actually go down then up before reaching the (local) optima, which also suggests the numerical solver can handle non-convexity reasonably well.

%\clearpage

\subsection{Invariance and Selectivity Path Potential}
More data related to the invariance and selectivity path potentials are presented in the following figure.

\begin{figure}[H]
\centering \includegraphics[width=0.60\textwidth]{Figs_supp/e_fig3a_2-nup-crop.pdf} 
\caption{{\bf Fitness-distance analyses of shallow and deep representations with random walk results.} %scalar
In addition to the invariance and selectivity paths, the random perturbation based characterization (\ie random walk results as green dots) of the landscape (\cite{jones1995fitness}, as the original method) was also adopted in our experiments.
A random walk direction can be defined as $f\left(\cos\left(\delta\right)\hat{\bf{x}} + \sin\left(\delta\right)\tilde{\bf{x}}\right)$ where $\tilde{\bf{x}}$ is any random stimulus such that $\left\| \tilde{\bf{x}} \right\| = 1$ and $\left\langle \hat{\bf{x}},\tilde{\bf{x}} \right\rangle = 0$, which can be obtained through random projection of $\mathrm{Null}\left(\hat{\bf{x}}\right)$.
1,000 random walks were executed on each shallow and deep neurons.
The results indicate random walks are not as effective as the numerically optimized results (\ie the invariance and selectivity paths) at characterizing the dynamic range of neurons. 
Nevertheless, comparing the gaps between the random-walk and selectivity curves in A and B also points out that the selectivity of deep neurons is in fact way more ``selective'' (\ie not noise-like) than of shallow neurons.
}
\label{fig:fda}
\end{figure}

\clearpage

\section{Results: Good \vs Bad Representations}

\subsection{Task-related Stimulus Encoding Accuracy}
%Encoding accuracy as a single measure can best explain the deep network's performance. Though bearing strong similarity to the invariance path search in terms of mathematical formulation, encoding accuracy is in fact a radically different and complementary measure, since its numerical optimization is not constrained through distances to the reference stimulus (while distance constraints can induce \emph{a priori} similarities); thus, it can better function as an unconstrained ``global'' characterization of the ``encoding landscape'' and estimate if the (inverted) stimuli that a deep network is invariant to are ``selective'' (\ie visually similar to original stimuli).
More related data are presented in the following 2 figures.

\begin{figure}[H]
\centering \includegraphics[width=0.55\textwidth]{Figs_supp/pic4.pdf} 
\caption{{\bf More examples of task-related stimulus encoding accuracy.}
Utilizing 16 face images in A as reference stimuli, 3 examples of inverted stimuli from deep networks corresponding to the best, second best, third best and worst average qualities (SSIM) of inverted stimuli are shown in B--E respectively.
Color of the boarder of an image indicates the SSIM between the image and reference stimulus, with color map following the definition in the main paper.
Though the inverted stimuli are in general of lower spatial frequencies, some of the better ones do capture the spatial structures of reference stimuli decently well (\eg stimulus 1, 5, \etc).
} % (reconstructed)
\label{refstim1}
\end{figure}

\begin{figure}[H]
\centering \includegraphics[width=0.60\textwidth]{Figs_supp/e_fig6b-crop.pdf}
\caption{
{\bf Encoding accuracies \vs performances of deep networks.} Each distribution from left to right shows the SSIM scores of the inverted stimuli of vector representations of a deep network against all the 16 reference stimuli (as in Fig.~\ref{refstim1}) in all 10 runs. Means of distributions are plotted as gray circles and the linear regression as dashed black line. %Significance of the slope of all means under permutation test has $p < 0.001$.
}
\end{figure}

\clearpage

\subsection{Invariance and Selectivity \vs Task-related Stimulus Subspace Alignment}
More data related to the subspace alignment measure are presented in the following 2 figures.

\begin{figure}[H]
\centering \includegraphics[width=0.55\textwidth]{Figs_supp/pic3.pdf} 
\caption{{\bf More examples of alignments between subspaces of invariance and selectivity paths and task-related stimuli.}
Using 16 face images in A as reference stimuli, the invariance paths at $\delta = 0.1\pi$ corresponding to the best alignments against task-related stimuli (\ie high $L^1$ sparsities of images represented in eigenface coordinates) among all deep networks are shown in B.1, and similarly, best alignments of selectivity paths in C.1, worst alignments of invariance paths in D.1, and worst alignments of selectivity paths in E.1.
Differences between A and B.1--E.1 are shown in B.2--E.2 for clearer visualization.
Color of the boarder of an image indicates the alignment of the image against task-related stimuli, with color map following the definition in the main paper.
While good alignments (B.1 and C.1) appear more natural (like structural deformations, lighting changes, \etc), bad alignments are mostly meaningless noises.
} %of subspaces % which are known to be important factors in visual recognition % accordingly
%Multiple runs of invariance and selectivity path searches are executed to form samples of the subspaces.
\label{refstim2}
\end{figure}

\begin{figure}[H]
\centering \includegraphics[width=0.60\textwidth]{Figs_supp/e_fig5d-crop.pdf}
\caption{ 
{\bf Invariance and selectivity subspace alignments \vs performances of deep networks.} Each red distribution from left to right shows the invariance subspace alignments against the task-related stimuli in vector representations of a deep network given the 16 reference stimuli (as in Fig.~\ref{refstim2}) in all 20 runs. Alignment scores are normalized by subtracting the corresponding reference stimuli's own alignment scores. Means of distributions are plotted as red circles and the linear regression as dashed red line. Blue distributions for selectivity subspace alignments follow the same definitions.
%Significances of slopes of means under permutation tests both have $p < 0.001$.
}
\end{figure}

\clearpage

\subsection{Optimal \vs Task-related Stimulus Alignment}
More data related to the optimal \vs task-related stimulus alignment measure are presented in the following 2 figures.

\begin{figure}[H]
\centering \includegraphics[width=0.90\textwidth]{Figs_supp/pic2.pdf} 
\caption{{\bf More examples of optimal \vs task-related stimulus alignment.}
Results from deep networks of the best 3 average alignment scores (A--C) and the worst average alignment score (D) are visualized.
Within each panel, the first row shows all 32 neurons' optimal stimuli from the network, the third to fifth rows show face images with the top 3 alignment scores (\ie inner-product distances between the images and the corresponding optimal stimuli), and the second row shows the average of images with top 100 alignment scores, out of 10,000 natural images.
Color of the boarder of an image indicates the alignment score of the image, with color map following the definition in the main paper.
}
\end{figure}

\begin{figure}[H]
\centering \includegraphics[width=0.60\textwidth]{Figs_supp/e_fig4b-crop.pdf} 
\caption{ 
{\bf Optimal stimulus alignments \vs performances of deep networks.} Each distribution from left to right shows the top 1\% alignment scores of all the 32 top-layer scalar representations of a deep network against the task-related stimuli. Means of distributions are plotted as gray circles and the linear regression as dashed black line.
%Significance of slope of means under permutation test has $p < 0.001$.
}
\end{figure}

\clearpage

\subsection{Invariance and Selectivity Path Potential of Vector Representations}
Since no baseline can be defined for vector representations as for scalar representations , the invariance (and selectivity, similarly) path potential is redefined as path integral on the fitness landscape $\int_{0}^{\frac{\pi}{2}}{\exp\left(-\left\|f\left(\bf{x}^{+}_{\delta}\right)-f\left(\hat{\bf{x}}\right)\right\|\right)}\mathrm{d}\delta \mathbin{/} {\frac{\pi}{2}}$.
More data related to path potential of vector representations are presented in the following figure.

\begin{figure}[H]
\centering \includegraphics[width=0.60\textwidth]{Figs_supp/e_fig5a-crop.pdf}
\caption{ 
{\bf Fitness-distance analysis of deep vector representations.} Red, blue, and green dots correspond to results of invariance and selectivity path searches, and random walks, starting from all the 16 reference stimuli (as in Fig.~\ref{refstim1}/\ref{refstim2}), where brighter shades of a color are from better performing networks, and darker shades from poorer performing networks. Means of results are plotted as solid lines in corresponding colors. As visualized, correlations between invariance and selectivity path potentials and performance are weak. Random walk results \vs performance, on the other hand, has $\rho = -0.43$ correlation which may arise from the differences in sensitivities to noises, though incorporating it into representation measures does not improve the final multiple correlation.}
\end{figure}

\subsection{Invariance Subspace Capacity of Vector Representations}
More data related to the invariance subspace capacity of vector representations are presented in the following figure.

\begin{figure}[H]
\centering \includegraphics[width=0.60\textwidth]{Figs_supp/e_fig5b-crop.pdf}
\caption{ 
{\bf Invariance subspace capacities \vs performances of deep networks.} Each distribution from left to right shows the invariance subspace capacities of vector representations of a deep network given the 16 reference stimuli (as in Fig.~\ref{refstim1}/\ref{refstim2}). Means of distributions are plotted as gray circles and the linear regression as dashed black line.
%Significance of slope of means under permutation test has $p < 0.001$.
}
\end{figure}

\clearpage

\subsection{Overall Comparison}

More data related to the measures used but not presented in the main paper are presented in the following figure.

\begin{figure}[H]
\centering \includegraphics[width=0.65\textwidth, trim=25ex 27ex 0 0, clip]{Figs_supp/e_fig7_scatter.pdf}
\caption{ 
{\bf Invariance and selectivity path potentials (INPP/SLPP), and invariance subspace capacity (INSC) of vector representations \vs performance of networks.} Spearman's rank correlations and their significance levels of INPP, SLPP and INSC are $-0.31^{\ast\ast}$, $-0.24^{\ast}$, $-0.39^{\ast\ast\ast}$, respectively.
Invariance and selectivity path potentials and invariance subspace capacity for deep vector representations, compared to invariance and selectivity subspace alignments, have lower correlations to network's performance (though still significant), possibly due to their nature of being less task-related than subspace alignment measures (INPP, SLPP, and INSC are measures of search results \emph{started} from task-related images, but not \emph{ended} compared against them, unlike ITSA and STSA).
Vector representation's INSC is negatively correlated to network's performance, mainly due to the fact that poorer performing network's invariance paths are usually noisier (as depicted in Fig.~\ref{refstim2}), and thus should be interpreted differently from scalar representation's INSC.
Task-unrelated measures, \ie OSSC, INPP, SLPP, and INSC of scalar representations as used in Sec.~\ref{sec2}, are significantly worse at explaining a network's performance as one would expect.
}
\end{figure}

As a comparison, Pearson's correlations of TSEA, ITSSA, STSSA, OTSA, INPP, SLPP, INSC, and all measures together, are $0.63$, $-0.41$, $-0.48$, $0.61$, $-0.13$, $-0.14$, $-0.43$, and $0.83$, respectively, which are basically consistent with the Spearman's rank correlations.

%(altogether with linear multiple correlation of only $0.59$, and when combined with the other task-unrelated measures, only bring up the correlation marginally from $0.84$ to $0.86$),

\section{Characterizing Face Pair Matching Decisions}

\begin{figure}[H]
\centering \includegraphics[width=\textwidth]{Figs/whole_face.pdf}
\caption{{\bf Characterizing face pair matching decisions.}
The best-performing decision function $\mathcal{D}$ (\ie the SVM classifier trained based on representations from the best-performing deep network $f_v$) {was} presented with the out-of-training-set reference stimuli, and the optimal stimulus searches, \ie $\arg\max_{\bf{X}}\mathcal{D}(| f_v\left({\bf{X}}\right) - f_v(\tilde{\bf{X}}) |^\frac{1}{2})$, initialized with the LFW-a average face {were} performed.
The resultant optimal stimuli for the ``match'' decisions, though far from photorealism, actually still well capture certain facial features, especially around the eye regions.
}
\label{fig:wface}
\end{figure}

Previous investigations have sought to visualize the optimal stimulus for the decision output of fully trained network \cite{zeiler2014visualizing, simonyan2013deep}, and the methods presented here can also be applied to the final classifier layer in our face matching network.  Fig.~\ref{fig:wface} shows examples of such a procedure, wherein an optimal ``matching'' stimulus was generated starting from an average face. The emergence of distinctive facial features in the optimized image could provide useful information about the features that the network uses to represent matches.

Unlike inverting vector representations, inverting the decision (\ie a single top-layer neuron) back into the entire image $\bf{X}$ is significantly harder, and using the average face (which does not ``match'' any of the reference stimuli) significantly helps with the speed and quality of convergence without biasing the results.
Both resultant optimal stimuli here surprisingly lead to very high matching scores, given that they are merely subtlety different from the average face.
Worth to note, the decision function {was} only trained on 6,000 face pairs.

%and visually carry similarities 
%Why blurry? not inv. rep. it thinks it's close enough. considering the fact that it is only trained on 6,000 face pairs.

{\small
\bibliographystyle{ieee}
\bibliography{ref_supp}
}

\end{document}

### 2015-06-05 nips2015.tex
final

@@
 Here, we examine two fundamental questions: first, why do deep networks outperform shallow networks, and second, among networks of the same depth, why do certain network architectures outperform the others.
-We used face pair matching as a simplified recognition test-bed, and conducted a large-scale experiment in which random shallow and deep networks were generated and analyzed to measure properties of their resulting representations of visual inputs.
-We describe a suite of measures of network selectivity, invariance, and representational quality that are able to account for up to 71\% of the variance in network performance, and we explore how these measures can yield insights into the factors that govern network performance.
+We used face pair matching as a simplified recognition test bed, and conducted a large-scale experiment in which random shallow and deep networks were generated and analyzed to measure properties of their resulting representations of visual inputs.
+We describe a suite of measures of network invariance, selectivity, and representational quality that are able to account for up to 71\% of the variance in network performance, and we explore how these measures can yield insights into the factors that govern network performance.
 \end{abstract}
@@
 More specifically, from the point of view that successful networks are forming better representations of visual inputs in which object categories are more separable, we want to identify the crucial properties that are required for forming (or learning) better representations.
-We adopted face pair matching (\ie identifying pairs of different pictures from the same person and rejecting those from different persons) as a benchmark task in our experiments, since previous investigations have shown that deep networks perform well in this domain \cite{taigman2014deepface}, and because face images can optionally be easily aligned and extracted from irreverent backgrounds, which can serve as a good source of visual stimuli for studying representations that fundamentally facilitates recognition.
+We adopted face pair matching (\ie identifying pairs of different pictures from the same person and rejecting those from different persons) as a benchmark task in our experiments, since previous investigations have shown that deep networks perform well in this domain \cite{taigman2014deepface, sun2014deep, schroff2015facenet}, and because face images can optionally be easily aligned and extracted from irreverent backgrounds, which can serve as a good source of visual stimuli for studying representations that fundamentally facilitates recognition.
 The techniques presented here, however, are in no way specific to face recognition.
@@
 Although it would be desirable to also characterize representations from the state-of-the-art face pair matching networks \cite{taigman2014deepface, sun2014deep, schroff2015facenet}, neither the trained networks nor their datasets are publicly available.
-Nonetheless, under the challenging protocol adopted here (\ie image-restricted, no outside data, no data augmentation, or multiple-crop evaluation), the highest-performing among the random networks still can achieve reasonable levels of performance (esp.
-if more layers are used) and thus we argue they are useful objects of study.
+Nonetheless, under the challenging protocol adopted here (\ie image-restricted, no outside data, no data augmentation, or multiple-crop evaluation), the highest-performing among the random networks still can achieve reasonable levels of performance (especially if more layers are used) and thus we argue they are useful objects of study.
 } 
@@
 To enable the rapid probing of a large number of candidate networks, we used simplified random (untrained) deep convolutional neural networks \cite{cox2011beyond, sthor} in our experiments.
-As with other deep networks, these networks consist of the standard cascade of convolution, nonlinear activation, pooling layers, \etc (which together define a single ``level'' in this work), however, the convolution kernel weights are randomly assigned rather than trained.
-Though not fully matching the performances of deep networks trained on massive quantities of data, such models nonetheless surprisingly effective in regimes where limited training data is available \cite{cox2011beyond}.
+As with other deep networks, these networks consist of the standard cascade of convolution, nonlinear activation, pooling layers, \etc (which together define a single ``level'' in this work); however, the convolution kernel weights are randomly assigned rather than trained.
+Though not fully matching the performances of deep networks trained on massive quantities of data, such models nonetheless are surprisingly effective in regimes where limited training data is available \cite{cox2011beyond, yamins2014performance}.
 %\eg the LFW-a setting we used \cite{krizhevsky2012imagenet} \cite{pinto2009high, cox2011beyond, viglarge}.
@@
 As further demonstrated in the fitness-distance diagrams \cite{jones1995fitness} (Fig.~\ref{fig:fda}), on average, although shallow neurons have good invariance (\ie red curve stays high), they do not have any selectivity (\ie blue curve drops slow), while deep neurons show both good invariance and selectivity (\ie red curve stays high and blue curve drops fast).
-For scalar representations, we can define the invariance (and selectivity, similarly) path potential as the area sandwiched between the invariance curve and the baseline curve (\ie the invariance curves of a single inner-product neuron\footnote{\defbaseline}---the ``baseline'' of all types of neural networks) on the diagram, \ie $\int_{0}^{0.5\pi} \left| \cos^{-1}(f_s(\bf{x}^{+}_{\delta} )) - \delta \right| \mathrm{d}\delta$ (in the $\cos^{-1}$ domain for equalizing both types of path potentials), which measures how invariant (and selective) a neuron is compared to the baseline.
+For scalar representations, we can define the invariance (and selectivity, similarly) path potential as the area sandwiched between the invariance curve and the baseline curve (\ie the invariance curves of a single inner-product neuron\footnote{\defbaseline}---the ``baseline'' of all types of neural networks) on the diagram, \ie $\int_{0}^{\frac{\pi}{2}} \left| \cos^{-1}(f_s(\bf{x}^{+}_{\delta} )) - \delta \right| \mathrm{d}\delta / \frac{\pi}{2}$ (in the $\cos^{-1}$ domain for equalizing both types of path potentials), which measures how invariant (and selective) a neuron is compared to the baseline.
 %/ {\frac{\pi}{2}}
@@
 {\bf Future work.}
-In this work, we studied the simplest forms of convolutional networks--- random ``shallow'' (one-layer) networks and the simplest ``deep'' (two-layer) network.
+In this work, we studied the simplest forms of convolutional networks---random ``shallow'' (one-layer) networks and the simplest ``deep'' (two-layer) network.
 However, it is obviously highly desirable to extend such investigations to deeper networks, and to networks that have been trained on large amounts of data.
@@
 
-{\small
+{%\small
 \bibliographystyle{ieee}

### 2015-06-05 nips2015_supp.tex
final

@@
 
-\subsection{Permutation Tests}
+\subsection{Permutation Tests \cite{fisher1935design, pitman1938significance}}
 
@@
 Considering two distributions $A$ and $B$, we can test how likely these two distributions are significantly different, rather than actually being from the same distribution (the null hypothesis), using a permutation test as follows.
-The original difference $d$ between $A$ and $B$ is first calculated, then a large number of trials are performed.
+The original difference $d$ between $A$ and $B$ is first calculated, then a large number of trials are performed (1,000 in our experiments).
 Within each trial, $A'$ and $B'$ (of the same size as $A$ and $B$) are formed by random sampling from $A \cup B$ (without replacement), and a new $d'$ is calculated.
@@
 Considering two sequences $a$ and $b$, we can test how likely these two sequences are significantly correlated, rather than actually not being correlated (the null hypothesis), using a permutation test as follows.
-The original correlation $c$ between $a$ and $b$ is first calculated, then a large number of trials are performed.
+The original correlation $c$ between $a$ and $b$ is first calculated, then a large number of trials are performed (1,000 in our experiments).
 Within each trial, $a'$ is formed by randomly shuffling $a$, and a new $c'$ is calculated between $a'$ and $b$.
@@
 
-\section{Characterizing Face Pair Matching Decisions}
+\section{Results: Characterizing Face Pair Matching Decisions}
 
